### 2022-04-25 paper/main.tex
added python package, paper outline

\documentclass[11pt]{extarticle}

\title{Assessing the Performance of Matching Methods in Observational Studies}
\author{Jack Potrykus}
\date{\today}

\input{preamble.tex}

\begin{document}

\maketitle

\section*{Abstract}
\tableofcontents
\newpage

\section{Introduction}

\section{Literature Review}

\subsection{Measuring Similarity}

\subsubsection{Propensity Scoring}

Rosenbaum and Rubin, 1983
\begin{itemize}
  \item The \emph{propensity score} is most often accredited to \ldots
  \item They offer a more general definition of a score: anything affording conditional independence
  \item Score seeks to replicated randomized trial
  \item Review of properties (particularly: when can we make unbiased estimates of ATE)
  \item Proposed use cases
\end{itemize}

Methods for calculating propensity scores: Garrido et al, 2014
Note: Blocking-based balance metrics

\paragraph{Extensions}
\begin{itemize}
  \item (Optimal) Caliper width: Austin 2011
  \item Prognostic score: Hansen
  \item Miettinen score is the root of the above
  \item Joint use of Prognostic, Propensity, + Mahalanobis, Leacy and Stuart 2014
  \item Imai and Van Dyk (make sure to include)
\end{itemize}

\subsubsection{(Coarsened) Exact Matching}

\begin{itemize}
  \item Iacus King Porro, 2011: MIB methods
  \item Iacus King Porro, 2012: Causal inference without balance checking
\end{itemize}

\subsection{Balance Assessment}

\subsection{Matching Algorithms}

\section{Experiments}

\subsection{Data Generation}

Papers with data generation:
\begin{itemize}
  \item Austin 2011 (optimal caliper widths)
  \item Stuart et al: 2018
\end{itemize}

\section{Results}

\section{Discussion}

\section{Conclusion}

\end{document}

### 2022-04-25 matching/_typing/sklearn_types.py

@@ -1,0 +1,39 @@
+from typing import Any, Protocol, runtime_checkable
+
+
+@runtime_checkable
+class _Fittable(Protocol):
+    """Anything with a `fit()` method. Most `sklearn` estimators satisfy this"""
+
+    def fit(self, X, y) -> Any:
+        ...
+
+
+@runtime_checkable
+class _Regressor(_Fittable, Protocol):
+    """Any `_Fittable` with a `predict` method"""
+
+    def predict(self, X) -> Any:
+        ...
+
+
+@runtime_checkable
+class _Classifier(_Regressor, Protocol):
+    """Any `_Fittable` with a `predict_proba` method"""
+
+    def predict_proba(self, X) -> Any:
+        ...
+
+
+@runtime_checkable
+class _Preprocessor(_Fittable, Protocol):
+    """Any `_Fittable` with `transform` and `fit_transform` methods"""
+
+    def transform(self, X) -> Any:
+        ...
+
+    def fit_transform(self, X, y=None, **fit_params) -> Any:
+        ...
+
+    def inverse_transform(self, X) -> Any:
+        ...

### 2022-04-25 matching/dataset.py

@@ -1,0 +1,196 @@
+from dataclasses import dataclass
+from typing import Hashable, Iterator, NamedTuple, Optional, Iterable, TypeVar, Union
+
+from numpy import typing as npt
+import numpy as np
+import pandas as pd
+
+
+class _MatchingDatasetNDArrays(NamedTuple):
+    """Container for the NDArray equivalent of each element of a :class:`MatchingDataset`.
+    Note that non-numeric columns of ``X`` are one-hot encoded.
+
+    Attributes
+    ----------
+    X: ndarray
+        of features, with non-numeric columns one-hot encoded with ``drop_first=True``
+    z: ndarray
+        of boolean treatment assignments
+    ids: ndarray
+        of string IDs
+    """
+
+    X: np.ndarray
+    z: np.ndarray
+    ids: np.ndarray
+
+
+"""TypeVar for :class:`MatchingDataset`, for methods which return another :class:`MatchingDataset`"""
+_MatchingDataset = TypeVar("_MatchingDataset", bound="MatchingDataset")
+
+
+@dataclass(init=False, eq=False)
+class MatchingDataset:
+    """Container that validates inputs to be used in bipartite matching
+
+    Attributes
+    ----------
+    X: DataFrame
+        of features to match on
+    z: Series
+        of boolean treatment assignments
+    ids: Series
+        of observation IDs. Note that these are coerced to ``str``
+    """
+
+    X: pd.DataFrame
+    z: pd.Series
+    ids: pd.Series
+
+    def __init__(
+        self,
+        X: npt.ArrayLike,
+        z: npt.ArrayLike,
+        ids: Optional[npt.ArrayLike] = None,
+        treatment_encoded_as: Hashable = True,
+    ):
+        """
+        Parameters
+        ----------
+        X: ArrayLike
+            of features to match on. Will be converted to a ``pd.DataFrame``
+        z: ArrayLike
+            of boolean treatment assignments. Will be converted to a ``pd.Series``
+        ids: ArrayLike
+            of observation IDs. Will be converted to a ``pd.Series``, and coerced to type ``str``
+        """
+        self.X = pd.DataFrame(X).copy().reset_index(drop=True)
+        self.z = pd.Series(np.asarray(z) == treatment_encoded_as, name="is_treatment").copy()
+
+        # Auto-generate IDs, if not supplied
+        if ids is None:
+            self.ids = pd.Series(np.arange(self.nobs)).astype(str)
+        else:
[…]

### 2022-04-25 matching/distance.py

@@ -1,0 +1,293 @@
+from abc import ABCMeta, abstractmethod
+from dataclasses import dataclass, field
+from typing import Iterable, List, NamedTuple, Union
+
+import numpy as np
+from numpy import typing as npt
+from scipy import spatial
+
+from matching.dataset import MatchingDataset
+
+"""In sparse representations of graphs, 0 values are taken to mean "no edge" by convention.
+In order to prevent the sparse representation from dropping edge weights of 0 (perfect matches),
+we clip the edge weights to this small floating point value
+"""
+DEFAULT_DISTANCE_TOL = 1e-16
+
+
+class _WeightedEdgeInfo(NamedTuple):
+    """``NamedTuple`` representing weighted graph edges.
+    ``Iterable[WeightedEdgeInfo]`` can be passed to ``networkx.Graph.add_weighted_edges_from``
+
+    Attributes
+    ----------
+    tid: str
+        treatment ID (first node)
+    cid: str
+        control ID (second node)
+    weight: float
+        weight of the edge
+    """
+
+    tid: str
+    cid: str
+    weight: float
+
+
+@dataclass
+class _GraphUpdateInfo:
+    """Intermediate data structure. Attributes can be used to construct a weighted bipartite graph
+
+    Attributes
+    ----------
+    tids: ArrayLike
+        of treatment ID nodes to be included in the graph, for which `"bipartite" == 1`
+    cids: ArrayLike
+        of control ID nodes to be included in the graph, for which `"bipartite" == 0`
+    edges: Iterable[:class:`WeightedEdgeInfo`]
+        Iterable of <tid, cid, edge weight> tuples
+    """
+
+    tids: npt.ArrayLike
+    cids: npt.ArrayLike
+    edges: Iterable[_WeightedEdgeInfo]
+
+
+# NOTE: type: ignore here because mypy doesn't like abstract dataclasses
+# NOTE: see https://github.com/python/mypy/issues/5374
+@dataclass  # type: ignore
+class _Distance(metaclass=ABCMeta):
+    """Distance metric base class
+
+    Attributes
+    ----------
+    tol: float
+        A small number to replace 0-weight edges (ie edges of perfect matches) with.
+        This is so that the sparse matrix representation does not drop matches.
+        See ``DEFAULT_DISTANCE_TOL`` for details
+    """
+
+    tol: float = field(default=DEFAULT_DISTANCE_TOL)
+
+    @abstractmethod
[…]

### 2022-04-25 matching/graph.py

@@ -1,0 +1,750 @@
+from collections import defaultdict
+from dataclasses import dataclass, field
+import enum
+from typing import Dict, Hashable, Iterable, Iterable, List, Optional, TypeVar, Union
+
+import networkx as nx
+from numpy import typing as npt
+import numpy as np
+from scipy import sparse
+
+from matching import graph_utils as gu
+from matching.dataset import MatchingDataset
+from matching.distance import _Distance
+
+
+class NotBipartiteError(Exception):
+    """Raised if the graph is not bipartite"""
+
+
+@dataclass
+class BipartiteGraphParameters:
+    """Container of various parameters relevant to initializing, plotting, and otherwise working with bipartite graphs
+    using networkx
+
+    Attributes
+    ----------
+    init_weight: float
+        The default weight between nodes in the naive graph, before ``set_edges`` is called
+    treatment_bipartite_attr: Hashable
+        What the "bipartite" attribute for will be set to for each treatment node in the graph
+    control_bipartite_attr: Hashable
+        What the "bipartite" attribute for will be set to for each control node in the graph
+    treatment_color: str
+        Color to plot treatment nodes as
+    control_color: str
+        Color to plot control nodes as
+    match_group_col: str
+        Name of column to add to the graph data, indicating which match the observation belongs to
+    """
+
+    init_weight: float = 1
+    treatment_bipartite_attr: Hashable = 1
+    control_bipartite_attr: Hashable = 0
+    treatment_color: str = "red"
+    control_color: str = "blue"
+    match_group_col: str = "match_group"
+
+
+class MatchingMethod(enum.Enum):
+    """Enum for valid matching methods, with some aliases"""
+
+    FAST = GREEDY = enum.auto()
+    HUNGARIAN = OPTIMAL = KUHN = MUNKRES = enum.auto()
+
+
+"""``TypeVar`` for :class:`MatchingGraph`, since many methods return ``self`` to enable method chaining"""
+_MatchingGraph = TypeVar("_MatchingGraph", bound="MatchingGraph")
+
+
+@dataclass(init=False)
+class MatchingGraph:
+    """:class:`MatchingGraph` offers a simple framework to
+    - calculate a distance metric between two sides of a bipartite graph,
+    - iteratively filter this graph by enforcing other constraints on nodes and edges,
+    - calculate a "matched" subset of this graph via greedy, grouping, or optimal matching strategies, and
+    - compare the balance between input and matched datasets.
+
[…]

### 2022-04-25 matching/graph_utils.py

@@ -1,0 +1,30 @@
+from copy import deepcopy
+from typing import List
+
+import networkx as nx
+
+
+def get_connected_subgraphs(G: nx.Graph, *, copy: bool = False, deep: bool = False, **kwargs) -> List[nx.Graph]:
+    """Split the contents of ``G`` into a list of connected subgraphs.
+
+    Parameters
+    ----------
+    copy: bool
+        Whether or not to make a copy of ``G`` before computing the subgraphs. ``False`` by default
+    deep: Optional[bool]
+        Whether or not to make a _deep_ copy of ``G``. Only used if ``copy == True``. ``False`` by default.
+    **kwargs
+        Passed to ``nx.Graph.copy()`` if ``copy=True`` and ``deep=False``
+
+    Returns
+    -------
+    List[nx.Graph]
+        Subgraphs of ``G``, in a list. Each subgraph is fully connected
+    """
+    if copy:
+        G2 = deepcopy(G) if deep else G.copy(**kwargs)
+    else:
+        G2 = G
+
+    assert isinstance(G2, nx.Graph), "Copy failed"
+    return [G2.subgraph(c) for c in nx.connected_components(G2)]

### 2022-04-25 matching/preprocessing.py

@@ -1,0 +1,149 @@
+from typing import Iterable, Optional
+
+from numpy import typing as npt
+import numpy as np
+import pandas as pd
+from scipy.special import logit
+from sklearn.linear_model import LogisticRegression
+from sklearn.preprocessing import KBinsDiscretizer
+
+from matching._typing import sklearn_types as skt
+
+
+class UnsupportedModel(Exception):
+    """Raised when an unsupported model is provided"""
+
+
+def _get_preds_from_fittable(model: skt._Fittable, X: npt.ArrayLike) -> np.ndarray:
+    """Utility function to get the predictions out of an sklearn model.
+    If the model is a `_Classifier` and predicts probabilities, we take `predict_proba` output, dropping the first column.
+    If the model is a `_Regressor`, we take `predict` output
+
+    Parameters
+    ----------
+    model: :class:`_Fittable`
+        Fitted object which we want to use to predict over `X`
+    X: ArrayLike
+        Of features to get predictions for
+
+    Returns
+    -------
+    np.ndarray
+        - if ``model`` is a :class:`_Classifier`, predicted probabilities, with the first column removed
+        - if ``model`` is a :class:`_Regressor`, predicted values
+    """
+    # TODO: Docstring
+    if isinstance(model, skt._Classifier):
+        return model.predict_proba(X)[:, 1:]
+    elif isinstance(model, skt._Regressor):
+        return model.predict(X)
+    else:
+        raise UnsupportedModel("Model must be of type `skt._Classifier` or `skt._Fittable`")
+
+
+def propensity_score(
+    X: npt.ArrayLike, z: npt.ArrayLike, model: Optional[skt._Fittable] = None, use_logit: bool = False
+) -> np.ndarray:
+    """Compute propensity scores for the data. Propensity scores are predicted treatment assignments (`z`) from `X`.
+    This fitted model is then used to get predicted scores over the WHOLE dataset.
+
+    Parameters
+    ----------
+    X: ArrayLike
+        of features to predict `y` with
+        of binary treatment assignments
+    model: Optional[:class:`_Fittable`]
+        Model to use in prediction. Can be a regressor (for continuous `z`) or a classifier (for discrete `z`)
+        By default, Logistic Regression with no penalty is used
+    use_logit: bool = False
[…]

### 2022-04-25 matching/graph.py
updated docstrings

@@ -38,12 +38,12 @@ class BipartiteGraphParameters:
         Name of column to add to the graph data, indicating which match the observation belongs to
     """
 
-    init_weight: float = 1
-    treatment_bipartite_attr: Hashable = 1
-    control_bipartite_attr: Hashable = 0
-    treatment_color: str = "red"
-    control_color: str = "blue"
-    match_group_col: str = "match_group"
+    init_weight: float = field(default=1)
+    treatment_bipartite_attr: Hashable = field(default=1)
+    control_bipartite_attr: Hashable = field(default=0)
+    treatment_color: str = field(default="red")
+    control_color: str = field(default="blue")
+    match_group_col: str = field(default="match_group")
 
 
 class MatchingMethod(enum.Enum):
@@ -115,7 +115,7 @@ class MatchingGraph:
     def __init__(
         self, X: npt.ArrayLike, z: npt.ArrayLike, ids: Optional[npt.ArrayLike] = None, *, init_naive_graph: bool = False
     ) -> None:
-        """Initialize a ``Similarity`` object with necessary data
+        """Initialize a :class:`MatchingGraph` object with necessary data
 
         Parameters
         ----------

### 2022-04-25 matching/graph.py
updated docstring

@@ -88,13 +88,14 @@ class MatchingGraph:
     Attributes
     ----------
     input_data: :class:`MatchingDataset`
-        The data which is supplied to the constructor (X, z, ids) is parsed into an ``input_data`` object, with attributes
-        ``X``, ``z``, and ``ids`. You can convert this to a ``pd.DataFrame`` via ``<MatchingGraph>.input_data.frame``
+        The data which is supplied to the constructor (``X``, ``z``, ``ids``) is parsed into an ``input_data`` object, 
+        with attributes ``X``, ``z``, and ``ids`. 
+        You can convert this to a ``pd.DataFrame`` via ``<MatchingGraphInst>.input_data.frame``
     graph_data: :class:`MatchingDataset`
-        Just the data for the nodes currently present on the graph. Has all the same properties as ``input_data``. Notes
-        that this is a derived property that is re-calculated each time it is accessed.
+        Just the data for the nodes currently present on the graph. Has all the same properties as ``input_data``. 
+        Note that this is a derived property that is re-calculated each time it is accessed.
     match_data: :class:`MatchingDataset`
-        This is the same as ``graph_data``, except it adds an additional column to ``X``: "match_group", an integer number
+        This is the same as ``graph_data``, except it adds an additional column to ``X``: "match_group", an integer
         indicating which matched subgraph the match belongs to
     graph: nx.Graph
         Bipartite graph, where nodes represent observation IDs and edge weights correspond to the cost of matching

### 2022-04-25 matching/dataset.py
fixed docstrings to follow numpydoc

@@ -8,12 +8,12 @@ import pandas as pd
 
 class _MatchingDatasetNDArrays(NamedTuple):
     """Container for the NDArray equivalent of each element of a :class:`MatchingDataset`.
-    Note that non-numeric columns of ``X`` are one-hot encoded.
+    Note that non-numeric columns of `X` are one-hot encoded.
 
     Attributes
     ----------
     X: ndarray
-        of features, with non-numeric columns one-hot encoded with ``drop_first=True``
+        of features, with non-numeric columns one-hot encoded with `drop_first=True`
     z: ndarray
         of boolean treatment assignments
     ids: ndarray
@@ -40,7 +40,7 @@ class MatchingDataset:
     z: Series
         of boolean treatment assignments
     ids: Series
-        of observation IDs. Note that these are coerced to ``str``
+        of observation IDs. Note that these are coerced to `str`
     """
 
     X: pd.DataFrame
@@ -58,11 +58,11 @@ class MatchingDataset:
         Parameters
         ----------
         X: ArrayLike
-            of features to match on. Will be converted to a ``pd.DataFrame``
+            of features to match on. Will be converted to a `pd.DataFrame`
         z: ArrayLike
-            of boolean treatment assignments. Will be converted to a ``pd.Series``
+            of boolean treatment assignments. Will be converted to a `pd.Series`
         ids: ArrayLike
-            of observation IDs. Will be converted to a ``pd.Series``, and coerced to type ``str``
+            of observation IDs. Will be converted to a `pd.Series`, and coerced to type `str`
         """
         self.X = pd.DataFrame(X).copy().reset_index(drop=True)
         self.z = pd.Series(np.asarray(z) == treatment_encoded_as, name="is_treatment").copy()
@@ -118,7 +118,7 @@ class MatchingDataset:
 
     @property
     def arrs(self) -> _MatchingDatasetNDArrays:
-        """Numpy array versions of ``X``, ``z``, and ``ids``. Non-numeric columns of ``X`` are one-hot encoded"""
+        """Numpy array versions of `X`, `z`, and `ids`. Non-numeric columns of `X` are one-hot encoded"""
         Xarr = pd.get_dummies(self.X, drop_first=True).values
         zarr = self.z.values.astype(bool)
         idsarr = self.ids.values.astype(str)
@@ -145,14 +145,14 @@ class MatchingDataset:
     def groupby(
[…]

### 2022-04-25 matching/distance.py

@@ -16,8 +16,8 @@ DEFAULT_DISTANCE_TOL = 1e-16
 
 
 class _WeightedEdgeInfo(NamedTuple):
-    """``NamedTuple`` representing weighted graph edges.
-    ``Iterable[WeightedEdgeInfo]`` can be passed to ``networkx.Graph.add_weighted_edges_from``
+    """`NamedTuple` representing weighted graph edges.
+    `Iterable[WeightedEdgeInfo]` can be passed to `networkx.Graph.add_weighted_edges_from`
 
     Attributes
     ----------
@@ -64,7 +64,7 @@ class _Distance(metaclass=ABCMeta):
     tol: float
         A small number to replace 0-weight edges (ie edges of perfect matches) with.
         This is so that the sparse matrix representation does not drop matches.
-        See ``DEFAULT_DISTANCE_TOL`` for details
+        See `DEFAULT_DISTANCE_TOL` for details
     """
 
     tol: float = field(default=DEFAULT_DISTANCE_TOL)
@@ -116,7 +116,7 @@ class Exact(_Distance):
 
 @dataclass
 class _Tuneable:
-    """A baseclass for distance metrics that can use ``min_distance`` and ``max_distance`` parameters to tune edge creation.
+    """A baseclass for distance metrics that can use `min_distance` and `max_distance` parameters to tune edge creation.
     Exists for MRO reasons with dataclasses, so that constuctor parameters take the desired order
 
     Attributes
@@ -133,7 +133,7 @@ class _Tuneable:
 
 @dataclass
 class _Norm:
-    """A baseclass for distance metrics norm-based distance metrics. ``p`` denotes which Lp Norm to use.
+    """A baseclass for distance metrics norm-based distance metrics. `p` denotes which Lp Norm to use.
     Exists for MRO reasons with dataclasses, so that constuctor parameters take the desired order
 
     Attributes
@@ -159,12 +159,12 @@ class Norm(_Distance, _Tuneable, _Norm):
         Maximum allowable edge weight. Can be used as a tuning parameter for identifying similar groups
     tol: float
         A small float. Edge weights of 0 (perfect matches) are replaced with this value, so that the sparse matrix
-        representation does not drop perfect matches (normally, 0 = no edge). Does not affect ``max_distance``.
-        See ``DEFAULT_DISTANCE_TOL`` for details
+        representation does not drop perfect matches (normally, 0 = no edge). Does not affect `max_distance`.
+        See `DEFAULT_DISTANCE_TOL` for details
     """
 
[…]

### 2022-04-25 matching/graph.py

@@ -25,7 +25,7 @@ class BipartiteGraphParameters:
     Attributes
     ----------
     init_weight: float
-        The default weight between nodes in the naive graph, before ``set_edges`` is called
+        The default weight between nodes in the naive graph, before `set_edges` is called
     treatment_bipartite_attr: Hashable
         What the "bipartite" attribute for will be set to for each treatment node in the graph
     control_bipartite_attr: Hashable
@@ -53,7 +53,7 @@ class MatchingMethod(enum.Enum):
     HUNGARIAN = OPTIMAL = KUHN = MUNKRES = enum.auto()
 
 
-"""``TypeVar`` for :class:`MatchingGraph`, since many methods return ``self`` to enable method chaining"""
+"""`TypeVar` for :class:`MatchingGraph`, since many methods return `self` to enable method chaining"""
 _MatchingGraph = TypeVar("_MatchingGraph", bound="MatchingGraph")
 
 
@@ -65,37 +65,37 @@ class MatchingGraph:
     - calculate a "matched" subset of this graph via greedy, grouping, or optimal matching strategies, and
     - compare the balance between input and matched datasets.
 
-    At construction, the user supplies three ``ArrayLike``s:
-    - ``X``, an array of features to match on;
-    - ``z``, an array of binary treatment assignments; and
-    - ``ids``, an optional array of observation IDs. If not supplied, default IDs are enumerated from 0 upwards.
+    At construction, the user supplies three `ArrayLike`s:
+    - `X`, an array of features to match on;
+    - `z`, an array of binary treatment assignments; and
+    - `ids`, an optional array of observation IDs. If not supplied, default IDs are enumerated from 0 upwards.
 
-    These inputs are parsed into the attribute ``input_data``. The user can then proceed to ``set_edges`` by supplying
-    a :class:`_Distance` measure, optionally specifying specific columns of ``X`` to ``include`` or ``exclude`` from the
+    These inputs are parsed into the attribute `input_data`. The user can then proceed to `set_edges` by supplying
+    a :class:`_Distance` measure, optionally specifying specific columns of `X` to `include` or `exclude` from the
     distance calculation.
 
     Once the edges are calculated, the user can then use one of the matching strategies provided to "match" their
-    dataset. Alternatively, the user can proceed to iteratively filter the graph further, via the ``filter_edges`` and
-    ``filter_nodes``.
-    - ``filter_edges`` allows the user to enforce additional distance metrics on the data, while preserving the edge
[…]

### 2022-04-25 matching/graph_utils.py

@@ -5,21 +5,21 @@ import networkx as nx
 
 
 def get_connected_subgraphs(G: nx.Graph, *, copy: bool = False, deep: bool = False, **kwargs) -> List[nx.Graph]:
-    """Split the contents of ``G`` into a list of connected subgraphs.
+    """Split the contents of `G` into a list of connected subgraphs.
 
     Parameters
     ----------
     copy: bool
-        Whether or not to make a copy of ``G`` before computing the subgraphs. ``False`` by default
+        Whether or not to make a copy of `G` before computing the subgraphs. `False` by default
     deep: Optional[bool]
-        Whether or not to make a _deep_ copy of ``G``. Only used if ``copy == True``. ``False`` by default.
+        Whether or not to make a _deep_ copy of `G`. Only used if `copy == True`. `False` by default.
     **kwargs
-        Passed to ``nx.Graph.copy()`` if ``copy=True`` and ``deep=False``
+        Passed to `nx.Graph.copy()` if `copy=True` and `deep=False`
 
     Returns
     -------
     List[nx.Graph]
-        Subgraphs of ``G``, in a list. Each subgraph is fully connected
+        Subgraphs of `G`, in a list. Each subgraph is fully connected
     """
     if copy:
         G2 = deepcopy(G) if deep else G.copy(**kwargs)

### 2022-04-25 matching/preprocessing.py

@@ -29,8 +29,8 @@ def _get_preds_from_fittable(model: skt._Fittable, X: npt.ArrayLike) -> np.ndarr
     Returns
     -------
     np.ndarray
-        - if ``model`` is a :class:`_Classifier`, predicted probabilities, with the first column removed
-        - if ``model`` is a :class:`_Regressor`, predicted values
+        - if `model` is a :class:`_Classifier`, predicted probabilities, with the first column removed
+        - if `model` is a :class:`_Regressor`, predicted values
     """
     # TODO: Docstring
     if isinstance(model, skt._Classifier):
@@ -124,12 +124,12 @@ def autocoarsen(X: npt.ArrayLike, n_bins: npt.ArrayLike = 5) -> np.ndarray:
     X: ArrayLike
         input array to coarsen
     n_bins: ArrayLike
-        either a single ``int``, or an array of ``int``s, specifying number of bins for each column
+        either a single `int`, or an array of `int`s, specifying number of bins for each column
 
     Returns
     -------
     ndarray
-        of coarsened ``X``
+        of coarsened `X`
     """
     Xarr = pd.get_dummies(X).values
     if not not isinstance(n_bins, Iterable) or np.asarray(n_bins).shape[0] == Xarr.shape[1]:

### 2022-04-26 matching/dataset.py
updated docstrings, with space before :

@@ -12,11 +12,11 @@ class _MatchingDatasetNDArrays(NamedTuple):
 
     Attributes
     ----------
-    X: ndarray
+    X : ndarray
         of features, with non-numeric columns one-hot encoded with `drop_first=True`
-    z: ndarray
+    z : ndarray
         of boolean treatment assignments
-    ids: ndarray
+    ids : ndarray
         of string IDs
     """
 
@@ -35,11 +35,11 @@ class MatchingDataset:
 
     Attributes
     ----------
-    X: DataFrame
+    X : DataFrame
         of features to match on
-    z: Series
+    z : Series
         of boolean treatment assignments
-    ids: Series
+    ids : Series
         of observation IDs. Note that these are coerced to `str`
     """
 
@@ -57,12 +57,14 @@ class MatchingDataset:
         """
         Parameters
         ----------
-        X: ArrayLike
+        X : ArrayLike
             of features to match on. Will be converted to a `pd.DataFrame`
-        z: ArrayLike
+        z : ArrayLike
             of boolean treatment assignments. Will be converted to a `pd.Series`
-        ids: ArrayLike
+        ids : ArrayLike, default=None
             of observation IDs. Will be converted to a `pd.Series`, and coerced to type `str`
+        treatment_encoded_as : Hashable, default=True
+            Value in `z` corresponding to treatment group membership.
         """
         self.X = pd.DataFrame(X).copy().reset_index(drop=True)
         self.z = pd.Series(np.asarray(z) == treatment_encoded_as, name="is_treatment").copy()
@@ -82,7 +84,7 @@ class MatchingDataset:
 
         Parameters
         ----------
-        __o: object
+        __o : object
             for comparison
 
         Returns
@@ -129,7 +131,7 @@ class MatchingDataset:
 
         Parameters
         ----------
-        indexer: ArrayLike
+        indexer : ArrayLike
             of boolean values, indicating which rows to keep
 
         Returns
@@ -149,7 +151,7 @@ class MatchingDataset:
 
         Parameters
         ----------
-        by: Union[Hashable, Iterable[Hashable], npt.ArrayLike]
+        by : Union[Hashable, Iterable[Hashable], npt.ArrayLike]
             A label, sequence of labels, or indexing array to supply to `pd.DataFrame.groupby`
         *args, **kwargs
             Passed to `pd.DataFrame.groupby`
@@ -172,7 +174,7 @@ class MatchingDataset:
 
         Parameters
         ----------
[…]

### 2022-04-26 matching/distance.py

@@ -21,11 +21,11 @@ class _WeightedEdgeInfo(NamedTuple):
 
     Attributes
     ----------
-    tid: str
+    tid : str
         treatment ID (first node)
-    cid: str
+    cid : str
         control ID (second node)
-    weight: float
+    weight : float
         weight of the edge
     """
 
@@ -40,11 +40,11 @@ class _GraphUpdateInfo:
 
     Attributes
     ----------
-    tids: ArrayLike
+    tids : ArrayLike
         of treatment ID nodes to be included in the graph, for which `"bipartite" == 1`
-    cids: ArrayLike
+    cids : ArrayLike
         of control ID nodes to be included in the graph, for which `"bipartite" == 0`
-    edges: Iterable[:class:`WeightedEdgeInfo`]
+    edges : Iterable[:class:`WeightedEdgeInfo`]
         Iterable of <tid, cid, edge weight> tuples
     """
 
@@ -61,7 +61,7 @@ class _Distance(metaclass=ABCMeta):
 
     Attributes
     ----------
-    tol: float
+    tol : float, default=`DEFAULT_DISTANCE_TOL`
         A small number to replace 0-weight edges (ie edges of perfect matches) with.
         This is so that the sparse matrix representation does not drop matches.
         See `DEFAULT_DISTANCE_TOL` for details
@@ -93,7 +93,7 @@ class Exact(_Distance):
 
     Attributes
     ----------
-    tol: float
+    tol : float, default=`DEFAULT_DISTANCE_TOL`
         A small number to replace 0-weight edges (ie edges of perfect matches) with.
         This is so that the sparse matrix representation does not drop matches.
         See DEFAULT_DISTANCE_TOL for details
@@ -121,9 +121,9 @@ class _Tuneable:
 
     Attributes
     ----------
-    min_distance: float
+    min_distance : float, default=0
         Minimum allowable edge weight. Can be used as a tuning parameter for identifying similar groups
-    max_distance: float
+    max_distance : float, default=inf
         Maximum allowable edge weight. Can be used as a tuning parameter for identifying similar groups
     """
 
@@ -138,7 +138,7 @@ class _Norm:
 
     Attributes
     ----------
-    p: Union[int, float]
+    p : Union[int, float]
         Which LpNorm to use. Either an integer value, or infinity
     """
 
@@ -151,13 +151,13 @@ class Norm(_Distance, _Tuneable, _Norm):
 
     Attributes
     ----------
[…]

### 2022-04-26 matching/graph.py

@@ -24,17 +24,17 @@ class BipartiteGraphParameters:
 
     Attributes
     ----------
-    init_weight: float
+    init_weight : float
         The default weight between nodes in the naive graph, before `set_edges` is called
-    treatment_bipartite_attr: Hashable
+    treatment_bipartite_attr : Hashable
         What the "bipartite" attribute for will be set to for each treatment node in the graph
-    control_bipartite_attr: Hashable
+    control_bipartite_attr : Hashable
         What the "bipartite" attribute for will be set to for each control node in the graph
-    treatment_color: str
+    treatment_color : str
         Color to plot treatment nodes as
-    control_color: str
+    control_color : str
         Color to plot control nodes as
-    match_group_col: str
+    match_group_col : str
         Name of column to add to the graph data, indicating which match the observation belongs to
     """
 
@@ -87,21 +87,21 @@ class MatchingGraph:
 
     Attributes
     ----------
-    input_data: :class:`MatchingDataset`
-        The data which is supplied to the constructor (`X`, `z`, `ids`) is parsed into an `input_data` object, 
-        with attributes `X`, `z`, and `ids`. 
+    input_data : :class:`MatchingDataset`
+        The data which is supplied to the constructor (`X`, `z`, `ids`) is parsed into an `input_data` object,
+        with attributes `X`, `z`, and `ids`.
         You can convert this to a `pd.DataFrame` via `<MatchingGraphInst>.input_data.frame`
-    graph_data: :class:`MatchingDataset`
-        Just the data for the nodes currently present on the graph. Has all the same properties as `input_data`. 
+    graph_data : :class:`MatchingDataset`
+        Just the data for the nodes currently present on the graph. Has all the same properties as `input_data`.
         Note that this is a derived property that is re-calculated each time it is accessed.
-    match_data: :class:`MatchingDataset`
+    match_data : :class:`MatchingDataset`
         This is the same as `graph_data`, except it adds an additional column to `X`: "match_group", an integer
         indicating which matched subgraph the match belongs to
-    graph: nx.Graph
+    graph : nx.Graph
         Bipartite graph, where nodes represent observation IDs and edge weights correspond to the cost of matching
-    distance: :class:`_Distance`
+    distance : :class:`_Distance`
         The distance measure used to calculate the edge weights on the graph
-    _params: :class:`BipartiteGraphParameters`
+    _params : :class:`BipartiteGraphParameters`
         Private attribute that provides a single container for default arguments to pass to graph functions
     """
[…]

### 2022-04-26 matching/graph_utils.py

@@ -9,10 +9,10 @@ def get_connected_subgraphs(G: nx.Graph, *, copy: bool = False, deep: bool = Fal
 
     Parameters
     ----------
-    copy: bool
-        Whether or not to make a copy of `G` before computing the subgraphs. `False` by default
-    deep: Optional[bool]
-        Whether or not to make a _deep_ copy of `G`. Only used if `copy == True`. `False` by default.
+    copy : bool, default=False
+        Whether or not to make a copy of `G` before computing the subgraphs
+    deep: Optional[bool], default=False
+        Whether or not to make a _deep_ copy of `G`. Only used if `copy == True`
     **kwargs
         Passed to `nx.Graph.copy()` if `copy=True` and `deep=False`
 

### 2022-04-26 matching/preprocessing.py

@@ -21,9 +21,9 @@ def _get_preds_from_fittable(model: skt._Fittable, X: npt.ArrayLike) -> np.ndarr
 
     Parameters
     ----------
-    model: :class:`_Fittable`
+    model : :class:`_Fittable`
         Fitted object which we want to use to predict over `X`
-    X: ArrayLike
+    X : ArrayLike
         Of features to get predictions for
 
     Returns
@@ -49,13 +49,14 @@ def propensity_score(
 
     Parameters
     ----------
-    X: ArrayLike
+    X : ArrayLike
         of features to predict `y` with
+    z : ArrayLike
         of binary treatment assignments
-    model: Optional[:class:`_Fittable`]
+    model : Optional[:class:`_Fittable`]
         Model to use in prediction. Can be a regressor (for continuous `z`) or a classifier (for discrete `z`)
         By default, Logistic Regression with no penalty is used
-    use_logit: bool = False
+    use_logit : bool, default=False
         Whether or not to take the logit of the predicted values. Only pertinent when `model` is a classifier (and predicts probabilities)
 
     Returns
@@ -85,16 +86,16 @@ def prognostic_score(
 
     Parameters
     ----------
-    X: ArrayLike
+    X : ArrayLike
         of features to predict `y` with
-    y: ArrayLike
+    y : ArrayLike
         of outcomes
-    z: ArrayLike
+    z : ArrayLike
         of binary treatment assignments
-    model: Optional[_Fittable]
+    model : Optional[_Fittable]
         Model to use in prediction. Can be a regressor (for continuous response) or a classifier (for discrete response)
         By default, Logistic Regression with no penalty is used
-    use_logit: bool = False
+    use_logit : bool = False
         Whether or not to take the logit of the predicted values. Only pertinent when `model` is a classifier (and predicts probabilities)
 
     Returns
@@ -121,9 +122,9 @@ def autocoarsen(X: npt.ArrayLike, n_bins: npt.ArrayLike = 5) -> np.ndarray:
 
     Parameters
     ----------
-    X: ArrayLike
+    X : ArrayLike
         input array to coarsen
-    n_bins: ArrayLike
+    n_bins : ArrayLike
         either a single `int`, or an array of `int`s, specifying number of bins for each column
 
     Returns

### 2022-04-26 matching/balance.py
Updated summary() method to mirror MatchIt

@@ -1,0 +1,73 @@
+import numpy as np
+from numpy import typing as npt
+
+
+def standardized_mean_difference(a: npt.ArrayLike, b: npt.ArrayLike):
+    """Computes the standardized mean difference between a and b.
+
+    .. math:: \frac{\overline{a} - \overline{b}}{\sqrt{\text{Var}(a) + \text{Var}(b)}}
+
+    Parameters
+    ----------
+    a, b : npt.ArrayLike
+        to compute the standardized mean difference between
+
+    Returns
+    -------
+    float
+        standardized mean difference
+    """
+    meana, stda = np.mean(a), np.std(a)
+    meanb, stdb = np.mean(b), np.std(b)
+
+    return (meana - meanb) / np.sqrt(stda**2 + stdb**2)
+
+
+def ecdf(x: npt.ArrayLike):
+    """Fits an eCDF to `x` and returns a callable to evaluate it at any point(s)
+
+    Parameters
+    ----------
+    x : ArrayLike
+        to fit the eCDF to
+
+    Returns
+    -------
+    inner : Callable
+        evaluates the eCDF at point `v`
+    """
+    x_sorted = np.sort(np.asarray(x))
+
+    def fitted(v: npt.ArrayLike) -> npt.ArrayLike:
+        """Evaluates the eCDF of `x` at point `v`
+
+        Parameters
+        ----------
+        v : ArrayLike
+            point(s) to evaluate the eCDF of fitted `x` at
+
+        Returns
+        -------
+        ArrayLike
+            of eCDF evaluations in [0, 1]
+        """
+        return np.searchsorted(x_sorted, v, side="right") / x_sorted.size
+
+    return fitted
+
+
+def variance_ratio(a: npt.ArrayLike, b: npt.ArrayLike) -> float:
+    """Compute the variance ratio between `a` and `b`, defined as
+    .. math:: \frac{\text{Var}(a)}{\text{Var}(b)}
+
+    Parameters
+    ----------
+    a, b : ArrayLike
+        To compute the variance ratio between
+
+    Returns
+    -------
+    float
+        variance ratio
+    """
+    return np.var(a) / np.var(b)

### 2022-04-26 matching/dataset.py

@@ -5,6 +5,8 @@ from numpy import typing as npt
 import numpy as np
 import pandas as pd
 
+from matching import balance
+
 
 class _MatchingDatasetNDArrays(NamedTuple):
     """Container for the NDArray equivalent of each element of a :class:`MatchingDataset`.
@@ -188,11 +190,57 @@ class MatchingDataset:
         return type(self)(X, z, ids)
 
     def summary(self) -> pd.DataFrame:
-        """Print a summary of the matching :class:`_MatchingDataset`"""
-        return pd.concat(
-            [
-                self.frame.groupby(self.z.name).describe(include=np.number).T,
-                self.frame.groupby(self.z.name).describe(include=np.object_).T,
-            ],
-            axis=0,
-        )
+        """Print a summary of the matching :class:`_MatchingDataset`.
+
+        Metrics include:
+        - min, 25%, 50%, 75%, max;
+        - mean;
+        - standard deviation;
+        - standardized mean difference;
+        - variance ratio;
+        - eCDF of mean.
+
+        Returns
+        -------
+        pd.DataFrame
+            Note that this is multi-indexed by feature name, then metric (i.e. "min").
+            Columns represent treatment assignment
+        """
+
+        # NOTE: OLD BEHAVIOR
+        # numeric_summary = self.frame.groupby(self.z.name).describe(include=np.number).T
+        # discrete_summary = self.frame.groupby(self.z.name).describe(include=np.object_).T
+        # return pd.concat([numeric_summary, discrete_summary], axis=0)
+
+        dummified_df = pd.get_dummies(self.X, drop_first=True)
+        describe_df = dummified_df.groupby(self.z).describe().T
+        subdfs = []
+        for colname, subdf in describe_df.groupby(describe_df.index.get_level_values(0)):
+            # Get data from dummified data
+            tdata, cdata = dummified_df.loc[self.z, colname], dummified_df.loc[~self.z, colname]
+
+            # Get mean within each group
+            tmean, cmean = np.mean(tdata.values), np.mean(cdata.values)
+
+            # Fit eCDF within each group
+            tecdf, cecdf = balance.ecdf(tdata), balance.ecdf(cdata)
+
+            # Compute standardized mean differences, add to table
+            subdf.loc[(colname, "standardized_difference"), :] = [
+                balance.standardized_mean_difference(tdata, cdata),
+                balance.standardized_mean_difference(cdata, tdata),
+            ]
+
[…]

### 2022-04-26 paper/main.tex
outlined most of lit review

@@
 
-\maketitle
+\maketitle{}
 
 \section*{Abstract}
-\tableofcontents
-\newpage
+\tableofcontents{}
+\newpage{}
 
 \section{Introduction}
+\label{sec:Introduction}
+
+In the context of observational data, a bipartite matching problem is minimally defined by:
+\begin{itemize}
+  \item \mat{X}, an $n \times p$ matrix of feature values;
+  \item \vec{z}, an $n$-vector of binary treatment assignments;
+  \item $d : \reals^p \times \reals^p \mapsto \reals^+$, a function which computes some ``distance'' or ``cost of matching'' between any two row vectors \vec{x_i} and \vec{x_j} of \mat{X}.
+\end{itemize}
+% The output
+
+\set{S}, \vec{b}, \mat{A}, \vec{\upbeta}
+test \cite{iacus_multivariate_2011}
 
@@
 \subsection{Measuring Similarity}
+Iacus, King, Porro 2011 - set the scene
 
@@
 
-Methods for calculating propensity scores: Garrido et al, 2014
+Methods for calculating propensity scores: \cite{garrido_methods_2014}
 Note: Blocking-based balance metrics
@@
 \begin{itemize}
-  \item (Optimal) Caliper width: Austin 2011
-  \item Prognostic score: Hansen
-  \item Miettinen score is the root of the above
-  \item Joint use of Prognostic, Propensity, + Mahalanobis, Leacy and Stuart 2014
-  \item Imai and Van Dyk (make sure to include)
+  \item (Optimal) Caliper width: \cite{austin_optimal_2011}
+  \item Prognostic score: \cite{hansen_prognostic_2008}
+  \item Miettinen score is the root of the above \cite{miettinen_stratification_1976}
+  \item Joint use of Prognostic, Propensity, + Mahalanobis, \cite{leacy_joint_2014}
+  \item \cite{imai_causal_2004}
 \end{itemize}
@@
 
+Garrido et al, 2014: don't use the outcome in the matching
+Garrido et al, 2014: balance in mean does not imply balance in scores
+
 \subsection{Matching Algorithms}
+
+\parencite{rosenbaum_optimal_1989}
+\parencite{ho_matchit_2011} (Greedy)
+\parencite{khan_efficient_2016} (hungarian)
+\parencite{munkres_algorithms_1957} (hungarian)
+Hungarian -- Munkres, Khan et al 2016
+
+\section{Python Package}
 
@@
 \begin{itemize}
-  \item Austin 2011 (optimal caliper widths)
-  \item Stuart et al: 2018
+  \item \cite{austin_optimal_2011}
+  \item \cite{stuart_prognostic_2013}
 \end{itemize}
@@
 
+% NOTE: SEE LINK BELOW
+% https://tex.stackexchange.com/questions/8458/making-the-bibliography-appear-in-the-table-of-contents
+% Special handling of bibliography awaits!!
+\cleardoublepage
+\phantomsection
+\addcontentsline{toc}{section}{Bibliography}
+\printbibliography
+
 \end{document}

### 2022-04-27 matching/dataset.py
small edit

@@ -214,7 +214,7 @@ class MatchingDataset:
 
         dummified_df = pd.get_dummies(self.X, drop_first=True)
         describe_df = dummified_df.groupby(self.z).describe().T
-        subdfs = []
+        subdescribe_dfs = []
         for colname, subdf in describe_df.groupby(describe_df.index.get_level_values(0)):
             # Get data from dummified data
             tdata, cdata = dummified_df.loc[self.z, colname], dummified_df.loc[~self.z, colname]
@@ -241,6 +241,7 @@ class MatchingDataset:
             subdf.loc[(colname, "ecdf_mean"), :] = [tecdf(tmean), cecdf(cmean)]
 
             # Append modified subdf
-            subdfs.append(subdf)
+            subdescribe_dfs.append(subdf)
 
-        return pd.concat(subdfs, axis=0)
+        # Concatenate dfs on top of each other
+        return pd.concat(subdescribe_dfs, axis=0)

### 2022-04-27 docs/source/conf.py
sphinx files

@@ -1,0 +1,82 @@
+import sys
+import pathlib
+
+CURRENT_FILE = pathlib.Path(__file__).resolve()
+REPOSITORY_ROOT = CURRENT_FILE.parent.parent.parent
+sys.path.insert(0, str(REPOSITORY_ROOT))
+
+# Configuration file for the Sphinx documentation builder.
+#
+# This file only contains a selection of the most common options. For a full
+# list see the documentation:
+# https://www.sphinx-doc.org/en/master/usage/configuration.html
+
+# -- Path setup --------------------------------------------------------------
+
+# If extensions (or modules to document with autodoc) are in another directory,
+# add these directories to sys.path here. If the directory is relative to the
+# documentation root, use os.path.abspath to make it absolute, like shown here.
+#
+# import os
+# import sys
+# sys.path.insert(0, '/Users/jackpotrykus/GitHub/propensity-score-matching-thesis/matching')
+
+
+# -- Project information -----------------------------------------------------
+
+project = "matching"
+copyright = "2022, Author"
+author = "Author"
+
+
+# -- General configuration ---------------------------------------------------
+
+# Add any Sphinx extension module names here, as strings. They can be
+# extensions coming with Sphinx (named 'sphinx.ext.*') or your custom
+# ones.
+extensions = [
+    "sphinx.ext.autodoc",
+    "sphinx.ext.napoleon",
+    "sphinx.ext.viewcode",
+    "sphinx.ext.todo",
+]
+napoleon_numpy_docstring = True
+napoleon_include_init_with_doc = True
+napoleon_include_private_with_doc = False
+
+
+# Add any paths that contain templates here, relative to this directory.
+templates_path = ["_templates"]
+
+# The language for content autogenerated by Sphinx. Refer to documentation
+# for a list of supported languages.
+#
+# This is also used if you do content translation via gettext catalogs.
+# Usually you set "language" from the command line for these cases.
+language = "en"
+
+# List of patterns, relative to source directory, that match files and
+# directories to ignore when looking for source files.
+# This pattern also affects html_static_path and html_extra_path.
+exclude_patterns = ["_build", "Thumbs.db", ".DS_Store"]
+
+
+# -- Options for HTML output -------------------------------------------------
+
+# The theme to use for HTML and HTML Help pages.  See the documentation for
+# a list of builtin themes.
+#
+html_theme = "alabaster"
+
+# Add any paths that contain custom static files (such as style sheets) here,
[…]

### 2022-04-29 docs/source/conf.py
updated helper docs

@@ -25,8 +25,8 @@ sys.path.insert(0, str(REPOSITORY_ROOT))
 # -- Project information -----------------------------------------------------
 
 project = "matching"
-copyright = "2022, Author"
-author = "Author"
+copyright = "2022, Jack Potrykus"
+author = "Jack Potrykus"
 
 
 # -- General configuration ---------------------------------------------------

### 2022-04-29 paper/main.tex
updated paper

--- 2022-04-25 paper/main.tex
@@
 
-\title{Assessing the Performance of Matching Methods in Observational Studies}
+\title{Development of a Python Package for Matching Observational Data}
 \author{Jack Potrykus}
@@
 
-\maketitle
+\maketitle{}
 
 \section*{Abstract}
-\tableofcontents
-\newpage
+\label{sec:abstract}
+
+Observational studies are widely used throughout econometrics, psychology, and medical research.
+Matching is a field in causal statistics concerned with algorithms to minimize the effect of selection bias in the observational data on analyses of treatment effects.
+In particular, in binary treatment/control studies, these algorithms work by matching each treatment observation to one or more ``nearby'' control observations.
+This paper breaks the matching procedure down into two key components: how distance is measured, and how matches are assigned.
+In doing so, it explores several distance metrics, in particular the propensity score \parencite{rosenbaum_central_1983}, the prognostic score \parencite{hansen_prognostic_2008}, and exact matching \parencite{iacus_causal_2012} and its machine learning extensions \parencite{liu_interpretable_2019, wang_flame_2021}; it then explores several matching algorithms that can be used to produce the matched subset once the distance is calculated, in particular the Hungarian algorithm \parencite{munkres_algorithms_1957} and greedy algorithms \parencite{ho_matchit_2011}.
+I then showcase the functionality of \texttt{matching}, an open-source Python package I have developed for matching observational data which takes a graph-centric approach, something which no other package offers.
+I finally explore some practical considerations of parameter tuning when matching via experiments on simulated data.
+
+\newpage{}
+\tableofcontents{}
+\newpage{}
 
 \section{Introduction}
+\label{sec:intro}
+
+Observational studies are of ever-increasing importance as the numerous amount of data collected each year continues to grow.
+However, of observational studies are at a major disadvantage compared to blocking-based study design or randomized control trials (RCTs), because the researchers do not have control over the distribution of confounding variables, known or unknown, in the data itself.
+As such, researchers must take preliminary steps to alter the data in some way in order to minimize the effects of selection bias, which results from heterogeneous distributions present in the observational data.
+
+Within the field of causal inference, the term \emph{matching} is often associated with observational studies with a binary treatment indicator.
+Using a similar framework to that of \textcite{iacus_multivariate_2011}, consider $\mat{X}_T \in \reals^{n \times k}$ and $\mat{X}_C \in \reals^{m \times k}$, matrices of $k$ features for $n$ treatment-group observations and $m$ control-group observations, $n \leq m$.
+We seek to produce a matching $\set{M}$ such that each observation in $\mat{X}_T$ is paired with one or more nearby observations from $\mat{X}_C$, according to a distance metric and a matching algorithm.
+
+The choice of distance measure and choice of matching algorithm are orthogonal, and this paper (and the Python package) will discuss them as such.
+Indeed, measuring distance, as well as \emph{balance diagnostics} between treatment and control distributions, are problems most associated with the field of statistics; in contrast, producing the optimally matched subset is a \emph{minimum cost flow} problem in the field of computer science \parencite{rosenbaum_optimal_1989}.
+
+I then introduce a Python package I have developed, \texttt{matching}, which provides an idiomatic and flexible tool for matching observational data.
+It achieves this by allowing distance measures and matching algorithms to be independently and iteratively applied, so that even incredibly bespoke matching procedures may be implemented succinctly in code.
+It is also the first Python library for matching observational data which is ``graph-centric'': the key data structure used is a bipartite graph, whose nodes represent observation IDs, and edge weights the distance measure between them.
+This enables further exploration into new matching procedures and balance diagnostics using graph metrics and algorithms from computer science.
+
+% Matching algorithms in practice come in two forms
 
 \section{Literature Review}
+\label{sec:litrev}
 
-\subsection{Measuring Similarity}
+\subsection{Measuring Distance}
 
-\subsubsection{Propensity Scoring}
+In the context of this paper, distance measures are functions $d: \reals^k \times \reals^k \mapsto \reals^+$ which produce a non-negative distance, or ``cost of matching'', between any vectors $\vec{x}_T, \vec{x}_C$ from the rows of $\mat{X}_T \in \reals^{n \times k}$ and $\mat{X}_C \in \reals^{m \times k}$, to judge the match quality.
+Distance measures are often combined with some preprocessing function $f(\mat{X})$, usually in the form of a dimension reduction or discretization of the data.
+Again using a framework akin to that of \textcite{iacus_multivariate_2011}, one may succinctly describe a distance measurement as
+\begin{equation}
+  \mathcal{D}_d (f(\mat{X}_T), f(\mat{X}_C)), \label{eqn:abstractdistance}
+\end{equation}
+where $\mathcal{D}_d : \reals^{n \times k} \times \reals^{m \times k} \mapsto \reals^{n \times m}$ is a ``vectorized'' version of $d$, producing a matrix $\mat{D}$ whose $(i, j)$th element is the distance between observation $i$ from $\mat{X}_T$ and observation $j$ from $\mat{X}_C$.
 
-Rosenbaum and Rubin, 1983
-\begin{itemize}
-  \item The \emph{propensity score} is most often accredited to \ldots
-  \item They offer a more general definition of a score: anything affording conditional independence
-  \item Score seeks to replicated randomized trial
-  \item Review of properties (particularly: when can we make unbiased estimates of ATE)
-  \item Proposed use cases
-\end{itemize}
+In the simplest case, $d$ is some norm (perhaps Euclidean) calculated on the difference between feature vectors, and $f$ is the identity.
+With this simple framework in mind, we are now ready to explore extensions of the simple case, and examine their proposed benefits.
 
-Methods for calculating propensity scores: Garrido et al, 2014
-Note: Blocking-based balance metrics
+\subsubsection{Balancing Scores}
 
-\paragraph{Extensions}
-\begin{itemize}
-  \item (Optimal) Caliper width: Austin 2011
-  \item Prognostic score: Hansen
-  \item Miettinen score is the root of the above
-  \item Joint use of Prognostic, Propensity, + Mahalanobis, Leacy and Stuart 2014
-  \item Imai and Van Dyk (make sure to include)
-\end{itemize}
+Balancing scores
+\footnote{Balancing scores come in a variety of forms: as \textcite{rosenbaum_central_1983} note, $b(X) = X$ is the simplest balancing score.
+However, this paper uses the term ``balancing score'' specifically with regards to 1-dimensional reductions of the data.
+That is to say, that we will consider balancing scores as a preprocessing function $b: \reals^{(n + m) \times k} \mapsto \reals^{n + m}$.}
+\parencite{rosenbaum_central_1983} are a versatile class of functions $b(X)$ satisfying the property
+\begin{equation}
+  \mat{X} \perp \vec{z} | b(\mat{X}), \label{eqn:defbalancingscore}
+\end{equation}
+where the matrix $\mat{X} \in \reals^{(n + m) \times k} = \mat{X}_T \cup \mat{X}_C$, and the vector $\vec{z} \in \reals^{n + m}$ contains binary treatment group assignments.
+In words, given the balancing score $b(\mat{X})$, the distribution of the feature vector $\mat{X}$ is independent of treatment assignment $\vec{z}$.
+These methods each set $d$ equal to the $L^1$ norm, or absolute, distance by convention.
+In the notation of~\eqref{eqn:abstractdistance}, then, balancing scores compose the class of distance measures of the form
+\begin{equation}
+  \mathcal{D}_{L^1}(b(\mat{X})_T, b(\mat{X})_C).
+\end{equation}
 
-\subsubsection{(Coarsened) Exact Matching}
+TODO DISCUSS STRONG IGNORABILITY, PROPERTIES, WHEN IS ATE ESTIMATE UNBIASED
++ TODO MAKE CLEAR IT IS REPLICATING A RANDOMIZED TRIAL
 
-\begin{itemize}
-  \item Iacus King Porro, 2011: MIB methods
-  \item Iacus King Porro, 2012: Causal inference without balance checking
-\end{itemize}
+\paragraph{The Propensity Score.} In the same paper, \textcite{rosenbaum_central_1983} propose the balancing score $b(\mat{X}) = \mathbbmss E \left [\vec{z} | \mat{X}\right ]$, also known as the propensity score.
+This is simply a predicted probability (or logit) that any given feature vector $\vec{x} \in \mat{X}$ belongs to the treatment group.
+This is the most prevalent balancing score in the literature, and indeed the most prevalent distance measurement overall.
+It is almost always estimated by logistic or probit regression \parencite{garrido_methods_2014}, but other classifiers are possible.
+By preprocessing $\mat{X}$ to a single dimension of scores $b(\mat{X})$, we not only reduce computation time necessary for $\mathcal{D}$, but also implicitly apply a weighting to the feature matrix $\mat{X}$: the features most predictive of treatment assignment (i.e., most heterogeneous between the two groups) will be the most closely matched.
+\cite{dehejia_causal_1999} evaluated the performance of propensity score methods, and found that, given the assumption that treatment depends only on pre-intervention observable features, propensity score methods can serve as a useful diagnostic, particularly for examining the degree of ``overlap'' in feature distributions between the two groups.
+
+\paragraph{Extensions of the Propensity Score.} Acknowledging that propensity score methods have thus far been limited to binary, multinomial, or ordinal treatment assignments in the literature (as they are considered in this paper), \textcite{imai_causal_2004} explored extensions of the propensity score into the realm of quantitative treatments.
+In their paper, they consider the treatment vector \emph{packyear}, the number of packs smoked by a smoker each year, and subclass the resulting scores into a varying number of bins.
+They found that conditioning on the subclass helped to reduce bias by between 16\% and 95\%.
+
+\paragraph{The Prognostic Score.} \textcite{hansen_prognostic_2008} discussed the \emph{prognostic score}, a balancing metric on outcomes.
+The only difference between the construction in~\eqref{eqn:defbalancingscore} is that we exchange $\vec{z}$ for $\vec{y}$, a vector of outcomes.
+\begin{equation}
+  \mat{X} \perp \vec{y} | b(\mat{X}) \iff b \text{ is a prognostic balancing score} \label{eqn:prognosticscoredef}
+\end{equation}
+The pgognostic score is constructed nearly in exactly the same way as the propensity score; however, in order to preserve~\eqref{eqn:prognosticscoredef}, the model (linear/logistic regression, random forest, etc.) is trained on the \emph{control data alone}.
+This model is then used to predict scores over the \emph{whole dataset}.
+
+The idea of using outcomes, either raw or ``preprocessed'', does not see unanimous support; \textcite{garrido_methods_2014} argues against incorporating any information about outcome into the matching process.
+\textcite{miettinen_stratification_1976} was a seminal paper in matching that used outcome stratification; this method has since been found to be suboptimal, and has been effectively replaced by prognostic scores and other methods \parencite{hansen_bias_2006} (Miettinen's ``multivariate confounder score'' used the \emph{full} dataset, not just the control, when fitting).
+\textcite{hansen_bias_2006} argues for its inclusion via a ``conditionality principle'', which suggests that if some statistic (in this case, $b(\mat{X})$) is known to be uninformative about the parameter being estimated (i.e.
+treatment effects), it should be included in inferential models.
+
+\textcite{stuart_prognostic_2013} found prognostic scores to be highly correlated with bias in estimates of treatment effects, and thus a useful diagnostic tool, even when the model was misspecified.
+Thus, prognostic scores can serve as a useful ``proxy'' for the reduction in bias of the estimates.
+Nonetheless, they also caution that their use in matching depends on the researcher's appetite to reduce the separation of data identification and analysis of outcomes in their study.
+
+\paragraph{Joint Use of Balancing Scores.} \textcite{leacy_joint_2014} further explores of calculating \emph{both} propensity and prognostic scores, stratifying each into a $k \times k$ grid, and considering observations within each of the $k^2$ grid squares as a \emph{matched strata}.
+However, they found that matching on prognostic scores alone performed the best, while matching on propensity scores and Mahalanobis distance (see \S\ref{subsubsec:normdist}) also performed admirably.
+The $k \times k$ strata performed better than stratification of either score alone, but could not beat matching on either un-discretized score alone.
+They also found that the percent bias of estimates created using this matching method increased with the non-linearity in the outcomes model.
+
+\paragraph{Calipers.} Propensity and prognostic scoring methods often add one more hyperparameter: a \emph{caliper}, $c$.
+The caliper is used to calculate \emph{caliper width}, equal to the product of $c$ and the standard deviation of the scores \parencite{ho_matchit_2011}.
+The caliper serves to establish a radius of acceptable matches around each observation's propensity score as a form of quality control: if the observations are too dissimilar (their distance is greater than the caliper width), then they may not match.
+This comes with a cost: if the caliper is sufficiently small, observations with outlying feature vectors $\vec{x}$ may not have any suitable matches, and the matched subset remaining for treatment effect analyses may be substantially smaller.
+
+\textcite{austin_optimal_2011} explored the question of choosing an optimal caliper width under a variety of regimes.
+When all features were iid Gaussian random variables, calipers $c \in [0.05, 0.15]$ maximized the reduction in bias, with each reduction on the order of $98.9\%$ or greater.
+\citeauthor{austin_optimal_2011} then introduced correlation when generating the feature matrix, as well as including a varying number of independent $\text{Bernoulli}(0.5)$.
+Under the correlated regime, the optimal caliper varied between $0.05$ and $0.30$, depending on the true risk reduction given the generated features.
+However, he also notes that the percent bias reduction decreased as the number of Bernoulli features increased, and that when \emph{all} features were Bernoulli-distributed, the choice of caliper had no effect on bias reduction.
+
+\paragraph{Criticism.} The most notable criticism of balancing score based distance measurement comes from \textcite{king_why_2019}.
+King and Nielsen's argument is twofold: first, from an experimental design perspective, they argue that \emph{fully blocked} study design, which other distance measures attempt to emulate (in particular: Mahalanobis distance, Coarsened Exact Matching), dominates \emph{complete randomization study design}, which balancing-score based methods attempt to emulate.
+Whereas each study design balances unobserved confounding features on average, fully blocked methods assure balance of observed confounding features, whereas complete randomization only promises balance \emph{on average}.
+They thus regard balancing score methods as having a ``lower standard''.
+Second, they observe the ``paradox of Propensity Score Matching'': consider the case where $\mat{X} \perp \vec{z}$, and so all predicted scores are the same constant: $\hat{\vec{z}}$.
+Then, all potential matches are indecipherable, and the individual matchings may be complete nonsensical.
+
+TODO VERIFY
+
+\subsubsection{Exact Matching and Almost Exact Matching}
+\label{subsubsec:exactmatching}
+
+Among other methods which seek to replicate fully blocked study design, \textcite{king_why_2019} argue for ``coarsened exact matching'', or CEM.
+This requires the researcher to first preprocess their data into discrete categories: for example, a vector ``age'' might be grouped into buckets.
+They then suggest exact matching on the now-discretized features into \emph{strata}, and applying observational weights to the data according to the formula
+\begin{equation}
+  w_{i}= \begin{cases}1, & i \in \text{treatment group} \\ \frac{m_{C}}{m_{T}} \frac{m_{T}^{s}}{m_{C}^{s}}, & i \in \text{control group} \end{cases},
+\end{equation}
+where $w_i$ denotes the observation weight for observation $i$, $m_T$ and $m_C$ are the total number of matched treatment and control observations, and $m_T^s$ and $m_C^s$ are the total number of matched treatment and control observations \emph{within observation $i$'s stratum, $s$} \parencite{iacus_causal_2012}.
+
+\textcite{iacus_multivariate_2011} introduces a formal definition for a class of distance measures which they call ``monotonic imbalance bounding'', or MIB.
+These methods are designed to require no assumptions about the distribution of the data, and to emphasize minimizing \emph{in-sample} imbalance, as opposed to \emph{sample} balance.
+These methods take a vector of tuning parameters $\vec{\uppi}$, such as a vector of calipers.
+These methods afford the benefit that the number of matches is a function of the tuning parameters alone, and given monotonicity of the maximum distance function $\gamma_{D_{d}}(\vec{\uppi})$, can be each adjusted independently other others: if $\gamma$ is vector-valued, returning a vector in $\reals^k$, the distance between each feature vector can be independently compared in a sequence of inequalities.
+CEM is an example of an MIB distance measure, but this class is clearly much more flexible, allowing for precise adjustments.
+
+\paragraph{Almost Exact matching.} Some of the most modern methods are DAME \parencite{liu_interpretable_2019} and FLAME \cite{wang_flame_2021}, which incorporate machine-learning generated weight vectors indicating the relative importance of features to extend exact matching into the field of \emph{almost exact matching}.
+These two algorithms are unified by their tendency to sometimes ``drop'' uninformative features from the matching process, in order to ensure that the most important features are matched on, and potentially increase the number of potential matches.
+
+TODO FINISH
+
+\subsubsection{Norms and Mahalanobis Distance}
+\label{subsubsec:normdist}
+
+Norms of the difference between feature vectors (i.e.
+$\| \vec{x}_T - \vec{x}_C \|_p$) themselves are usually used in the form of Mahalanobis distance, which first rescales all features to have mean zero and unit variance, and then calculates distances using the $L^2$ norm.
+This ensures that all features have equal weight in the matching, which may or may not be desirable; in either case, it almost certainly better than the no-rescaling case, where a feature's variance would determine its relative importance.
+As \textcite{king_why_2019} notes, norm-based distance metrics attempt to replicate fully-blocked design, and thus offer some theoretical advantages.
+However, many are attracted to the ability of balancing score methods, as well as modern methods such as DAME and FLAME, to discern which features are \emph{most important} to match on.
+
+Nonetheless, \textcite{leacy_joint_2014} finds Mahalanobis distance to perform well when the number of features was small, each of which was approximately normally distributed.
+Additionally, many modern matching software packages allow for initially establishing a ``perimeter'' of viable matches using a caliper and a propensity score, and subsequently matching on the Mahalanobis distance between the feature vectors \parencite{ho_matchit_2011}.
+
+\subsection{Matching Algorithms}
+
+Once the distance measure has been agreed upon, the researcher must then decide \emph{how} to match.
+A computer scientist could describe this problem as finding the minimum weight full matching of a \emph{bipartite graph}, whose nodes represent observation IDs, and whose edge weights represent the distance between them.
+Here, the term ``bipartite'' means that all edges in the graph map $T \leftrightarrow C$; within each disjoint subset, no two nodes share an edge.
+This paper will explore the optimal solution to this problem in both the $1:1$ and $1:k$ matching scenarios, as well as fast approximations.
+
+\paragraph{Optimal Matching.} \textcite{rosenbaum_optimal_1989} argues for optimal matching according to the Hungarian algorithm, also known as the Kuhn-Munkres or simply Munkres algorithm \parencite{munkres_algorithms_1957}.
+We will denote a matching a set $\set{M}$ of tuples $(i, j)$, with $i$ representing a treatment-group observation ID, and $j$ a control-group observation ID.
+The Hungarian algorithm solves the \emph{linear assignment problem},which in this setting may be expressed as
+\begin{equation}
+  \min_{\set{M}} \sum_{i, j \in \set{M}} d(\vec{x}_i, \vec{x}_j), \text{ such that } \forall (i, j), (k, l) \in \set{M}, i = k \iff j = k;
+\end{equation}
+the ``such that'' condition simply specifies that the matching is one-to-one.
+Optimal matching is the true minimizer of the objective function, and can be extended to the $1:k$ matching case via $b$-matching while retaining optimality.
+This process involves adding $k$ ``copy'' nodes for each node representing a $T$ datapoint, each with identical edges as the original node, and then perform the optimal matching on the resulting graph.
+However, as \textcite{khan_efficient_2016} notes, this can lead to calculations which quickly become time consuming as the total number of data points increases.
+\textcite{khan_efficient_2016} goes on further to explore efficient approximation algorithms of the $b$-matching, such as the $b$-\textsc{Suitor} algorithm.
+
+\paragraph{Greedy Matching.} Greedy algorithms have the benefit of running much quicker than optimal matching algorithms, and are often the default in other observational data matching software, such as \texttt{MatchIt} \parencite{ho_matchit_2011}.
+These methods are equivalent to sorting the edges of the graph in ascending order, and naively popping off the top of the list the next match.
+While these methods benefit from speed, they guarantee nothing with regards to optimality.
+This can be particularly problematic in the $1:k$ scenario \cite{rosenbaum_central_1983}, where the order in matches are assigned can have matches implications; compare this to optimal matching, which is a simultaneous match.
 
@@
 
-\subsection{Matching Algorithms}
+Researchers may wish to assess the balance of their matching, particularly when attempting to heuristically optimize a hyperparameter, such as the caliper.
+Common numerical methods include standardized mean-differences, variance ratios, and empirical CDF (eCDF) comparison between the two groups \parencite{greifer_assessing_2022}.
+Most researchers do not check for convergence of moments beyond the second, and as \textcite{basu_use_2008} notes, these moments can take much longer (i.e.
+require more data) to converge, the effects of which can be particularly detrimental when outcome depends on some non-linear function of the joint distribution of $\mat{X}$.
+For this reason, \textcite{zhu_kernel-based_2018} recommend using density-based metrics, such as total variation distance or Kolmogorov-Smirnov distance.
+It is also important to note that balance of the $i$th moment does not imply balance of the $j$th moment across the two groups for $j \geq i$ \parencite{garrido_methods_2014}.
+
+There are also many graphical diagnostics for assessing balance.
+QQ-plots, eCDFs, and density estimates for for the unmatched data and the matched subset, perhaps overlaid, effectively visually convey the balance improvement from matching.
+One may also plot the improvement in standardized mean differences between the two groups, referred to as a Love plot \parencite{greifer_assessing_2022}.
+One could also plot the bipartite graph to understand where clusters are appearing, or how large groups of ``similar'' observations are, which observations are most ``dissimilar'' from the rest.
+
+\section{\texttt{matching}: a Python Package for Matching Observational Data}
+
+The most popular implementation of bipartite matching for observational studies is the \texttt{MatchIt} package for R \parencite{ho_matchit_2011}.
+The GitHub repository\footnote{\url{https://github.com/kosukeimai/MatchIt}} boasts an impressive $37000$ downloads per month, and with good reason: the authors of the package are also authors of many of the most popular papers in the field of matching \parencite{ho_matching_2007, imai_causal_2004, king_why_2019, stuart_matching_2010}.
+
+\texttt{MatchIt} is largely used through a single function: \texttt{Matchit::matchit}, which minimally accepts a \texttt{formula} (like those used in \texttt{lm}), \texttt{data}, and a \texttt{method}.
+The function also allows for iterative matching procedures, complex filters on the suitability of potential matches, and a selection of bipartite matching algorithms.
+The output of the function is furthermore easily inspected via the ubiquitous \texttt{summary} function, or plotted via \texttt{plot}.
+
+However, R is not as popular ``in-industry'' as it once was.
+Whereas academics have largely favored R as their programming language of choice, Python has become the \emph{lingua franca} of data scientists.
+The TIOBE index\footnote{\url{https://www.tiobe.com/tiobe-index/}} currently lists Python as the most popular programming language in the world, with nearly ten times the market share of R.
+It is then surprising that there is no package for matching observational data that is as flexible and feature-complete as \texttt{MatchIt}.
+This is what lead me to developing \texttt{matching}, a Python library for matching observational data.
+
+\subsection{Design Goals}
+
+After exploring \texttt{MatchIt}'s capabilities, as well as the Python landscape of matching packages for causal inference, I set out to create a package that is:
+\begin{itemize}
+  \item inspectable. The user should be able to numerically and graphically inspect each step of the matching procedure.
+  \item flexible. The package should provide ``plug-and-play'' tools to customize and conduct an arbitrarily complex matching procedure, without the user needing to implement new classes.
+  \item familiar.
+The package should mimic idioms from popular data science packages\footnote{e.g.
+\texttt{numpy}, \texttt{pandas}, \texttt{sklearn}, \texttt{networkx}}, and all outputs should be easily coercible to a \texttt{pandas.DataFrame} or a \texttt{numpy.ndarray} for further analysis.
+  \item purpose-built. The package should only implement matching procedures, leaving model-building and effect estimation to the user.
+    The user likely already knows what model they wish to fit to the data, and predicting every foreseeable model is difficult.
+    Narrowing in on the matching procedure alone also helps with the first three goals.
+\end{itemize}
+In particular, I focused on separating responsibility between how the distance between feature vectors is calculated, and how the matches are then assigned.
+Being able to easily adjust distance measure and matching procedure independently is especially valuable when assessing the matching quality via diagnostic tools when tuning a hyperparameter, e.g.
+the maximum allowable distance, or the number of matches $k$ to find for each treatment observation.
+
+\subsection{Architecture}
+
+The package is organized into the following modules.
+\begin{itemize}
+  \item \texttt{matching.balance}: implementations of common balance measures, such as standardized mean difference, variance ratio, and eCDFs.
+  \item \texttt{matching.dataset}: implementation of the \texttt{MatchingDataset} class, which parses user data into an easily manageable format.
+  \item \texttt{matching.distance}: the \texttt{\_Distance} base class and concrete implementations of distance measures, such as \texttt{Norm}.
+  \item \texttt{matching.graph}: the \texttt{MatchingGraph} class implementation; this is the main class that the user interacts with.
+  \item \texttt{matching.graph\_utils}: functional graph utilities.
+  \item \texttt{matching.preprocessing}: functional preprocessing utilities, such as propensity scoring, prognostic scoring, and auto-coarsening.
+\end{itemize}
+
+Most users will only need to use \texttt{matching.graph} and \texttt{matching.distance}, perhaps in addition to some light preprocessing with \texttt{matching.preprocessing} as necessary.
+These two submodules represent the algorithmic matching, and distance calculations separately.
+This separation of responsibility into two separate class hierarchies allows for high flexibility in the matching procedure, as each component can be changed independent of the other.
+
+\begin{figure}[h!]
+  \center{%
+    \includegraphics[width=\textwidth]{../docs/mermaid/img/relationships.png}%
+  }
+  \caption{\label{fig:arch_flow} The relationships between the different modules of \texttt{matching}}.
+\end{figure}
+
+\begin{figure}[h!]
+  \center{%
+    \includegraphics[width=\textwidth]{../docs/mermaid/img/distance_uml.png}%
+  }
+  \caption{\label{fig:distance_uml} UML Diagram for the \texttt{matching.distance} submodule}
+\end{figure}
+
+\newpage
+\subsection{Usage}
+\label{subsec:usage}
+
+Typical usage of the \texttt{Matching} package is outlined in Figure~\ref{fig:matching_usage}. A written explanation of each step follows.
+
+\begin{figure}[h!]
+  \center{%
+    \includegraphics[width=\textwidth]{../docs/mermaid/img/usage_flow.png}%
+  }
+  \caption{\label{fig:matching_usage} Flow diagram describing typical usage of the \texttt{matching} package}
+\end{figure}
+
+\paragraph{Initialization.} 
+The \texttt{MatchingGraph} is minimally initialized with an array-like (e.g., a \texttt{pandas.DataFrame}, or a \texttt{numpy.ndarray}) of features \texttt{X} and an array-like of binary treatment assignments \texttt{z}.
+The user may first wish to preprocess their data (i.e.
+calculating propensity scores) using \texttt{matching.preprocessing}, but this is not necessary.
+\begin{minted}[linenos, fontsize=\small]{python3}
+from matching.graph import MatchingGraph
+from matching.preprocessing import propensity_score
+
+# Possibly some preprocessing... (not shown)
+# X is an array of propensity scores, z is an array of booleans
+mg = MatchingGraph(X, z)
+\end{minted}
+\paragraph{Setting Edges.} The user should then make a call to the \texttt{set\_edges} method to calculate the distance between each observation the two groups, and set the graph nodes and edge weights.
+Here, the user supplies a \texttt{\_Distance} measure from \texttt{matching.distance}.
+The user may supply a \texttt{min\_distance} and/or a \texttt{max\_distance} for an edge to be allowed; in graph terms, this can be considered the edges' \emph{capacity}.
+\begin{minted}[linenos, fontsize=\small]{python3}
+import numpy as np
+from matching.distance import L1Norm
+
+# Calculating the caliper width. Recall that X is an array of scores
+caliper = 0.05
+caliper_width = caliper * np.std(X)
+mg.set_edges(distance=L1Norm(max_distance=caliper_width))
+\end{minted}
+\paragraph{Iterative Filtering.} It is possible to implement iterative distance measures as well.
+Consider \texttt{X} as a dataframe with columns \texttt{"score"}, containing propensity scores, and \texttt{"is\_nice"}, a boolean series.
+We would like to match on propensity scores within \texttt{caliper\_width}, but also cannot assign any nice people to any naughty people.
+We can make use of the \texttt{filter\_edges} method and the \texttt{include} keyword argument to accomplish this.
+\begin{minted}[linenos, fontsize=\small]{python3}
+from matching.distance import Exact, L1Norm
+from matching.graph import MatchingGraph
+from matching.preprocessing import propensity_score
+
+# Take caliper_width as given
+# X has columns "score" and "is_nice", z is treatment assignments
+mg = MatchingGraph(X, z)
+# NOTE: there is also a keyword argument "exclude"
+mg.set_edges(distance=L1Norm(max_distance=caliper_width), include=["score"])
+mg.filter_edges(distance=Exact(), include=["is_nice"])
+\end{minted}
+The \texttt{filter\_edges} command will drop any potential matches who do not match exactly on \texttt{"is\_nice"} by simply deleting the edge between them.
+This extends beyond exact-match filtering.
+One could filter using \texttt{L2Norm}, for instance, to drop potential matches whose $L^2$ norm distance was above some maximum threshold.
+
+There are also methods to \texttt{filter\_nodes} by label or order, as well as to \texttt{filter\_subgraphs}, which filters the disjoint subgraphs of the graph by a variety of metrics; they will not be discussed here, but readers are referred to \texttt{matching} documentation hosted on the GitHub\footnote{\url{https://github.com/jackpotrykus/propensity-score-matching-thesis}}.
+
+\paragraph{Matching.} Once the user has set the edges of the graph and completed their filtering, they may wish to conduct a $1:k$ matching between treatment and control, for some integer $k \geq 1$\footnote{Note that this is not always necessary, particularly in the case of coarsened exact matching (CEM)}.
+The user does this via a call to the \texttt{match} method.
+The user can control $k$ with \texttt{n\_match}, the maximum number of matches to find for each observation.
+Some treatment observations will not be able to be matched with this many control observations; they can be automatically pruned with the \texttt{min\_match} parameter.
+The parameter \texttt{replace} is a boolean, indicating whether multiple treatment observations may match with the same control observation, which is \texttt{False} by default.
+Finally, the user has a choice of greedy (keyword argument: \texttt{"greedy"} or \texttt{"fast"}) or optimal matches (keyword argument: \texttt{"optimal"}, \texttt{"hungarian"}, \texttt{"kuhn"}, or \texttt{"munkres"}).
+\begin{minted}[linenos, fontsize=\small]{python3}
+# Now we should match the graph
+# By default: n_match=1, min_match=1, replace=False
+mg.match(n_match=2, min_match=2, method="optimal")
+\end{minted}
+
+\paragraph{Balance Assessment.} The user can then compare the balance of the \texttt{input\_data} and the \texttt{match\_data}\footnote{In the case of Exact matching edges, the match\_data should consist of all subgraphs of order at least 2}.
+Note that \texttt{match\_data} is a subset of \texttt{input\_data}, plus a new column, \texttt{"match\_group"}.
+\begin{minted}[linenos, fontsize=\small]{python3}
+# Can get dataframes of input_data and match_data
+input_df = mg.input_data.frame
+match_df = mg.match_data.frame
+
+# Assess balance improvement via the MatchingDataset.summary() method
+input_balance = mg.input_data.summary()
+match_balance = mg.match_data.summary()
+
+# ... or import the balance functions themselves
+# Let's compare the eCDF of scores between the input_data and the match_data
+from matplotlib import pyplot as plt
+import numpy as np
+from matching.balance import ecdf
+
+# Fit eCDFs to the scores in each dataframe
+input_score_ecdf = ecdf(input_df["scores"])
+match_score_ecdf = ecdf(match_df["scores"])
+
+# Evaluate the eCDFs at 101 points between 0 and 1
+xs = np.linspace(0, 1, 101)
+input_score_ecdf_at_xs = input_score_ecdf(xs)
+match_score_ecdf_at_xs = match_score_ecdf(xs)
+
+# Plot them to compare
+plt.plot(xs, input_score_ecdf_at_xs)
+plt.plot(xs, match_score_ecdf_at_xs)
+plt.show()
+\end{minted}
+
+\subsection{Comparison to Other Causal Inference Packages}
+
+Within the Python landscape, there are three existing packages capable of matching for causal inference: \texttt{DoWhy}, \texttt{pymatch}, and \texttt{dame-flame}.
+None of them directly use a graph data structure as their primary data structure for conducting the matching procedure, something which makes \texttt{matching} unique.
+\texttt{matching} also offers the simplest API for iterative matching procedures, as described in \S\ref{subsec:usage}.
+
+Below, I discuss the package in decreasing order of number of stars on its GitHub repo, as a proxy variable for its prevalence among Python data scientists alone (not necessarily a marker of quality!).
+
+\paragraph{\texttt{DoWhy}.} An ``End-to-End Library for Causal Inference'' developed by Microsoft \parencite{dowhypaper}, \texttt{DoWhy} supports matching on a variety of distance metrics, in addition to numerous other methods in the field of Causal Inference (e.g.
+instrumental variables), and can match the data, build a predictive model for outcomes, and estimate the treatment effect in one shot.
+The main drawback of this approach is that matching procedures\footnote{e.g.
+\texttt{dowhy.causal\_estimators\_propensity\_score\_estimator}} are rigid, ``black-box'' steps that occur as part of treatment effect estimation.
+The matching is neither easy to inspect nor flexibly alter without sub-classing \texttt{CausalEstimator}.
+In short, \texttt{DoWhy} is concerned with estimating treatment effects, and treats the matching procedure as a means to this end, not a process of note in itself.
+\texttt{matching} could be used to pre-subset the observational data before passing it to \texttt{DoWhy} for further analysis, but the fundamental aims of these packages differ: \texttt{DoWhy} is a high-level API for end-to-end causal inference, whereas \texttt{matching} enables low-level exploration of the matching process.
+
+\paragraph{\texttt{pymatch}.} Purpose-built by researchers working on an observational study \parencite{miroglio_pymatch_2022}, \texttt{pymatch} supports only one distance metric, the propensity score, and the only matching algorithms on offer are ``random'' and ``greedy''; no optimal matching is available.
+The package also lacks support iterative matching procedures.
+The codebase is short and succinct, and certainly works well for the specific use-case the researchers intended it for, but it is not feature-complete nor flexible enough to be considered a proper alternative to \texttt{MatchIt} or \texttt{matching}.
+
+\paragraph{\texttt{dame-flame}.} Developed by Duke's \emph{Almost Matching Exactly Lab} \parencite{gupta_dame-flame_2021}, \texttt{dame-flame} implements its two titular matching procedures: DAME \parencite{liu_interpretable_2019} and FLAME \parencite{wang_flame_2021}.
+As discussed in \S\ref{subsubsec:exactmatching}, these methods extend CEM using machine learning, calculating a weight vector \vec{w} indicating the relative importance of matching on each feature.
+These algorithms each run fast, and are easier to inspect than \texttt{DoWhy}.
+However, since these each seek to replicate CEM, and since these are the \emph{only} matching methods provided by \texttt{dame-flame}, this means that \texttt{dame-flame} can only be used with discrete datasets; for users unwilling to coarsen/discretize their data, or users seeking a matching procedure other than DAME and FLAME, \texttt{dame-flame} is not an option.
+\texttt{matching} does not currently support these algorithms, but they are ``in-the-works''; it can certainly be implemented as a new \texttt{matching.distance.\_Distance} subclass.
 

### 2022-04-29 matching/balance.py

@@ -5,7 +5,9 @@ from numpy import typing as npt
 def standardized_mean_difference(a: npt.ArrayLike, b: npt.ArrayLike):
     """Computes the standardized mean difference between a and b.
 
-    .. math:: \frac{\overline{a} - \overline{b}}{\sqrt{\text{Var}(a) + \text{Var}(b)}}
+    .. math::
+
+        \\frac{\\overline{a} - \\overline{b}}{\\sqrt{\\text{Var}(a) + \\text{Var}(b)}}
 
     Parameters
     ----------
@@ -58,7 +60,10 @@ def ecdf(x: npt.ArrayLike):
 
 def variance_ratio(a: npt.ArrayLike, b: npt.ArrayLike) -> float:
     """Compute the variance ratio between `a` and `b`, defined as
-    .. math:: \frac{\text{Var}(a)}{\text{Var}(b)}
+
+    .. math:: 
+
+        \\frac{\\text{Var}(a)}{\\text{Var}(b)}
 
     Parameters
     ----------

### 2022-04-29 matching/graph.py

@@ -60,12 +60,14 @@ _MatchingGraph = TypeVar("_MatchingGraph", bound="MatchingGraph")
 @dataclass(init=False)
 class MatchingGraph:
     """:class:`MatchingGraph` offers a simple framework to
+
     - calculate a distance metric between two sides of a bipartite graph,
     - iteratively filter this graph by enforcing other constraints on nodes and edges,
     - calculate a "matched" subset of this graph via greedy, grouping, or optimal matching strategies, and
     - compare the balance between input and matched datasets.
 
     At construction, the user supplies three `ArrayLike`s:
+
     - `X`, an array of features to match on;
     - `z`, an array of binary treatment assignments; and
     - `ids`, an optional array of observation IDs. If not supplied, default IDs are enumerated from 0 upwards.
@@ -77,6 +79,7 @@ class MatchingGraph:
     Once the edges are calculated, the user can then use one of the matching strategies provided to "match" their
     dataset. Alternatively, the user can proceed to iteratively filter the graph further, via the `filter_edges` and
     `filter_nodes`.
+
     - `filter_edges` allows the user to enforce additional distance metrics on the data, while preserving the edge
     weights from `set_edges`. For example, a user may calculate edge weights on a propensity score, but then require
     an exact match on certain categorical features, as well as a cutoff for the Mahalanobis distance between the feature
@@ -524,6 +527,7 @@ class MatchingGraph:
             match with the same control group observation. `False` by default
         method : Union[str, MatchingMethod], default="greedy"
             Matching method to use. Valid options are (case-insensitive):
+
             - "greedy" == "fast"
             - "optimal" == "hungarian" == "kuhn" == "munkres"
 

### 2022-05-02 paper/main.tex
updated draft -- most everything done except experiments + conclusion

@@
 \end{equation}
-The pgognostic score is constructed nearly in exactly the same way as the propensity score; however, in order to preserve~\eqref{eqn:prognosticscoredef}, the model (linear/logistic regression, random forest, etc.) is trained on the \emph{control data alone}.
+The prognostic score is constructed nearly in exactly the same way as the propensity score; however, in order to preserve~\eqref{eqn:prognosticscoredef}, the model (linear/logistic regression, random forest, etc.) is trained on the \emph{control data alone}.
 This model is then used to predict scores over the \emph{whole dataset}.
@@
 Then, all potential matches are indecipherable, and the individual matchings may be complete nonsensical.
+The authors nonetheless concede that propensity scores (and other balancing scores, by extension) do indeed have nice theoretical properties, and may serve as a useful diagnostic; they contend only that they should not be used for matching.
 
@@
 
-\paragraph{Almost Exact matching.} Some of the most modern methods are DAME \parencite{liu_interpretable_2019} and FLAME \cite{wang_flame_2021}, which incorporate machine-learning generated weight vectors indicating the relative importance of features to extend exact matching into the field of \emph{almost exact matching}.
+\paragraph{Almost Exact Matching.} Some of the most modern methods are DAME \parencite{liu_interpretable_2019} and FLAME \cite{wang_flame_2021}, which incorporate machine-learning generated weight vectors indicating the relative importance of features to extend exact matching into the field of \emph{almost exact matching}.
 These two algorithms are unified by their tendency to sometimes ``drop'' uninformative features from the matching process, in order to ensure that the most important features are matched on, and potentially increase the number of potential matches.
@@
 
-\subsubsection{Norms and Mahalanobis Distance}
+\subsubsection{$L^p$ Norms and Mahalanobis Distance}
 \label{subsubsec:normdist}
@@
 
-\paragraph{Optimal Matching.} \textcite{rosenbaum_optimal_1989} argues for optimal matching according to the Hungarian algorithm, also known as the Kuhn-Munkres or simply Munkres algorithm \parencite{munkres_algorithms_1957}.
+\subsubsection{Optimal Matching}
+
+\textcite{rosenbaum_optimal_1989} argues for optimal matching according to the Hungarian algorithm, also known as the Kuhn-Munkres or simply Munkres algorithm \parencite{munkres_algorithms_1957}.
 We will denote a matching a set $\set{M}$ of tuples $(i, j)$, with $i$ representing a treatment-group observation ID, and $j$ a control-group observation ID.
@@
 \begin{equation}
-  \min_{\set{M}} \sum_{i, j \in \set{M}} d(\vec{x}_i, \vec{x}_j), \text{ such that } \forall (i, j), (k, l) \in \set{M}, i = k \iff j = k;
+  \min_{\set{M}} \sum_{(i, j) \in \set{M}} d(\vec{x}_{T_{i}}, \vec{x}_{C_{j}}), \text{ such that } \forall (i, j), (k, l) \in \set{M}, i = k \iff j = l;
 \end{equation}
 the ``such that'' condition simply specifies that the matching is one-to-one.
-Optimal matching is the true minimizer of the objective function, and can be extended to the $1:k$ matching case via $b$-matching while retaining optimality.
+As such, optimal matching via the Hungarian algorithm minimizes the total sum of distances between matches in the matched subset.
+
+\paragraph{Extensions.} The Hungarian algorithm can be extended to the $1:k$ matching case via $b$-matching while retaining optimality.
 This process involves adding $k$ ``copy'' nodes for each node representing a $T$ datapoint, each with identical edges as the original node, and then perform the optimal matching on the resulting graph.
@@
 
-\paragraph{Greedy Matching.} Greedy algorithms have the benefit of running much quicker than optimal matching algorithms, and are often the default in other observational data matching software, such as \texttt{MatchIt} \parencite{ho_matchit_2011}.
+\subsubsection{Greedy Matching}
+
+Greedy algorithms have the benefit of running much quicker than optimal matching algorithms, and are often the default in other observational data matching software, such as \texttt{MatchIt} \parencite{ho_matchit_2011}.
 These methods are equivalent to sorting the edges of the graph in ascending order, and naively popping off the top of the list the next match.
 While these methods benefit from speed, they guarantee nothing with regards to optimality.
-This can be particularly problematic in the $1:k$ scenario \cite{rosenbaum_central_1983}, where the order in matches are assigned can have matches implications; compare this to optimal matching, which is a simultaneous match.
+This can be particularly problematic in the $1:k$ scenario \parencite{rosenbaum_optimal_1989}, where the order in matches are assigned can have matches implications; compare this to optimal matching, which is a simultaneous match.
 
@@
 Common numerical methods include standardized mean-differences, variance ratios, and empirical CDF (eCDF) comparison between the two groups \parencite{greifer_assessing_2022}.
-Most researchers do not check for convergence of moments beyond the second, and as \textcite{basu_use_2008} notes, these moments can take much longer (i.e.
-require more data) to converge, the effects of which can be particularly detrimental when outcome depends on some non-linear function of the joint distribution of $\mat{X}$.
+Most researchers do not check for convergence of moments beyond the second, and as \textcite{basu_use_2008} report, these moments exhibit much slower convergence, the effects of which can be particularly detrimental when outcome depends on some non-linear function of the joint distribution of $\mat{X}$.
 For this reason, \textcite{zhu_kernel-based_2018} recommend using density-based metrics, such as total variation distance or Kolmogorov-Smirnov distance.
@@
 The most popular implementation of bipartite matching for observational studies is the \texttt{MatchIt} package for R \parencite{ho_matchit_2011}.
-The GitHub repository\footnote{\url{https://github.com/kosukeimai/MatchIt}} boasts an impressive $37000$ downloads per month, and with good reason: the authors of the package are also authors of many of the most popular papers in the field of matching \parencite{ho_matching_2007, imai_causal_2004, king_why_2019, stuart_matching_2010}.
+The GitHub repository\footnote{\url{https://github.com/kosukeimai/MatchIt}} boasts an impressive $37000$ downloads per month, and with good reason: the authors of the package are also authors of many of the most popular papers in the field of matching (e.g.
+\textcite{ho_matching_2007, imai_causal_2004, king_why_2019, stuart_matching_2010}).
 
@@
 \texttt{numpy}, \texttt{pandas}, \texttt{sklearn}, \texttt{networkx}}, and all outputs should be easily coercible to a \texttt{pandas.DataFrame} or a \texttt{numpy.ndarray} for further analysis.
-  \item purpose-built. The package should only implement matching procedures, leaving model-building and effect estimation to the user.
-    The user likely already knows what model they wish to fit to the data, and predicting every foreseeable model is difficult.
+  \item ``depth over breadth''.
+The package should only implement matching procedures, leaving model-building and effect estimation to the user.
 Narrowing in on the matching procedure alone also helps with the first three goals.
@@
 
-There are also methods to \texttt{filter\_nodes} by label or order, as well as to \texttt{filter\_subgraphs}, which filters the disjoint subgraphs of the graph by a variety of metrics; they will not be discussed here, but readers are referred to \texttt{matching} documentation hosted on the GitHub\footnote{\url{https://github.com/jackpotrykus/propensity-score-matching-thesis}}.
+There are also methods to \texttt{filter\_nodes} by label or order, as well as to \texttt{filter\_subgraphs}, which filters the disjoint subgraphs of the graph by a variety of metrics\footnote{\texttt{filter\_subgraphs} is particularly relevant for strata-based matching methods like CEM, which often filter strata based on the total number of observations (from each group), the ratio of treatment to control}; they will not be discussed here, but readers are referred to \texttt{matching} documentation hosted on the GitHub\footnote{\url{https://github.com/jackpotrykus/propensity-score-matching-thesis}}.
 
-\paragraph{Matching.} Once the user has set the edges of the graph and completed their filtering, they may wish to conduct a $1:k$ matching between treatment and control, for some integer $k \geq 1$\footnote{Note that this is not always necessary, particularly in the case of coarsened exact matching (CEM)}.
+\paragraph{Matching.} Once the user has set the edges of the graph and completed their filtering, they may wish to conduct a $1:k$ matching between treatment and control, for some integer $k \geq 1$\footnote{This is not always necessary, particularly in the case of CEM}.
 The user does this via a call to the \texttt{match} method.
@@
 
-\paragraph{Balance Assessment.} The user can then compare the balance of the \texttt{input\_data} and the \texttt{match\_data}\footnote{In the case of Exact matching edges, the match\_data should consist of all subgraphs of order at least 2}.
+\paragraph{Balance Assessment.} The user can then compare the balance of the \texttt{input\_data} and the \texttt{match\_data}\footnote{In the case of Exact matching edges, the \texttt{match\_data} should consist of all subgraphs of order at least 2}.
 Note that \texttt{match\_data} is a subset of \texttt{input\_data}, plus a new column, \texttt{"match\_group"}.
@@
 
+\cleardoublepage
+\section*{Addenda}
+\addcontentsline{toc}{section}{Addenda}
+
 % NOTE: SEE LINK BELOW

### 2022-05-03 matching/distance.py
minor tweaks; reintroduced propensity-score distance

@@ -244,50 +244,49 @@ Euclidean = L2Norm
 
 
 # TODO: Refactor the below
-# from sklearn.preprocessing import StandardScaler
-# @dataclass
-# class Mahalanobis(Norm):
-#     """Mahalanobis distance. This is equivalent to standard-scaling all vectors to have mean 0 and variance 1, and then
-#     using :class:`L2Norm` distance on the standard-scaled differences
+from sklearn.preprocessing import StandardScaler
+from sklearn.linear_model import LogisticRegression
+from matching import preprocessing
+from matching._typing import sklearn_types as skt
 
-#     Attributes
-#     ----------
-#     p: Union[int, float]
-#         Which LpNorm to use. Either an integer value, or infinity. This is pre-set to 2 for Mahalanobis norm.
-#     max_distance: float
-#         Max allowable edge weight. Can be used as a tuning parameter for identifying similar groups
-#     tol: float
-#         A small float. Edge weights of 0 (perfect matches) are replaced with this value, so that the sparse matrix
-#         representation does not drop perfect matches (normally, 0 = no edge). Does not affect max_distance
-#         See DEFAULT_DISTANCE_TOL for details
-#     """
+@dataclass
+class Mahalanobis(Norm):
+    """Mahalanobis distance. This is equivalent to standard-scaling all vectors to have mean 0 and variance 1, and then
+    using :class:`L2Norm` distance on the standard-scaled differences
 
-#     p: Union[int, float] = field(init=False, default=2)
+    Attributes
+    ----------
+    p: Union[int, float]
+        Which LpNorm to use. Either an integer value, or infinity. This is pre-set to 2 for Mahalanobis norm.
+    max_distance: float
+        Max allowable edge weight. Can be used as a tuning parameter for identifying similar groups
+    tol: float
+        A small float. Edge weights of 0 (perfect matches) are replaced with this value, so that the sparse matrix
+        representation does not drop perfect matches (normally, 0 = no edge). Does not affect max_distance
+        See DEFAULT_DISTANCE_TOL for details
+    """
 
-#     def _preprocess_data(self, data: MatchingDataset) -> MatchingDataset:
-#         data.X = pd.DataFrame(StandardScaler().fit_transform(data.Xarr))
-#         return data
+    model: skt._Preprocessor = field(init=False)
+    p: Union[int, float] = field(init=False, default=2)
 
[…]

### 2022-05-03 matching/graph.py

@@ -2,6 +2,7 @@ from collections import defaultdict
 from dataclasses import dataclass, field
 import enum
 from typing import Dict, Hashable, Iterable, Iterable, List, Optional, TypeVar, Union
+import warnings
 
 import networkx as nx
 from numpy import typing as npt
@@ -629,7 +630,10 @@ class MatchingGraph:
                 ro, co = cid_order, tid_order
 
             # Convert the subgraph into a biadjacency matrix
-            bm = nx.algorithms.bipartite.biadjacency_matrix(H, row_order=ro, column_order=co)
+            with warnings.catch_warnings():
+                # `biadjacency_matrix` will soon return a scipy sparse array (instead of a matrix); we already ensure this
+                warnings.simplefilter(action='ignore', category=FutureWarning)
+                bm = sparse.csr_array(nx.algorithms.bipartite.biadjacency_matrix(H, row_order=ro, column_order=co))
             if not isinstance(bm, (sparse.csr_array, sparse.csr_matrix)):
                 raise nx.NetworkXError(f"Failed to create biadjacency matrix, got {type(bm)=}")
 

### 2022-05-03 experiments/__main__.py
experiments module

@@ -1,0 +1,45 @@
+from experiments import experiments
+import numpy as np
+
+
+def main():
+    experiments.clean_output_base_dir()
+
+    SIZE1 = 250
+    SIZE0 = 750
+    P = 5
+    NITER_PER_DGM = 50
+
+    CALIPERS = np.round(np.linspace(0.05, 0.5, 10), 2)
+
+    # Setting up experiment 1: caliper vs correlations
+    RHOS = np.round(np.linspace(0, 0.8, 9), 2)
+    experiment1 = experiments.CaliperExperiment(
+        experiment_name="caliper_vs_correlation_big",
+        size0=SIZE0,
+        size1=SIZE1,
+        p=P,
+        niter_per_dgm=NITER_PER_DGM,
+        calipers=CALIPERS,
+        rhos=RHOS,
+    )
+    experiment1.run()
+    experiment1.plot_rho_heatmaps()
+
+    # Setting up experiment 2: caliper vs imbalance
+    THETA1S = np.round(np.linspace(0, 1, 11), 2)
+    experiment2 = experiments.CaliperExperiment(
+        experiment_name="caliper_vs_imbalance_big",
+        size0=SIZE0,
+        size1=SIZE1,
+        p=P,
+        niter_per_dgm=NITER_PER_DGM,
+        calipers=CALIPERS,
+        theta1s=THETA1S,
+    )
+    experiment2.run()
+    experiment2.plot_theta1_heatmaps()
+
+
+if __name__ == "__main__":
+    main()

### 2022-05-03 experiments/data_generation.py

@@ -1,0 +1,169 @@
+from dataclasses import dataclass
+from typing import Tuple, Type, TypeVar
+
+import numpy as np
+from numpy import typing as npt
+
+
+def generate_constant_correlation_matrix(p: int, rho: float) -> np.ndarray:
+    """Generate a correlation/covariance matrix with unit entries along the main diagonal, and `rho` elsewhere.
+
+    Note that matrix may not be SPD if `rho < 0`"""
+    K = np.zeros((p, p))
+    for idx in range(p):
+        for jdx in range(p):
+            if idx == jdx:
+                K[idx, jdx] = 1
+            else:
+                K[idx, jdx] = rho
+    return K
+
+def generate_imbalanced_X(
+    size0: int, size1: int, mean0: npt.ArrayLike, mean1: npt.ArrayLike, K: np.ndarray
+) -> np.ndarray:
+    """Generate `size0` observations with mean `mean0` and covariance `K`, and `size1` observations with mean `mean1`
+    and covariance `K`.
+
+    Observations from group 0 constitute the first `size0` rows, and observations from group 1 the last `size1` rows
+    """
+    p = K.shape[0]
+
+    mu1 = np.zeros(p) + np.asarray(mean0)
+    mu2 = np.zeros(p) + np.asarray(mean1)
+
+    X0 = np.random.multivariate_normal(mean=mu1, cov=K, size=size0)
+    X1 = np.random.multivariate_normal(mean=mu2, cov=K, size=size1)
+    return np.vstack([X0, X1])
+
+
+_ExperimentDataModel = TypeVar("_ExperimentDataModel", bound="ExperimentDataModel")
+
+@dataclass
+class ExperimentDataModel:
+    """Handles data generation and organization for the experiments.
+    
+    THIS IS HIGHLY SPECIFIC TO THE TYPES OF EXPERIMENTS I'M RUNNING.
+    IT IS ONLY HERE FOR CONVENIENCE, AND NOT GENERALIZABLE
+
+    Attributes
+    ----------
+    size0: int
+        Number of group 0 (control) observations
+    size1: int
+        Number of group 1 (treatment) observations
+    p: int
+        Number of features; the feature matrix `X` has shape `(size0 + size1, p)`
+    theta0: float
+        Location parameter to a normal distribution, from which group 0 feature means are drawn
+    theta1: float
+        Location parameter to a normal distribution, from which group 1 feature means are drawn
+    rho: float
+        Non-diagonal entries of the covariance matrix `K` are filled with `rho`.
[…]

### 2022-05-03 experiments/experiments.py

@@ -1,0 +1,376 @@
+from collections import defaultdict
+from dataclasses import dataclass, field
+import os
+import pathlib
+import shutil
+import sys
+from typing import Any, Callable, Dict, Iterator, List, Tuple
+
+from matplotlib import pyplot as plt
+import numpy as np
+import pandas as pd
+import tqdm
+import seaborn as sns
+
+from experiments.data_generation import ExperimentDataModel
+from matching import distance
+from matching import preprocessing
+from matching.graph import MatchingGraph
+
+"""Base directory to save output to"""
+OUTPUT_BASE_DIR = (pathlib.Path(__file__).parent / "output").resolve()
+
+
+def clean_output_base_dir() -> None:
+    """Clean (i.e., delete) the contents of the output base directory"""
+    if OUTPUT_BASE_DIR.exists():
+        shutil.rmtree(OUTPUT_BASE_DIR)
+
+
+@dataclass
+class ExperimentDataLogger:
+    """Handles creating directories to save experiment data to
+
+    Attributes
+    ----------
+
+    experiment_name: str
+        Name of the experiment, which will be the subdirectory of `OUTPUT_BASE_DIR` we save to
+    """
+
+    experiment_name: str
+
+    def __post_init__(self) -> None:
+        """Create the directories, if they don't exist already"""
+        self.output_dir.mkdir(parents=True, exist_ok=True)
+        self.csv_dir.mkdir(parents=True, exist_ok=True)
+        self.plot_dir.mkdir(parents=True, exist_ok=True)
+
+    @property
+    def output_dir(self) -> pathlib.Path:
+        """Base directory for output"""
+        return OUTPUT_BASE_DIR / self.experiment_name
+
+    @property
+    def csv_dir(self) -> pathlib.Path:
+        """Directory to write CSVs to"""
+        return self.output_dir / "csvs"
+
+    @property
+    def plot_dir(self) -> pathlib.Path:
+        """Directory to write plots to"""
+        return self.output_dir / "plots"
+
+    def save_df(self, df: pd.DataFrame, filename: str) -> None:
+        """Save `df` to a CSV file. Just a wrapper around `to_csv` method, with some additional defaults
+
+        Parameters
+        ----------
+        df: pd.DataFrame
+            to save
+        filename: str
[…]

### 2022-05-03 experiments/__main__.py
finalized experiments

@@ -3,12 +3,10 @@ import numpy as np
 
 
 def main():
-    experiments.clean_output_base_dir()
-
     SIZE1 = 250
     SIZE0 = 750
     P = 5
-    NITER_PER_DGM = 50
+    NITER_PER_DGM = 30
 
     CALIPERS = np.round(np.linspace(0.05, 0.5, 10), 2)
 

### 2022-05-03 experiments/experiments.py

@@ -1,10 +1,7 @@
-from collections import defaultdict
 from dataclasses import dataclass, field
-import os
 import pathlib
 import shutil
-import sys
-from typing import Any, Callable, Dict, Iterator, List, Tuple
+from typing import Any, Callable, Iterator, Tuple
 
 from matplotlib import pyplot as plt
 import numpy as np
@@ -14,7 +11,6 @@ import seaborn as sns
 
 from experiments.data_generation import ExperimentDataModel
 from matching import distance
-from matching import preprocessing
 from matching.graph import MatchingGraph
 
 """Base directory to save output to"""
@@ -42,6 +38,8 @@ class ExperimentDataLogger:
 
     def __post_init__(self) -> None:
         """Create the directories, if they don't exist already"""
+        if self.output_dir.exists():
+            shutil.rmtree(self.output_dir)
         self.output_dir.mkdir(parents=True, exist_ok=True)
         self.csv_dir.mkdir(parents=True, exist_ok=True)
         self.plot_dir.mkdir(parents=True, exist_ok=True)
@@ -264,113 +262,3 @@ class CaliperExperiment:
         self.log.save_plot(_plot_heatmap_zero_arg(self.rho_caliper_mean_abs_smd_matrix, fmt="f"), "rho_caliper_mean_abs_smd", figsize=figsize)
         self.log.save_plot(_plot_heatmap_zero_arg(self.rho_caliper_max_abs_smd_matrix, fmt="f"), "rho_calipermax_abs_smd", figsize=figsize)
         # fmt: on
-
-
-def experiment1(n_treat: int, n_control: int, p: int) -> None:
-    """Experiment 1 investigates the effect of increasing the correlation between features `X`
-    on choosing an "optimal" caliper for propensity scores"""
-    edl = ExperimentDataLogger("experiment1")
-
-    N_RHO = 17
-    N_CALIPER = 20
-
-    rhos = np.round(np.linspace(0, 0.8, N_RHO), 2)
-    calipers = np.round(np.linspace(0.05, 1, N_CALIPER), 2)
-
-    N_MATCH_MATRIX = np.zeros((N_RHO, N_CALIPER), dtype=int)
-    MEAN_ABS_SMD_MATRIX = np.zeros((N_RHO, N_CALIPER))
-    MAX_ABS_SMD_MATRIX = np.zeros((N_RHO, N_CALIPER))
-
-    for rdx, rho in enumerate(tqdm.tqdm(rhos)):
[…]

### 2022-05-03 experiments/__main__.py
minor tweaks

@@ -3,11 +3,13 @@ import numpy as np
 
 
 def main():
+    # Set constant hyperparameters
     SIZE1 = 250
     SIZE0 = 750
     P = 5
     NITER_PER_DGM = 30
 
+    # We'll use the same calipers for each experiment too
     CALIPERS = np.round(np.linspace(0.05, 0.5, 10), 2)
 
     # Setting up experiment 1: caliper vs correlations

### 2022-05-03 experiments/experiments.py

@@ -130,7 +130,7 @@ class CaliperExperiment:
     theta1s: np.ndarray = field(default_factory=lambda: np.array([1], dtype=np.float_))
     niter_per_dgm: int = field(default=1)
 
-    log: ExperimentDataLogger = field(init=False)
+    logger: ExperimentDataLogger = field(init=False)
     rho_caliper_n_match_matrix: np.ndarray = field(init=False)
     rho_caliper_mean_abs_smd_matrix: np.ndarray = field(init=False)
     rho_caliper_max_abs_smd_matrix: np.ndarray = field(init=False)
@@ -140,7 +140,7 @@ class CaliperExperiment:
 
     def __post_init__(self) -> None:
         """Initializes matrices"""
-        self.log = ExperimentDataLogger(self.experiment_name)
+        self.logger = ExperimentDataLogger(self.experiment_name)
         self.rho_caliper_n_match_matrix = np.zeros([self.nrho, self.ncaliper], dtype=np.int_)
         self.rho_caliper_mean_abs_smd_matrix = np.zeros([self.nrho, self.ncaliper], dtype=np.float_)
         self.rho_caliper_max_abs_smd_matrix = np.zeros([self.nrho, self.ncaliper], dtype=np.float_)
@@ -203,7 +203,7 @@ class CaliperExperiment:
                     # Sometimes a matching cannot be found; we should increase sample size
                     # Print some info about parameters
                     print(f"{self.experiment_name} FAILED: {theta1=}, {rho=}, {caliper=}")
-                    raise
+                    continue
 
                 n_match = mg.graph_data.nobs
                 standardized_mean_differences = np.asarray(graph_subdf[True].values)
@@ -234,9 +234,9 @@ class CaliperExperiment:
 
         figsize = (self.ntheta1 * 2, self.ncaliper * 2)
         # fmt: off
-        self.log.save_plot(_plot_heatmap_zero_arg(self.theta1_caliper_n_match_matrix, fmt="d"), "theta1_caliper_n_match", figsize=figsize)
-        self.log.save_plot(_plot_heatmap_zero_arg(self.theta1_caliper_mean_abs_smd_matrix, fmt="f"), "theta1_caliper_mean_abs_smd", figsize=figsize)
[…]

### 2022-05-05 experiments/__main__.py
updated experiments

@@ -7,7 +7,7 @@ def main():
     SIZE1 = 250
     SIZE0 = 750
     P = 5
-    NITER_PER_DGM = 30
+    NITER_PER_DGM = 100
 
     # We'll use the same calipers for each experiment too
     CALIPERS = np.round(np.linspace(0.05, 0.5, 10), 2)
@@ -15,7 +15,7 @@ def main():
     # Setting up experiment 1: caliper vs correlations
     RHOS = np.round(np.linspace(0, 0.8, 9), 2)
     experiment1 = experiments.CaliperExperiment(
-        experiment_name="caliper_vs_correlation_big",
+        experiment_name="caliper_vs_correlation_big2",
         size0=SIZE0,
         size1=SIZE1,
         p=P,
@@ -29,7 +29,7 @@ def main():
     # Setting up experiment 2: caliper vs imbalance
     THETA1S = np.round(np.linspace(0, 1, 11), 2)
     experiment2 = experiments.CaliperExperiment(
-        experiment_name="caliper_vs_imbalance_big",
+        experiment_name="caliper_vs_imbalance_big2",
         size0=SIZE0,
         size1=SIZE1,
         p=P,

### 2022-05-05 experiments/experiments.py

@@ -131,20 +131,20 @@ class CaliperExperiment:
     niter_per_dgm: int = field(default=1)
 
     logger: ExperimentDataLogger = field(init=False)
-    rho_caliper_n_match_matrix: np.ndarray = field(init=False)
+    rho_caliper_prop1_match_matrix: np.ndarray = field(init=False)
     rho_caliper_mean_abs_smd_matrix: np.ndarray = field(init=False)
     rho_caliper_max_abs_smd_matrix: np.ndarray = field(init=False)
-    theta1_caliper_n_match_matrix: np.ndarray = field(init=False)
+    theta1_caliper_prop1_match_matrix: np.ndarray = field(init=False)
     theta1_caliper_mean_abs_smd_matrix: np.ndarray = field(init=False)
     theta1_caliper_max_abs_smd_matrix: np.ndarray = field(init=False)
 
     def __post_init__(self) -> None:
         """Initializes matrices"""
         self.logger = ExperimentDataLogger(self.experiment_name)
-        self.rho_caliper_n_match_matrix = np.zeros([self.nrho, self.ncaliper], dtype=np.int_)
+        self.rho_caliper_prop1_match_matrix = np.zeros([self.nrho, self.ncaliper], dtype=np.float_)
         self.rho_caliper_mean_abs_smd_matrix = np.zeros([self.nrho, self.ncaliper], dtype=np.float_)
         self.rho_caliper_max_abs_smd_matrix = np.zeros([self.nrho, self.ncaliper], dtype=np.float_)
-        self.theta1_caliper_n_match_matrix = np.zeros([self.ntheta1, self.ncaliper], dtype=np.int_)
+        self.theta1_caliper_prop1_match_matrix = np.zeros([self.ntheta1, self.ncaliper], dtype=np.float_)
         self.theta1_caliper_mean_abs_smd_matrix = np.zeros([self.ntheta1, self.ncaliper], dtype=np.float_)
         self.theta1_caliper_max_abs_smd_matrix = np.zeros([self.ntheta1, self.ncaliper], dtype=np.float_)
 
@@ -167,15 +167,17 @@ class CaliperExperiment:
         self, tdx: int, cdx: int, n_match: int, mean_abs_smd: float, max_abs_smd: float
     ) -> None:
[…]

### 2022-05-05 paper/main.tex
rough draft of paper finished

@@
 
+\input{preamble.tex}
+
 \title{Development of a Python Package for Matching Observational Data}
+\subtitle{Advisors: Jingshu Wang, Sanjay Krishnan}
 \author{Jack Potrykus}
@@
 
-\input{preamble.tex}
-
 \begin{document}
 
-\maketitle{}
+% Insert some text making clear that this is an MS thesis
+\begin{center}{\large \bfseries Master’s Paper for the University of Chicago Department of Statistics}\end{center}
+
+% Relax `\maketitle` to allow text on same page. https://tex.stackexchange.com/questions/86249/maketitle-text-before-title
+{\let\newpage\relax\maketitle}
 
@@
 
+% Insert a spot for the advisor's signature
+\vfill{\null}
+Approved by: \hrulefill~Date: \hrulefill
+
 \newpage{}
@@
 Observational studies are of ever-increasing importance as the numerous amount of data collected each year continues to grow.
-However, of observational studies are at a major disadvantage compared to blocking-based study design or randomized control trials (RCTs), because the researchers do not have control over the distribution of confounding variables, known or unknown, in the data itself.
+However, of observational studies are at a major disadvantage compared to blocking-based study design or randomized control trials, because the researchers do not have control over the distribution of confounding variables, known or unknown, in the data itself.
 As such, researchers must take preliminary steps to alter the data in some way in order to minimize the effects of selection bias, which results from heterogeneous distributions present in the observational data.
@@
 Within the field of causal inference, the term \emph{matching} is often associated with observational studies with a binary treatment indicator.
-Using a similar framework to that of \textcite{iacus_multivariate_2011}, consider $\mat{X}_T \in \reals^{n \times k}$ and $\mat{X}_C \in \reals^{m \times k}$, matrices of $k$ features for $n$ treatment-group observations and $m$ control-group observations, $n \leq m$.
+Using a similar framework to that of \textcite{iacus_multivariate_2011}, consider $\mat{X}_T \in \reals^{n \times p}$ and $\mat{X}_C \in \reals^{m \times p}$, matrices of $p$ features for $n$ treatment-group observations and $m$ control-group observations, $n \leq m$.
 We seek to produce a matching $\set{M}$ such that each observation in $\mat{X}_T$ is paired with one or more nearby observations from $\mat{X}_C$, according to a distance metric and a matching algorithm.
@@
 It achieves this by allowing distance measures and matching algorithms to be independently and iteratively applied, so that even incredibly bespoke matching procedures may be implemented succinctly in code.
-It is also the first Python library for matching observational data which is ``graph-centric'': the key data structure used is a bipartite graph, whose nodes represent observation IDs, and edge weights the distance measure between them.
+It is also the first Python library for matching observational data which is ``graph-centric'': the key data structure used is a bipartite graph, whose nodes represent observations, and edge weights the distance measure between them.
 This enables further exploration into new matching procedures and balance diagnostics using graph metrics and algorithms from computer science.
@@
 
-In the context of this paper, distance measures are functions $d: \reals^k \times \reals^k \mapsto \reals^+$ which produce a non-negative distance, or ``cost of matching'', between any vectors $\vec{x}_T, \vec{x}_C$ from the rows of $\mat{X}_T \in \reals^{n \times k}$ and $\mat{X}_C \in \reals^{m \times k}$, to judge the match quality.
+In the context of this paper, distance measures are functions $d: \reals^p \times \reals^p \mapsto \reals^+$ which produce a non-negative distance, or ``cost of matching'', between any vectors $\vec{x}_T, \vec{x}_C$ from the rows of $\mat{X}_T \in \reals^{n \times p}$ and $\mat{X}_C \in \reals^{m \times p}$, to judge the match quality.
 Distance measures are often combined with some preprocessing function $f(\mat{X})$, usually in the form of a dimension reduction or discretization of the data.
@@
 \end{equation}
-where $\mathcal{D}_d : \reals^{n \times k} \times \reals^{m \times k} \mapsto \reals^{n \times m}$ is a ``vectorized'' version of $d$, producing a matrix $\mat{D}$ whose $(i, j)$th element is the distance between observation $i$ from $\mat{X}_T$ and observation $j$ from $\mat{X}_C$.
-
-In the simplest case, $d$ is some norm (perhaps Euclidean) calculated on the difference between feature vectors, and $f$ is the identity.
-With this simple framework in mind, we are now ready to explore extensions of the simple case, and examine their proposed benefits.
+where $\mathcal{D}_d : \reals^{n \times p} \times \reals^{m \times p} \mapsto \reals^{n \times m}$ is a ``vectorized'' version of $d$, producing a matrix $\mat{D}$ whose $(i, j)$th element is the distance between observation $i$ from $\mat{X}_T$ and observation $j$ from $\mat{X}_C$.
 
@@
 However, this paper uses the term ``balancing score'' specifically with regards to 1-dimensional reductions of the data.
-That is to say, that we will consider balancing scores as a preprocessing function $b: \reals^{(n + m) \times k} \mapsto \reals^{n + m}$.}
+That is to say, that we will consider balancing scores as a preprocessing function $b: \reals^{(n + m) \times p} \mapsto \reals^{n + m}$.}
 \parencite{rosenbaum_central_1983} are a versatile class of functions $b(X)$ satisfying the property
@@
 \end{equation}
-where the matrix $\mat{X} \in \reals^{(n + m) \times k} = \mat{X}_T \cup \mat{X}_C$, and the vector $\vec{z} \in \reals^{n + m}$ contains binary treatment group assignments.
-In words, given the balancing score $b(\mat{X})$, the distribution of the feature vector $\mat{X}$ is independent of treatment assignment $\vec{z}$.
-These methods each set $d$ equal to the $L^1$ norm, or absolute, distance by convention.
+where the matrix $\mat{X} \in \reals^{(n + m) \times p} = \mat{X}_T \cup \mat{X}_C$, and the vector $\vec{z} \in \reals^{n + m}$ contains binary treatment group assignments.
+In words, given the balancing scores $b(\mat{X})$, the distribution of the feature matrix $\mat{X}$ is independent of treatment assignment $\vec{z}$.
+These methods each set $d$ equal to the $L^1$ norm, or absolute distance, by convention.
 In the notation of~\eqref{eqn:abstractdistance}, then, balancing scores compose the class of distance measures of the form
@@
 
-TODO DISCUSS STRONG IGNORABILITY, PROPERTIES, WHEN IS ATE ESTIMATE UNBIASED
-+ TODO MAKE CLEAR IT IS REPLICATING A RANDOMIZED TRIAL
+Balancing scores afford observational data some of the desirable properties of randomized trials.
+Namely, in a randomized trial, $\mat{X}$ should contain all features used in determining treatment assignment, affording conditional independence of outcomes from treatment:
+\begin{equation}
+  \vec{y}_0, \vec{y}_1 \perp \vec{z} | \mat{X}, \label{eqn:defstrongignorable}
+\end{equation}
+where $\vec{y}_0$ and $\vec{y}_1$ are vectors of outcomes, for which the $i$th element is the outcome of observation $i$ when assigned to group 0 (control) or 1 (treatment); note that in practice we almost always only observe $y_{0,i}$ \emph{or} $y_{1,i}$ for each $i$.
+If the statement in~\eqref{eqn:defstrongignorable} holds, then the treatment assignment is \emph{strongly ignorable}.
+\textcite{rosenbaum_central_1983} then prove that if~\eqref{eqn:defstrongignorable} holds (that is, treatment $\vec{z}$ is strongly ignorable given features $\mat{X}$), then treatment assignment is \emph{also} strongly ignorable given any balancing score $b(X)$!
+Conditioning on balancing scores, we gain independence of treatment from features, as well as the strong ignorability of treatment; these useful properties form the key motivation for the use of balancing scores in causal inference.
 
@@
 
-\paragraph{Joint Use of Balancing Scores.} \textcite{leacy_joint_2014} further explores of calculating \emph{both} propensity and prognostic scores, stratifying each into a $k \times k$ grid, and considering observations within each of the $k^2$ grid squares as a \emph{matched strata}.
+\paragraph{Joint Use of Balancing Scores.} \textcite{leacy_joint_2014} further explores of calculating \emph{both} propensity and prognostic scores, stratifying each into a $p \times p$ grid, and considering observations within each of the $p^2$ grid squares as a \emph{matched strata}.
 However, they found that matching on prognostic scores alone performed the best, while matching on propensity scores and Mahalanobis distance (see \S\ref{subsubsec:normdist}) also performed admirably.
-The $k \times k$ strata performed better than stratification of either score alone, but could not beat matching on either un-discretized score alone.
+The $p \times p$ strata performed better than stratification of either score alone, but could not beat matching on either un-discretized score alone.
 They also found that the percent bias of estimates created using this matching method increased with the non-linearity in the outcomes model.
@@
 They thus regard balancing score methods as having a ``lower standard''.
-Second, they observe the ``paradox of Propensity Score Matching'': consider the case where $\mat{X} \perp \vec{z}$, and so all predicted scores are the same constant: $\hat{\vec{z}}$.
-Then, all potential matches are indecipherable, and the individual matchings may be complete nonsensical.
+Second, they observe the ``Paradox of Propensity Score Matching'': consider the case where $\mat{X} \perp \vec{z}$, and so all predicted scores are the same constant: $\hat{\vec{z}}$.
+Then, all potential matches are indecipherable, and the individual matchings may be complete nonsensical, and in fact may \emph{increase} the imbalance of features within the matched subset.
 The authors nonetheless concede that propensity scores (and other balancing scores, by extension) do indeed have nice theoretical properties, and may serve as a useful diagnostic; they contend only that they should not be used for matching.
-
-TODO VERIFY
 
@@
 These methods take a vector of tuning parameters $\vec{\uppi}$, such as a vector of calipers.
-These methods afford the benefit that the number of matches is a function of the tuning parameters alone, and given monotonicity of the maximum distance function $\gamma_{D_{d}}(\vec{\uppi})$, can be each adjusted independently other others: if $\gamma$ is vector-valued, returning a vector in $\reals^k$, the distance between each feature vector can be independently compared in a sequence of inequalities.
+These methods afford the benefit that the number of matches is a function of the tuning parameters alone, and given monotonicity of the maximum distance function $\gamma_{\mathcal{D}_{d}}(\vec{\uppi})$, can be each adjusted independently other others: if $\gamma$ is vector-valued, the element-wise differences of feature vectors can be independently compared in a system of inequalities.
 CEM is an example of an MIB distance measure, but this class is clearly much more flexible, allowing for precise adjustments.
@@
 These two algorithms are unified by their tendency to sometimes ``drop'' uninformative features from the matching process, in order to ensure that the most important features are matched on, and potentially increase the number of potential matches.
-
-TODO FINISH
+The chief difference can be summarized as follows: DAME choosing a subset of features to match on by minimizing predictive error, whereas FLAME maximizes a match quality statistic that the researcher can tune via a hyperparameter; this statistic is a function of predictive error and a \emph{balance factor}, which represents the relative size of the strata of observations matched on this subset on features \parencite{gupta_dame-flame_2021}.
 
@@
 
+\subsubsection{Iterative Distance Calculations}
+
+Often, researchers will employ an ``iterative'' distance measure.
+They may, for example, first calculate propensity scores $\widehat{\vec{z}}$, filter potential matches using a caliper $c$, and then match on Mahalanobis distance \emph{within} clusters; or, in the absence of a caliper, within strata of $\widehat{\vec{z}}$.
+This idea was explored by \textcite{rosenbaum_constructing_1985}, who found it to perform better than matching on either propensity score or Mahalanobis distance alone.
+
+\textcite{baltar_mahalanobis_2014} explored the efficacy of this iterative method in practice, relative to propensity score and Mahalanobis distance alone.
+In the paper, they compare seek to estimate the effect of Brazilian towns' social agenda on public health.
+They found both the propensity score alone and iterative method to perform admirably, whereas Mahalanobis distance performed poorly; they do not take this as an indictment of Mahalanobis distance matching, and pose for further research to identify conditions under which Mahalnobis distance may perform \emph{better}.
+The iterative method achieved the greatest reduction in bias for all features, except the propensity score, for which propensity score distance alone (somewhat obviously) saw the greatest bias reduction.
+However, as \textcite{austin_optimal_2011} and \textcite{greifer_assessing_2022} note, balance of the scores does not imply balance of the features, and balancing score balance metrics should only be used as a supplementary diagnostic; the focus of balance assessment should be balance of the features.
+
 \subsection{Matching Algorithms}
@@
 Once the distance measure has been agreed upon, the researcher must then decide \emph{how} to match.
-A computer scientist could describe this problem as finding the minimum weight full matching of a \emph{bipartite graph}, whose nodes represent observation IDs, and whose edge weights represent the distance between them.
+A computer scientist could describe this problem as finding the minimum weight full matching of a \emph{bipartite graph}, whose nodes represent observations, and whose edge weights represent the distance between them.
 Here, the term ``bipartite'' means that all edges in the graph map $T \leftrightarrow C$; within each disjoint subset, no two nodes share an edge.
@@
 \textcite{rosenbaum_optimal_1989} argues for optimal matching according to the Hungarian algorithm, also known as the Kuhn-Munkres or simply Munkres algorithm \parencite{munkres_algorithms_1957}.
-We will denote a matching a set $\set{M}$ of tuples $(i, j)$, with $i$ representing a treatment-group observation ID, and $j$ a control-group observation ID.
-The Hungarian algorithm solves the \emph{linear assignment problem},which in this setting may be expressed as
+We will denote a matching a set $\set{M}$ of tuples $(i, j)$, with $i$ representing a treatment-group observations, and $j$ a control-group observations.
+The Hungarian algorithm solves the \emph{linear assignment problem}, which in this setting may be expressed as
 \begin{equation}
@@
 One may also plot the improvement in standardized mean differences between the two groups, referred to as a Love plot \parencite{greifer_assessing_2022}.
-One could also plot the bipartite graph to understand where clusters are appearing, or how large groups of ``similar'' observations are, which observations are most ``dissimilar'' from the rest.
+One could also plot the bipartite graph to understand where clusters are appearing, or how large groups of ``similar'' observations are, or which observations are most ``dissimilar'' from the rest.
 
@@
   \item ``depth over breadth''.
-The package should only implement matching procedures, leaving model-building and effect estimation to the user.
-Narrowing in on the matching procedure alone also helps with the first three goals.
+The package should only implement matching procedures, leaving model-building and effect estimation to the user; narrowing in on the matching procedure alone also helps to achieve the first three goals.
 \end{itemize}
@@
   \center{%
-    \includegraphics[width=\textwidth]{../docs/mermaid/img/relationships.png}%
+    \includegraphics[width=0.8\textwidth]{../docs/mermaid/img/relationships.png}%
   }
@@
 The \texttt{MatchingGraph} is minimally initialized with an array-like (e.g., a \texttt{pandas.DataFrame}, or a \texttt{numpy.ndarray}) of features \texttt{X} and an array-like of binary treatment assignments \texttt{z}.
-The user may first wish to preprocess their data (i.e.
+The user may first wish to preprocess their data (e.g.
 calculating propensity scores) using \texttt{matching.preprocessing}, but this is not necessary.
@@
 The user may supply a \texttt{min\_distance} and/or a \texttt{max\_distance} for an edge to be allowed; in graph terms, this can be considered the edges' \emph{capacity}.
+For the \texttt{PropensityScore} distance metric, \texttt{max\_distance} is instead set by specifying a \texttt{caliper}; the \texttt{max\_distance} is later inferred, once the scores are calculated.
 \begin{minted}[linenos, fontsize=\small]{python3}
 import numpy as np
-from matching.distance import L1Norm
+from matching.distance import PropensityScore
 
-# Calculating the caliper width. Recall that X is an array of scores
-caliper = 0.05
-caliper_width = caliper * np.std(X)
-mg.set_edges(distance=L1Norm(max_distance=caliper_width))
+# Will calculate propensity score distance automatically
+mg.set_edges(distance=PropensityScore(caliper=0.05))
 \end{minted}
@@
 
-There are also methods to \texttt{filter\_nodes} by label or order, as well as to \texttt{filter\_subgraphs}, which filters the disjoint subgraphs of the graph by a variety of metrics\footnote{\texttt{filter\_subgraphs} is particularly relevant for strata-based matching methods like CEM, which often filter strata based on the total number of observations (from each group), the ratio of treatment to control}; they will not be discussed here, but readers are referred to \texttt{matching} documentation hosted on the GitHub\footnote{\url{https://github.com/jackpotrykus/propensity-score-matching-thesis}}.
+There are also methods to \texttt{filter\_nodes} by label or order, as well as to \texttt{filter\_subgraphs}, which filters the disjoint subgraphs of the graph by a variety of metrics\footnote{\texttt{filter\_subgraphs} is particularly relevant for strata-based matching methods like CEM, which often filter strata based on the total number of observations (from each group), the ratio of treatment to control.}; they will not be discussed here, but readers are referred to \texttt{matching} documentation hosted on the GitHub\footnote{\url{https://github.com/jackpotrykus/propensity-score-matching-thesis}}.
 
-\paragraph{Matching.} Once the user has set the edges of the graph and completed their filtering, they may wish to conduct a $1:k$ matching between treatment and control, for some integer $k \geq 1$\footnote{This is not always necessary, particularly in the case of CEM}.
+\paragraph{Matching.} Once the user has set the edges of the graph and completed their filtering, they may wish to conduct a $1:k$ matching between treatment and control, for some integer $k \geq 1$\footnote{This is not always necessary, particularly in the case of CEM.}
 The user does this via a call to the \texttt{match} method.
@@
 \begin{minted}[linenos, fontsize=\small]{python3}
-# Now we should match the graph
-# By default: n_match=1, min_match=1, replace=False
+# Continue from the previous excerpt. Now we want to conduct 1:k matching
+# NOTE: By default: n_match=1, min_match=1, replace=False
 mg.match(n_match=2, min_match=2, method="optimal")
@@
 
-\paragraph{Balance Assessment.} The user can then compare the balance of the \texttt{input\_data} and the \texttt{match\_data}\footnote{In the case of Exact matching edges, the \texttt{match\_data} should consist of all subgraphs of order at least 2}.
+\paragraph{Balance Assessment.} The user can then compare the balance of the \texttt{input\_data} and the \texttt{match\_data}\footnote{In the case of Exact matching edges, the \texttt{match\_data} should consist of all subgraphs of order at least 2.}.
 Note that \texttt{match\_data} is a subset of \texttt{input\_data}, plus a new column, \texttt{"match\_group"}.
@@
 
+\paragraph{Drawing the Graph.} After edges have been set using \texttt{set\_edges}, users may draw the graph at any point in the matching process.
+\texttt{networkx} provides a myriad of functions to draw graphs, but the \texttt{MatchingGraph} class has two convenient graph-drawing methods built-in.
+These are the methods \texttt{draw} and \texttt{draw\_bipartite}.
+The main difference between these two is that \texttt{draw} will draw the graph such that edge length is proportional to the distance between observations, whereas \texttt{draw\_bipartite} draws the graph in the \emph{bipartite layout}, with all treatment group nodes on one side, and all control group nodes on the other.
+Example output of each of these is displayed in Figure~\ref{fig:draw_examples}.
+\begin{figure}[h!]
+    \centering
+    \begin{subfigure}[t]{0.5\textwidth}
+        \centering
+        \includegraphics[width=0.95\textwidth]{./img/draw_examples/draw.png}
+        \caption{Graph produced by \texttt{mg.draw()}. Edge length corresponds to the distance between observations.}
+    \end{subfigure}%
+    ~
+    \begin{subfigure}[t]{0.5\textwidth}
+        \centering
+        \includegraphics[width=0.95\textwidth]{./img/draw_examples/draw_bipartite.png}
+        \caption{Graph produced by \texttt{mg.draw\_bipartite()}. The bipartite structure of the graph is clear, but edge lengths are arbitrary.}
+    \end{subfigure}
+    \caption{Two visualizations of the \emph{same} graph.
+This is a graph that resulted from $1:3$ optimal matching.
+Each graph automatically applies the color-coding of treatment nodes in {\color{red}red} and control nodes in {\color{blue} blue}.
+Observation IDs/labels may be displayed in either plot by passing \texttt{with\_labels=True}.}
+    \label{fig:draw_examples}
+\end{figure}
+
 \subsection{Comparison to Other Causal Inference Packages}
@@
 However, since these each seek to replicate CEM, and since these are the \emph{only} matching methods provided by \texttt{dame-flame}, this means that \texttt{dame-flame} can only be used with discrete datasets; for users unwilling to coarsen/discretize their data, or users seeking a matching procedure other than DAME and FLAME, \texttt{dame-flame} is not an option.
-\texttt{matching} does not currently support these algorithms, but they are ``in-the-works''; it can certainly be implemented as a new \texttt{matching.distance.\_Distance} subclass.
+\texttt{matching} does not currently support these algorithms, but they are ``in-the-works''; it can certainly be implemented as a new \texttt{matching.distance.\_Distance} subclass, and their impressive performance merits their inclusion.
 
 \section{Experiments}
+\label{sec:experiments}
+
+To demonstrate the use and capabilities of the \texttt{matching} package, I have also conducted the following numerical experiments, which offer practical lessons on parameter-tuning when propensity score matching.
 
@@
 
-Papers with data generation:
+The data were generated according to the following hyperparameters. This follows similar frameworks in the literature, such as those of \textcite{austin_optimal_2011}.
 \begin{itemize}
-  \item \cite{austin_optimal_2011}
-  \item \cite{stuart_prognostic_2013}
+  \item \texttt{size0} ($m$): number of control observations to generate.
+  \item \texttt{size1} ($n$): number of treatment observations to generate.
+  \item \texttt{p} ($p$): number of features to generate.
+  \item \texttt{theta0} ($\theta_0$): the feature means for the control group are distributed as follows:
+    \begin{equation}
+      \texttt{mean0} = \vec{\upmu}_0 \in \reals^p \sim \mathcal{N}(\theta_0 \cdot \vec{1}_p, 0.25 \cdot \mat{I}_p).
+    \end{equation}
+  \item \texttt{theta1} ($\theta_1$): the feature means for the treatment group are distributed as follows:
+    \begin{equation}
+      \texttt{mean1} = \vec{\upmu}_1 \in \reals^p \sim \mathcal{N}(\theta_1 \cdot \vec{1}_p, 0.25 \cdot \mat{I}_p).
+    \end{equation}
+  \item \texttt{rho} ($\rho$): the feature covariance matrix (for both groups) is set as follows:
+    \begin{equation}
+      \texttt{K[i][j]} = \mat{\Sigma}_{i j} = \begin{cases}
+        1    & \text{ if } i = j \\
+        \rho & \text{ otherwise.}
+      \end{cases}
+      \label{eqn:dgm_K}
+    \end{equation}
 \end{itemize}
+We then draw features
+\begin{equation}
+  \{ \vec{x}_{C_j} \}_{j=1}^{m} \sim \mathcal N(\vec{\upmu}_0, \mat{\Sigma}), \text{ and }
+  \{ \vec{x}_{T_i} \}_{i=1}^{n} \sim \mathcal N(\vec{\upmu}_1, \mat{\Sigma}).
+\end{equation}
+The treatment assignments \texttt{z} were deterministically set according to which mean vector the rows were drawn from.
+For each set of hyperparameters (as listed above), 30 datasets were generated, and summary statistics to judge the ``quality of matching'' were averaged across these trials.
 
-\section{Results}
+The two experiments each test the robustness and suitability of varying calipers with regards to a strenuous data setting: one of increasing correlation between the features, and one of increasing imbalance of the features (between treatment and control groups).
 
-\section{Discussion}
+In each experiment, propensity scores were fit using standard Logistic Regression, and then optimal $1:1$ matching was conducted, meaning the maximum size of the matched subset would be 500 observations.
+Each experiment varies one hyperparameter against caliper values $c \in \{0.05, 0.10, 0.15, \ldots, 0.50\}$.
+To assess the balance of the matched subset, we record the mean \emph{absolute} standardized mean difference of the features in the matched subset; out of practical considerations, we also record the proportion of treatment observations matched $\in [0, 1]$.
+
+\subsection{Experiment 1: Caliper vs Imbalance.}
+This experiment seeks to investigate the practical truth of the ``Paradox of Propensity Score Matching'' \parencite{king_why_2019}.
+The Paradox suggests that matching quality could (not necessarily \emph{must}) deteriorate as the imbalance between the treatment and control groups decreases; this is a logical extrapolation from that earlier-noted fact that when $\mat{X} \perp \vec{z}$, the classifier will assign a constant propensity score to all observations.
+
+For this experiment, we hold constant \texttt{size0=750}, \texttt{size1=250}, \texttt{p=5}, \texttt{theta0=0}, \texttt{rho=0} for all simulated data.
+Then, we vary $\texttt{theta1} = \theta_1 \in \{0, 0.1, 0.2, \ldots, 1\}$.
+
+\begin{figure}[h!]
+  \centering
+  \begin{center}
+    \begin{subfigure}[t]{0.49\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{./img/output30/caliper_vs_imbalance_big/plots/theta1_caliper_prop1_match.png}
+        \caption{Proportion of treatment observations matched}
+        \label{subfig:exp1_nmatch}
+    \end{subfigure}
+    \begin{subfigure}[t]{0.49\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{./img/output30/caliper_vs_imbalance_big/plots/theta1_caliper_mean_abs_smd.png}
+        \caption{Mean absolute standardized mean difference}
+        \label{subfig:exp1_mean_abs_smd}
+    \end{subfigure}
+    \caption{$\rho$ is varied from $0.0$ to $0.8$ in increments of $0.1$ \emph{down} the $y$-axis. $c$ is varied from $0.05$ to $0.5$ in increments of $0.05$ along the $x$-axis.}
+  \end{center}
+\end{figure}
+
+\paragraph{Analysis of Subfigure~\ref{subfig:exp1_nmatch}.} Of course, the proportion of treatment observations matched is monotonically increasing in $c$.
+This heatmap makes clear that when heterogeneity of the features is too large (even just a single standard deviation apart), the proportion of treatment observations matched may decrease substantially when using the same caliper $c$ (here, we see a decrease of 50\% for $c = 0.05$).
+\paragraph{Analysis of Subfigure~\ref{subfig:exp1_mean_abs_smd}.} First, we note that the mean absolute standardized mean difference is nearly constant down the leftmost column (with calipers of $c=0.05$); the cost of imbalance was primarily paid in the size of the matched subset.
+In fact, the test statistics increase very little with imbalance for most values of $c$; only for $c = 0.45$ and $c = 0.50$ is the increase in imbalance from $\theta_1 = 0$ to $\theta_1 = 1$ greater than $0.05$.
+This suggests some robustness of the caliper, for normally distributed means that differ by no more than a standard deviation of the data.
+Moreover, we do not observe the Paradox; the propensity score matching performs very well when $\theta_0 = \theta_1 = 0$.
+This does not disqualify the Paradox -- it could certainly happen -- it just was not present in this multivariate Normally distributed data.
+
+Finally, we should note that the value of $c$ minimizing the test statistic was not necessarily $0.05$ in all cases; in fact, almost all datasets saw equal or better balance achieved by $c = 0.10, c = 0.15,$ and $c = 0.20$.
+This underscores the importance of thorough parameter search, as well as balance assessment.
+
+\subsection{Experiment 2: Caliper vs Correlation.}
+This experiment seeks to assess the impact that varying correlation among the features has on selection of an optimal caliper.
+Specifically, it hypothesizes that the optimal choice of caliper will vary significantly as correlation changes; this would be in line with results of the aforementioned \textcite{austin_optimal_2011}.
+
+For this experiment, we hold constant \texttt{size0=750}, \texttt{size1=250}, \texttt{p=5}, \texttt{theta0=0}, \texttt{theta1=1} for all simulated data.
+Then, we vary $\texttt{rho} = \rho \in \{0, 0.1, 0.2, \ldots, 0.8\}$.
+
+\begin{figure}[h!]
+  \centering
+  \begin{center}
+    \begin{subfigure}[t]{0.49\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{./img/output30/caliper_vs_correlation_big/plots/rho_caliper_prop1_match.png}
+        \caption{Proportion of treatment observations matched}
+        \label{subfig:exp2_nmatch}
+    \end{subfigure}
+    \begin{subfigure}[t]{0.49\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{./img/output30/caliper_vs_correlation_big/plots/rho_caliper_mean_abs_smd.png}
+        \caption{Mean absolute standardized mean difference}
+        \label{subfig:exp2_mean_abs_smd}
+    \end{subfigure}
+    \caption{$\theta_1$ is varied from $0.0$ to $1.0$ in increments of $0.1$ \emph{down} the $y$-axis. $c$ is varied from $0.05$ to $0.5$ in increments of $0.05$ along the $x$-axis.}
+    \label{fig:exp2}
+  \end{center}
+\end{figure}
+\paragraph{Analysis of Subfigure~\ref{subfig:exp2_nmatch}.} There is not much discernable pattern in the proportion of observations matched with respect to $\rho$ until $\rho=0.8$, which performs very badly.
+Interestingly, moderate values of $\rho$ (e.g.
+$\rho=0.3$) afforded the highest proportion of treatment observations, in some cases much higher than in the case of independent features ($\rho = 0$).
+\paragraph{Analysis of Subfigure~\ref{subfig:exp2_mean_abs_smd}.} There is again not much pattern in the data with respect to $\rho$.
+This time, even $\rho=0.8$ performs adequately; it has the highest mean absolute standardized mean difference for most values of $c$, but it is generally not ``much worse'' than any of the other cases.
+In fact, we find that the test statistic is minimized at modest values of $\rho$ (e.g.
+$\rho=0.3$); this is no doubt correlated with the high-proportion of treatment group observations matched for those same values of $\rho$.
+
+This analysis suggests that propensity score matching was able to produce high-quality matches even when the data generating model consisted of highly correlated features.
+This is in line with results presented in other research, such as \textcite{austin_optimal_2011}.
 
 \section{Conclusion}
+
+\texttt{matching} was developed with the real-world researcher responsibilities of tuning their matching process's parameters and assessing its balance in mind.
+By allowing for constant access to visual and numerical diagnostics throughout the procedure, the user has instant access to diagnostics informing how they should tune their hyperparameters, such as the regularization of the propensity scoring model, or the caliper of matching.
+Perhaps the most common theme present throughout all numerical investigation and comparisons of matching methods is the extreme data-dependence of the match quality; by offering a simple-to-use, plug-and-play framework for specifying matching procedures, users are empowered to explore match quality in great detail.
+
+The graph-centric nature of \texttt{matching} also affords many advantages: the third-party dependency (\texttt{networkx}) which handles graph initialization and manipulation is highly memory-optimized.
+Since we only need to keep track of the \emph{distance between} treatment observations and control observations (and furthermore dropping those greater than a specified \texttt{max\_distance}), we can precisely represent the similarity between even large datasets, avoiding multiplicative growth of necessary space.
+
+Finally, it is important to remember that diagnostic checks must be chosen to suit the situation.
+In the case of the experiments conducted in \S\ref{sec:experiments}, mean absolute standardized mean difference was useful because we knew the true difference between $\theta_1$ and $\theta_0$; in practice, this quantity is unknown.
+As such, researchers should use robust methods of balance assessment; for example, checking for convergence of higher-order moments, or using density-based metrics such as those of \textcite{zhu_kernel-based_2018}.
+But even then, one cannot possibly hope to analyze the balance of the \emph{unobserved} confounding features; for these, we are only promised balance \emph{on average}.
+Future research in matching could certainly explore the integration with other causal inference methods, such as the difference-in-differences method or the usage of instrumental variables, which seek to account for the effects of unobserved features.
 
@@
 \addcontentsline{toc}{section}{Addenda}
+
+\begin{itemize}
+  \item Extensive documentation for the \texttt{matching} package can be found on the GitHub at this \href{https://github.com/jackpotrykus/propensity-score-matching-thesis/blob/d47cd58ae7c6cd3fc8a5afc940212e2674d6e7ba/docs/matching.pdf}{link}.
+This PDF documents every class, function, and method available to the user in great detail, and will serve as a very useful tool for anyone looking to learn how to use the package.
+\end{itemize}
 

### 2022-05-05 presentation/presentation.tex
presentation finished

\documentclass[11pt, compress]{beamer}

\usepackage[utf8]{inputenc}
\renewcommand\mathfamilydefault{\rmdefault}
\renewcommand{\familydefault}{\sfdefault}
\usepackage{sansmath}
\sansmath
\setlength\parindent{0pt}
\usepackage{microtype}
\usepackage{amsmath}
\usepackage{minted}
\usepackage{breakcites}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{bm}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{bbm}
\usepackage{fancyhdr}
\usepackage{hyperref}
\usepackage{parskip}
\usepackage{listings}
\usepackage{graphicx}
\usepackage[style=authoryear]{biblatex}
\usepackage{upgreek}
\addbibresource{../paper/bibs/Matching.bib}

\usemintedstyle{manni}

\hypersetup{
    breaklinks=true,
    colorlinks=true,
    citecolor=magenta,
    linkcolor=blue,
    % linkcolor=magenta,
    filecolor=magenta,      
    urlcolor=blue,
    % urlcolor=magenta,
}

% Shortcuts for math
\newcommand{\reals}{\ensuremath{\mathbbmss R}}
\renewcommand{\vec}[1]{\ensuremath{\boldsymbol{\mathsfbf{#1}}}}
\newcommand{\mat}[1]{\ensuremath{#1}}
\newcommand{\set}[1]{\ensuremath{\mathcal{#1}}}

\title{}
\date{\today}
\title{Development of a Python Package for Matching Observational Data}
\subtitle{Master's Thesis Presentation}
\author{Jack Potrykus}
\date{\today}
\institute{University of Chicago Department of Statistics}

% Section slides: https://tex.stackexchange.com/questions/178800/creating-sections-each-with-title-pages-in-beamers-slides
\AtBeginSection[]{
  \begin{frame}
  \vfill
  \centering
  \begin{beamercolorbox}[sep=8pt,center,shadow=true,rounded=true]{title}
    \usebeamerfont{title}\insertsectionhead\par%
  \end{beamercolorbox}
  \vfill
  \end{frame}
}

% Subsection slides: https://tex.stackexchange.com/questions/108898/beamer-presentation-show-toc-at-beginning-of-unnumbered-subsection-subsection
% \AtBeginSubsection[
%   {\frame<beamer>{\frametitle{Outline}   
%     \tableofcontents[currentsection,currentsubsection]}}%
% ]%
% {
%   \frame<beamer>{ 
%     \frametitle{Outline}   
%     \tableofcontents[currentsection,currentsubsection]}
% }

\begin{document}

\maketitle

\begin{frame}{Contents}
	\tableofcontents
\end{frame}

% First section

\section{Introduction}
\begin{frame}{Introduction}{Problem Setting}
	\begin{itemize}
		\item Observational studies frequent in econometrics, psychology, and medical research
		\item This presentation: binary treatment/control variable, $\vec{z}$.
		\item Goal is to estimate ATE: $\mathbbmss E [y_{i1} - y_{i0}]$
		\item Solve this problem via \emph{matching} treatment observations to control observations
		\item We break this down into two orthogonal problems:
			\begin{itemize}
				\item How is distance measured?
				\item How are matches assigned?
			\end{itemize}
		\item \texttt{matching}: Python package for matching observational data.
	\end{itemize}
\end{frame}

\begin{frame}{Introduction}{Notation}
	Now, we will introduce some notation, so we can precisely talk about construction of different matching methods.
	\begin{itemize}
		\item $\mat{X} \in \reals^{(n + m) \times p}$ of features; $\mat{X}_T \in \reals^{n \times p}, \mat{X}_C \in \reals^{m \times p}$
		\item $\vec{z} \in \{0, 1\}^{n + m}$ of binary treatment assignments;
		\item $d: \reals^p \times \reals^p \to \reals^+$ is a distance measure between feature vectors, e.g. $d(\vec{x}_{T_{i}}, \vec{x}_{C_{j}})$;
		\item $\mathcal{D}_d : \reals^{n \times p} \times \reals^{m \times p} \to \reals^{n \times m}$ produces \emph{biadjacency matrix};
		\item Some procedures use a preprocessing function $f$ (usually: dimension reduction or coarsening of the data).
	\end{itemize}
	Our framework is similar to that of~\textcite{iacus_multivariate_2011}:
	\begin{equation}
		\mathcal{D}_d(f(\mat{X})_T, f(\mat{X})_C). \label{eqn:beamerabstractdistance}
	\end{equation}
\end{frame}

\section{Literature Review}

\subsection{Measuring Distance}
\begin{frame}{Measuring Distance}{Balancing Scores}
	\begin{definition}[balancing score, \textcite{rosenbaum_central_1983}]
		A function $b(\mat{X}): \reals^{(n + m) \times p} \mapsto \reals^{n + m}$ is a balancing score iff%
		\begin{equation}
			\mat{X} \perp \vec{z} | b(\mat{X}). \label{eqn:beamerdefbalance}%
		\end{equation}%
	\end{definition}%
	\begin{itemize}
		\item Idea: balancing score makes observational studies more like randomized studies
		\item If \( \vec{y}_0, \vec{y}_1 \perp \vec{z} | \mat{X} \), \emph{then} \( \vec{y}_0, \vec{y}_1 \perp \vec{z} | b(\mat{X}) \)
	\end{itemize}
	These methods use $L^1$ distance by convention:
	\begin{equation}
		\mathcal{D}_{L^1}(b(\mat{X})_T, b(\mat{X})_C). \label{eqn:beamerbalancingscoredistance}
	\end{equation}
\end{frame}

\begin{frame}{Measuring Distance}{The Propensity Score}
	\begin{definition}[propensity score, \textcite{rosenbaum_central_1983}]
		The \emph{propensity score} is the balancing score \( b(\mat{X}) = \mathbbmss E [\vec{z} | \mat{X}] \).
	\end{definition}%
	\begin{itemize}
		\item Even if not used for matching, a useful diagnostic~\parencite{dehejia_causal_1999}.
		\item Implicitly ``weights'' features by heterogeneity wrt $\vec{z}$.
		\item Continuous scores: \textcite{imai_causal_2004}.
		\item Often prune potential matches with a caliper $c$.
	\end{itemize}
\end{frame}

\begin{frame}{Measuring Distance}{The Prognostic Score}
	\begin{definition}[prognostic score, \textcite{hansen_prognostic_2008}]
		The \emph{prognostic score} is the function \( b(\mat{X}) = \mathbbmss E_{C} [\vec{y} | \mat{X} ] \).
	\end{definition}
	\begin{itemize}
		\item Note that the expectation is with respect to control data $C$.
		\begin{itemize}
			\item Previous attempts to match on outcomes failed, such as \textcite{miettinen_stratification_1976}.
		\end{itemize}
		\item Implicitly ``weights'' features by heterogeneity wrt $\vec{y}$.
		\item  ``conditionality principle'' suggests usefulness \textcite{hansen_prognostic_2008}.
		\item Joint use of propensity and prognostic scores \textcite{leacy_joint_2014}.
	\end{itemize}
\end{frame}

\begin{frame}{Measuring Distance}{(Almost) Exact Matching}
	Most notable criticism of balance scores comes from~\textcite{king_why_2019}.
	\begin{itemize}
		\item ``Lower standard''.
		\item ``Paradox of Propensity Score Matching''.
	\end{itemize}
	What do they suggest instead?
	\begin{itemize}
		\item Coarsened Exact Matching (CEM)~\parencite{iacus_causal_2012}. No balance checks required!
		\item Observation weights:
			\begin{equation}
				w_{i}= \begin{cases}1, & i \in \text{treatment group} \\ \frac{m_{C}}{m_{T}} \frac{m_{T}^{s}}{m_{C}^{s}}, & i \in \text{control group} \end{cases}.
			\end{equation}
		\item ML extensions~\parencite{gupta_dame-flame_2021}: DAME~\parencite{liu_interpretable_2019}, FLAME~\parencite{wang_flame_2021}
	\end{itemize}
\end{frame}

\begin{frame}{Measuring Distance}{$L^P$ norms and Mahalanobis distance}
	Norm-based measurements are rarely used without some preprocessing function $f$.
	Why? Scaling.

	However, Mahalanobis is often used as a \emph{second} step.
	\begin{itemize}
		\item Initial filter: balancing score with a caliper. Then, match on Mahalanobis distance.
		\item ``Iterative'' approach performs well in practice~\parencite{baltar_mahalanobis_2014}. 
		\item \texttt{MatchIt} has limited support~\parencite{ho_matchit_2011}.
	\end{itemize}
\end{frame}

\begin{frame}{Bipartite Matching Algorithms}{}
	Define a matching $\set{M}$ as a set of tuples $(i, j)$ indicating a matching between observations $T_i$ and $C_j$. 
	``Optimal'' matching via the Hungarian algorithm \parencite{munkres_algorithms_1957}.
	\begin{equation}
		\min _{\mathcal{M}} \sum_{(i, j) \in \mathcal{M}} d\left(\vec{x}_{T_{i}}, \vec{x}_{C_{j}}\right), \text { s.t. } \forall(i, j),(k, l) \in \mathcal{M}, i=k \iff j=l.
	\end{equation}
	\begin{itemize}
		\item Extendable to $1:k$ matching via $b$-matching.
		\item Slow: $O((n + m)^3)$; approximation: $b$-\textsc{Suitor}~\parencite{khan_efficient_2016}.
	\end{itemize}

	Greedy matching: sort edges by distance, pop matches off the top.
	\begin{itemize}
		\item No optimality assured.
		\item Problems when matching $1:k$ \parencite{rosenbaum_optimal_1989}.
	\end{itemize}
\end{frame}

\section{\texttt{matching}: a Python Package for Matching Observational Data}
\begin{frame}{\texttt{matching}}{Design Goals}
	\begin{itemize}
		\item The main data structure used is a bipartite graph.
		\item It supports iterative distance measures of arbitrary length/complexity.
		\item It supports filtering by individual matches (edge), observation (node), and match-group/strata (subgraphs)
		\item Includes a \texttt{preprocessing} module for coarsening data and a \texttt{balance} assessment module.
	\end{itemize}
\end{frame}

\begin{frame}{\texttt{matching}}{Drawing Graphs}
	Example graphs after conducting $1:3$ optimal matching.
	\begin{center}
		\includegraphics[width=0.49\textwidth]{../paper/img/draw_examples/draw.png}
		\includegraphics[width=0.49\textwidth]{../paper/img/draw_examples/draw_bipartite.png}
	\end{center}
\end{frame}

\section{Experiments}
\begin{frame}{Experiments}{Overview}
	\begin{itemize}
		\item Hyperparameters $n$, $m$, $p$, $\theta_0$, $\theta_1$, and $\rho$.
	\begin{align}
		\vec{\upmu}_0 \sim \mathcal N(\theta_0, 0.25), \vec{\upmu}_1 \sim \mathcal N(\theta_1, 0.25) \\
		\{\vec{x}_C\}_{j=1}^{m} \sim \mathcal N(\vec{\upmu}_0, \Sigma), \{\vec{x}_T\}_{i=1}^{n} \sim \mathcal N(\vec{\upmu}_1, \Sigma) \\
		\text{ where } \Sigma_{a b} = \begin{cases}
			1 & \text{ if } a = b \\
			\rho & \text{ otherwise.}
		\end{cases}
	\end{align}
		\item Balance metrics: mean ASMD\footnote{Absolute Standardized Mean Difference}, maximum ASMD, proportion of treatment observations matched.
		\begin{itemize}
			\item In practice: use more robust metrics \parencite{basu_use_2008, zhu_kernel-based_2018}.
		\end{itemize}
		\item Vary calipers $c \in \{0.05, 0.10, 0.15, \ldots, 0.50\}$.
	\end{itemize}
\end{frame}

\subsection{Caliper vs. Imbalance}
\begin{frame}{Experiments}{Caliper vs. Imbalance}
	This will consider the effects of increased heterogeneity between $T$ and $C$.
	\begin{itemize}
		\item $m = 750$
		\item $n = 250$
		\item $p = 5$
		\item $\rho = 0$
		\item $\theta_0 = 0$
	\end{itemize}
	We vary $\theta_1 \in \{0, 0.1, 0.2, \ldots, 1\}$.
\end{frame}

\begin{frame}{Experiments}{Caliper vs. Imbalance: Mean Absolute Standardized Mean Difference}
	\begin{center}
		\includegraphics[width=\textwidth]{../paper/img/output30/caliper_vs_imbalance_big/plots/theta1_caliper_mean_abs_smd.png}
	\end{center}
\end{frame}
\begin{frame}{Experiments}{Caliper vs. Imbalance: Maximum Absolute Standardized Mean Difference}
	\begin{center}
		\includegraphics[width=\textwidth]{../paper/img/output30/caliper_vs_imbalance_big/plots/theta1_calipermax_abs_smd.png}
	\end{center}
\end{frame}
\begin{frame}{Experiments}{Caliper vs. Imbalance: Proportion of Treatment Observations Matched}
	\begin{center}
		\includegraphics[width=\textwidth]{../paper/img/output30/caliper_vs_imbalance_big/plots/theta1_caliper_prop1_match.png}
	\end{center}
\end{frame}

\subsection{Caliper vs. Correlation}
\begin{frame}{Experiments}{Caliper vs. Correlation}
	This will consider the effects of increased correlation between the features of $X$
	\begin{itemize}
		\item $m = 750$
		\item $n = 250$
		\item $p = 5$
		\item $\theta_0 = 0$
		\item $\theta_1 = 1$
	\end{itemize}
	We vary $\rho \in \{0, 0.1, 0.2, \ldots, 0.8\}$.
\end{frame}
\begin{frame}{Experiments}{Caliper vs. Correlation: Mean Absolute Standardized Mean Difference}
	\begin{center}
		\includegraphics[width=\textwidth]{../paper/img/output30/caliper_vs_correlation_big/plots/rho_caliper_mean_abs_smd.png}
	\end{center}
\end{frame}
\begin{frame}{Experiments}{Caliper vs. Correlation: Maximum Absolute Standardized Mean Difference}
	\begin{center}
		\includegraphics[width=\textwidth]{../paper/img/output30/caliper_vs_correlation_big/plots/rho_calipermax_abs_smd.png}
	\end{center}
\end{frame}
\begin{frame}{Experiments}{Caliper vs. Correlation: Proportion of Treatment Observations Matched}
	\begin{center}
		\includegraphics[width=\textwidth]{../paper/img/output30/caliper_vs_correlation_big/plots/rho_caliper_prop1_match.png}
	\end{center}
\end{frame}

\section{Conclusion}
\begin{frame}{Conclusion}
	\begin{itemize}
		\item Common theme of numerical studies: data-dependence.
		\item Emphasizes the importance of
			\begin{itemize}
				\item quality of balance metrics;
				\item inspectability of matching.
			\end{itemize}
		\item \texttt{matching} offers extreme flexibility in design of matching process.
			\begin{itemize}
				\item The current state of the graph is always inspectable.
				\item Can easily tune matches.
			\end{itemize}
	\end{itemize}
\end{frame}

\begin{frame}[fragile]
\frametitle{\texttt{matching} Example Code}
\begin{minted}[escapeinside=||, linenos, fontsize=\scriptsize]{python3}
from matplotlib import pyplot as plt
from matching.distance import Exact, L1Norm
from matching.graph import MatchingGraph
from matching.preprocessing import propensity_score

# Take `caliper_width` as given
# `X` has columns "score" and "is_nice", `z` is treatment assignments
mg = MatchingGraph(X, z)

# NOTE: there is also a keyword argument "exclude"
mg.set_edges(distance=L1Norm(max_distance=caliper_width), include=["score"])
mg.filter_edges(distance=Exact(), include=["is_nice"])

# Conduct optimal matching
mg.match(n_match=3, min_match=3, method="optimal")

# Get the matched data as a `pandas.DataFrame`
match_df = mg.match_data.frame

# Draw the graph of matches
mg.draw()
plt.show()
\end{minted}
\end{frame}

% \section*{References}
% \begin{frame}[allowframebreaks]
% 	\frametitle{References}
% 	\printbibliography
% \end{frame}

\end{document}

### 2022-05-05 paper/main.tex
updated thesis cover page
fixed typo: subset *on* features -> subset *of* features

@@
 \title{Development of a Python Package for Matching Observational Data}
-\subtitle{Advisors: Jingshu Wang, Sanjay Krishnan}
+\subtitle{{\Large Advisors: Jingshu Wang, Sanjay Krishnan}}
 \author{Jack Potrykus}
@@
 \begin{center}{\large \bfseries Master’s Paper for the University of Chicago Department of Statistics}\end{center}
+\vspace{5cm}
 
@@
 {\let\newpage\relax\maketitle}
+
+% Insert a spot for the advisor's signature
+\vfill{\null}
+{\Large 
+  Approved by:~\hrulefill\vspace{0.5cm}\\%
+  Date:~\hrulefill%
+}
+\newpage{}
 
@@
 
-% Insert a spot for the advisor's signature
-\vfill{\null}
-Approved by: \hrulefill~Date: \hrulefill
-
-\newpage{}
 \tableofcontents{}
@@
 These two algorithms are unified by their tendency to sometimes ``drop'' uninformative features from the matching process, in order to ensure that the most important features are matched on, and potentially increase the number of potential matches.
-The chief difference can be summarized as follows: DAME choosing a subset of features to match on by minimizing predictive error, whereas FLAME maximizes a match quality statistic that the researcher can tune via a hyperparameter; this statistic is a function of predictive error and a \emph{balance factor}, which represents the relative size of the strata of observations matched on this subset on features \parencite{gupta_dame-flame_2021}.
+The chief difference can be summarized as follows: DAME choosing a subset of features to match on by minimizing predictive error, whereas FLAME maximizes a match quality statistic that the researcher can tune via a hyperparameter; this statistic is a function of predictive error and a \emph{balance factor}, which represents the relative size of the strata of observations matched on this subset of features \parencite{gupta_dame-flame_2021}.
 

### 2022-05-05 paper/main.tex
last few updates

@@
 This paper breaks the matching procedure down into two key components: how distance is measured, and how matches are assigned.
-In doing so, it explores several distance metrics, in particular the propensity score \parencite{rosenbaum_central_1983}, the prognostic score \parencite{hansen_prognostic_2008}, and exact matching \parencite{iacus_causal_2012} and its machine learning extensions \parencite{liu_interpretable_2019, wang_flame_2021}; it then explores several matching algorithms that can be used to produce the matched subset once the distance is calculated, in particular the Hungarian algorithm \parencite{munkres_algorithms_1957} and greedy algorithms \parencite{ho_matchit_2011}.
+In doing so, it explores several distance metrics, in particular the propensity score~\parencite{rosenbaum_central_1983}, the prognostic score~\parencite{hansen_prognostic_2008}, and exact matching~\parencite{iacus_causal_2012} and its machine learning extensions~\parencite{liu_interpretable_2019, wang_flame_2021}; it then explores several matching algorithms that can be used to produce the matched subset once the distance is calculated, in particular the Hungarian algorithm~\parencite{munkres_algorithms_1957} and greedy algorithms~\parencite{ho_matchit_2011}.
 I then showcase the functionality of \texttt{matching}, an open-source Python package I have developed for matching observational data which takes a graph-centric approach, something which no other package offers.
@@
 Within the field of causal inference, the term \emph{matching} is often associated with observational studies with a binary treatment indicator.
-Using a similar framework to that of \textcite{iacus_multivariate_2011}, consider $\mat{X}_T \in \reals^{n \times p}$ and $\mat{X}_C \in \reals^{m \times p}$, matrices of $p$ features for $n$ treatment-group observations and $m$ control-group observations, $n \leq m$.
+Using a similar framework to that of~\textcite{iacus_multivariate_2011}, consider $\mat{X}_T \in \reals^{n \times p}$ and $\mat{X}_C \in \reals^{m \times p}$, matrices of $p$ features for $n$ treatment-group observations and $m$ control-group observations, $n \leq m$.
 We seek to produce a matching $\set{M}$ such that each observation in $\mat{X}_T$ is paired with one or more nearby observations from $\mat{X}_C$, according to a distance metric and a matching algorithm.
@@
 The choice of distance measure and choice of matching algorithm are orthogonal, and this paper (and the Python package) will discuss them as such.
-Indeed, measuring distance, as well as \emph{balance diagnostics} between treatment and control distributions, are problems most associated with the field of statistics; in contrast, producing the optimally matched subset is a \emph{minimum cost flow} problem in the field of computer science \parencite{rosenbaum_optimal_1989}.
+Indeed, measuring distance, as well as \emph{balance diagnostics} between treatment and control distributions, are problems most associated with the field of statistics; in contrast, producing the optimally matched subset is a \emph{minimum cost flow} problem in the field of computer science~\parencite{rosenbaum_optimal_1989}.
 
@@
 Distance measures are often combined with some preprocessing function $f(\mat{X})$, usually in the form of a dimension reduction or discretization of the data.
-Again using a framework akin to that of \textcite{iacus_multivariate_2011}, one may succinctly describe a distance measurement as
+Again using a framework akin to that of~\textcite{iacus_multivariate_2011}, one may succinctly describe a distance measurement as
 \begin{equation}
@@
 Balancing scores
-\footnote{Balancing scores come in a variety of forms: as \textcite{rosenbaum_central_1983} note, $b(X) = X$ is the simplest balancing score.
+\footnote{Balancing scores come in a variety of forms: as~\textcite{rosenbaum_central_1983} note, $b(X) = X$ is the simplest balancing score.
 However, this paper uses the term ``balancing score'' specifically with regards to 1-dimensional reductions of the data.
@@
 Balancing scores afford observational data some of the desirable properties of randomized trials.
-Namely, in a randomized trial, $\mat{X}$ should contain all features used in determining treatment assignment, affording conditional independence of outcomes from treatment:
+Namely, in a randomized trial, $\mat{X}$ should contain all features used in determining treatment assignment and all features known to be related to outcomes $\vec{y}_0, \vec{y}_1$; this affords conditional independence of outcomes from treatment:
 \begin{equation}
@@
 where $\vec{y}_0$ and $\vec{y}_1$ are vectors of outcomes, for which the $i$th element is the outcome of observation $i$ when assigned to group 0 (control) or 1 (treatment); note that in practice we almost always only observe $y_{0,i}$ \emph{or} $y_{1,i}$ for each $i$.
-If the statement in~\eqref{eqn:defstrongignorable} holds, then the treatment assignment is \emph{strongly ignorable}.
-\textcite{rosenbaum_central_1983} then prove that if~\eqref{eqn:defstrongignorable} holds (that is, treatment $\vec{z}$ is strongly ignorable given features $\mat{X}$), then treatment assignment is \emph{also} strongly ignorable given any balancing score $b(X)$!
-Conditioning on balancing scores, we gain independence of treatment from features, as well as the strong ignorability of treatment; these useful properties form the key motivation for the use of balancing scores in causal inference.
+If the statement in~\eqref{eqn:defstrongignorable} holds, then the treatment assignment is \emph{strongly ignorable}.~\textcite{rosenbaum_central_1983} then prove that if~\eqref{eqn:defstrongignorable} holds, then
+\begin{equation}
+  \vec{y}_0, \vec{y}_1 \perp \vec{z} | b(\mat{X})
+\end{equation}
+holds for \emph{any} balancing score $b(\mat{X})$!
+As such, the simple difference of sample means for observations at some fixed level of $b(\mat{X})$ will offer an unbiased estimate of the average treatment effect (ATE).
 
-\paragraph{The Propensity Score.} In the same paper, \textcite{rosenbaum_central_1983} propose the balancing score $b(\mat{X}) = \mathbbmss E \left [\vec{z} | \mat{X}\right ]$, also known as the propensity score.
+\paragraph{The Propensity Score.} In the same paper,~\textcite{rosenbaum_central_1983} propose the balancing score $b(\mat{X}) = \mathbbmss E \left [\vec{z} | \mat{X}\right ]$, also known as the propensity score.
 This is simply a predicted probability (or logit) that any given feature vector $\vec{x} \in \mat{X}$ belongs to the treatment group.
 This is the most prevalent balancing score in the literature, and indeed the most prevalent distance measurement overall.
-It is almost always estimated by logistic or probit regression \parencite{garrido_methods_2014}, but other classifiers are possible.
+It is almost always estimated by logistic or probit regression~\parencite{garrido_methods_2014}, but other classifiers are possible.
 By preprocessing $\mat{X}$ to a single dimension of scores $b(\mat{X})$, we not only reduce computation time necessary for $\mathcal{D}$, but also implicitly apply a weighting to the feature matrix $\mat{X}$: the features most predictive of treatment assignment (i.e., most heterogeneous between the two groups) will be the most closely matched.
@@
 
-\paragraph{Extensions of the Propensity Score.} Acknowledging that propensity score methods have thus far been limited to binary, multinomial, or ordinal treatment assignments in the literature (as they are considered in this paper), \textcite{imai_causal_2004} explored extensions of the propensity score into the realm of quantitative treatments.
+\paragraph{Extensions of the Propensity Score.} Acknowledging that propensity score methods have thus far been limited to binary, multinomial, or ordinal treatment assignments in the literature (as they are considered in this paper),~\textcite{imai_causal_2004} explored extensions of the propensity score into the realm of quantitative treatments.
 In their paper, they consider the treatment vector \emph{packyear}, the number of packs smoked by a smoker each year, and subclass the resulting scores into a varying number of bins.
@@
 
-\paragraph{The Prognostic Score.} \textcite{hansen_prognostic_2008} discussed the \emph{prognostic score}, a balancing metric on outcomes.
+\paragraph{The Prognostic Score.}~\textcite{hansen_prognostic_2008} discussed the \emph{prognostic score}, a balancing metric on outcomes.
 The only difference between the construction in~\eqref{eqn:defbalancingscore} is that we exchange $\vec{z}$ for $\vec{y}$, a vector of outcomes.
@@
 
-The idea of using outcomes, either raw or ``preprocessed'', does not see unanimous support; \textcite{garrido_methods_2014} argues against incorporating any information about outcome into the matching process.
-\textcite{miettinen_stratification_1976} was a seminal paper in matching that used outcome stratification; this method has since been found to be suboptimal, and has been effectively replaced by prognostic scores and other methods \parencite{hansen_bias_2006} (Miettinen's ``multivariate confounder score'' used the \emph{full} dataset, not just the control, when fitting).
-\textcite{hansen_bias_2006} argues for its inclusion via a ``conditionality principle'', which suggests that if some statistic (in this case, $b(\mat{X})$) is known to be uninformative about the parameter being estimated (i.e.
+The idea of using outcomes, either raw or ``preprocessed'', does not see unanimous support;~\textcite{garrido_methods_2014} argues against incorporating any information about outcome into the matching process.~\textcite{miettinen_stratification_1976} was a seminal paper in matching that used outcome stratification; this method has since been found to be suboptimal, and has been effectively replaced by prognostic scores and other methods~\parencite{hansen_bias_2006} (Miettinen's ``multivariate confounder score'' used the \emph{full} dataset, not just the control, when fitting).~\textcite{hansen_bias_2006} argues for its inclusion via a ``conditionality principle'', which suggests that if some statistic (in this case, $b(\mat{X})$) is known to be uninformative about the parameter being estimated (i.e.
 treatment effects), it should be included in inferential models.
@@
 
-\paragraph{Joint Use of Balancing Scores.} \textcite{leacy_joint_2014} further explores of calculating \emph{both} propensity and prognostic scores, stratifying each into a $p \times p$ grid, and considering observations within each of the $p^2$ grid squares as a \emph{matched strata}.
+\paragraph{Joint Use of Balancing Scores.}~\textcite{leacy_joint_2014} further explores of calculating \emph{both} propensity and prognostic scores, stratifying each into a $p \times p$ grid, and considering observations within each of the $p^2$ grid squares as a \emph{matched strata}.
 However, they found that matching on prognostic scores alone performed the best, while matching on propensity scores and Mahalanobis distance (see \S\ref{subsubsec:normdist}) also performed admirably.
@@
 \paragraph{Calipers.} Propensity and prognostic scoring methods often add one more hyperparameter: a \emph{caliper}, $c$.
-The caliper is used to calculate \emph{caliper width}, equal to the product of $c$ and the standard deviation of the scores \parencite{ho_matchit_2011}.
+The caliper is used to calculate \emph{caliper width}, equal to the product of $c$ and the standard deviation of the scores~\parencite{ho_matchit_2011}.
 The caliper serves to establish a radius of acceptable matches around each observation's propensity score as a form of quality control: if the observations are too dissimilar (their distance is greater than the caliper width), then they may not match.
@@
 
-\paragraph{Criticism.} The most notable criticism of balancing score based distance measurement comes from \textcite{king_why_2019}.
+\paragraph{Criticism.} The most notable criticism of balancing score based distance measurement comes from~\textcite{king_why_2019}.
 King and Nielsen's argument is twofold: first, from an experimental design perspective, they argue that \emph{fully blocked} study design, which other distance measures attempt to emulate (in particular: Mahalanobis distance, Coarsened Exact Matching), dominates \emph{complete randomization study design}, which balancing-score based methods attempt to emulate.
@@
 
-Among other methods which seek to replicate fully blocked study design, \textcite{king_why_2019} argue for ``coarsened exact matching'', or CEM.
+Among other methods which seek to replicate fully blocked study design,~\textcite{king_why_2019} argue for ``coarsened exact matching'', or CEM.
 This requires the researcher to first preprocess their data into discrete categories: for example, a vector ``age'' might be grouped into buckets.
@@
 \end{equation}
-where $w_i$ denotes the observation weight for observation $i$, $m_T$ and $m_C$ are the total number of matched treatment and control observations, and $m_T^s$ and $m_C^s$ are the total number of matched treatment and control observations \emph{within observation $i$'s stratum, $s$} \parencite{iacus_causal_2012}.
+where $w_i$ denotes the observation weight for observation $i$, $m_T$ and $m_C$ are the total number of matched treatment and control observations, and $m_T^s$ and $m_C^s$ are the total number of matched treatment and control observations \emph{within observation $i$'s stratum, $s$}~\parencite{iacus_causal_2012}.
 
@@
 
-\paragraph{Almost Exact Matching.} Some of the most modern methods are DAME \parencite{liu_interpretable_2019} and FLAME \cite{wang_flame_2021}, which incorporate machine-learning generated weight vectors indicating the relative importance of features to extend exact matching into the field of \emph{almost exact matching}.
-These two algorithms are unified by their tendency to sometimes ``drop'' uninformative features from the matching process, in order to ensure that the most important features are matched on, and potentially increase the number of potential matches.
-The chief difference can be summarized as follows: DAME choosing a subset of features to match on by minimizing predictive error, whereas FLAME maximizes a match quality statistic that the researcher can tune via a hyperparameter; this statistic is a function of predictive error and a \emph{balance factor}, which represents the relative size of the strata of observations matched on this subset of features \parencite{gupta_dame-flame_2021}.
+\paragraph{Almost Exact Matching.} Some of the most modern methods are DAME~\parencite{liu_interpretable_2019} and FLAME \cite{wang_flame_2021}, which incorporate machine-learning generated weight vectors indicating the relative importance of features to extend exact matching into the field of \emph{almost exact matching}.
+Each of these algorithms incorporates machine learning to compute a weight vector $\vec{w}$, indicating the relative importance of exact-matching on each feature.
+The algorithms each search for full exact matches first; when this is no longer possible, the weight vector is used to determine a subset of the features to search for exact matches on.
+The key difference is that DAME choses a subset the features to match on by minimizing predictive error, whereas FLAME maximizes a match quality statistic that the researcher can tune via a hyperparameter; this statistic is a function of predictive error and a \emph{balance factor}, which represents the relative size of the strata of observations matched on this subset of features~\parencite{gupta_dame-flame_2021}.
 
@@
 
-Norms of the difference between feature vectors (i.e.
-$\| \vec{x}_T - \vec{x}_C \|_p$) themselves are usually used in the form of Mahalanobis distance, which first rescales all features to have mean zero and unit variance, and then calculates distances using the $L^2$ norm.
+Another valid distance measure is the $L^p$ norms of the difference between feature vectors, i.e. $\| \vec{x}_T - \vec{x}_C \|_p$. 
+As~\textcite{king_why_2019} note, norm-based distance metrics attempt to replicate fully-blocked design, and thus offer some theoretical advantages.
+However, they do not incorporate any weighting about a feature's ``importance'', as we saw with balancing scores and almost exact matching.
+Hence, norms are most often used in the form of Mahalanobis distance, which first rescales all features to have mean zero and unit variance, and then calculates distances using the $L^2$ norm.
 This ensures that all features have equal weight in the matching, which may or may not be desirable; in either case, it almost certainly better than the no-rescaling case, where a feature's variance would determine its relative importance.
-As \textcite{king_why_2019} notes, norm-based distance metrics attempt to replicate fully-blocked design, and thus offer some theoretical advantages.
-However, many are attracted to the ability of balancing score methods, as well as modern methods such as DAME and FLAME, to discern which features are \emph{most important} to match on.
 
-Nonetheless, \textcite{leacy_joint_2014} finds Mahalanobis distance to perform well when the number of features was small, each of which was approximately normally distributed.
-Additionally, many modern matching software packages allow for initially establishing a ``perimeter'' of viable matches using a caliper and a propensity score, and subsequently matching on the Mahalanobis distance between the feature vectors \parencite{ho_matchit_2011}.
+\textcite{leacy_joint_2014} finds Mahalanobis distance to perform well when the number of features was small, each of which was approximately normally distributed.
+Additionally, many modern matching software packages allow for initially establishing a ``perimeter'' of viable matches using a caliper and a propensity score, and subsequently matching on the Mahalanobis distance between the feature vectors~\parencite{ho_matchit_2011}.
 
@@
 They may, for example, first calculate propensity scores $\widehat{\vec{z}}$, filter potential matches using a caliper $c$, and then match on Mahalanobis distance \emph{within} clusters; or, in the absence of a caliper, within strata of $\widehat{\vec{z}}$.
-This idea was explored by \textcite{rosenbaum_constructing_1985}, who found it to perform better than matching on either propensity score or Mahalanobis distance alone.
+This idea was explored by~\textcite{rosenbaum_constructing_1985}, who found it to perform better than matching on either propensity score or Mahalanobis distance alone.
 
@@
 The iterative method achieved the greatest reduction in bias for all features, except the propensity score, for which propensity score distance alone (somewhat obviously) saw the greatest bias reduction.
-However, as \textcite{austin_optimal_2011} and \textcite{greifer_assessing_2022} note, balance of the scores does not imply balance of the features, and balancing score balance metrics should only be used as a supplementary diagnostic; the focus of balance assessment should be balance of the features.
+However, as~\textcite{austin_optimal_2011} and~\textcite{greifer_assessing_2022} note, balance of the scores does not imply balance of the features, and balancing score balance metrics should only be used as a supplementary diagnostic; the focus of balance assessment should be balance of the features.
 
@@
 Here, the term ``bipartite'' means that all edges in the graph map $T \leftrightarrow C$; within each disjoint subset, no two nodes share an edge.
-This paper will explore the optimal solution to this problem in both the $1:1$ and $1:k$ matching scenarios, as well as fast approximations.
 
@@
 
-\textcite{rosenbaum_optimal_1989} argues for optimal matching according to the Hungarian algorithm, also known as the Kuhn-Munkres or simply Munkres algorithm \parencite{munkres_algorithms_1957}.
+\textcite{rosenbaum_optimal_1989} argues for optimal matching according to the Hungarian algorithm, also known as the Kuhn-Munkres or simply the Munkres algorithm~\parencite{munkres_algorithms_1957}.
 We will denote a matching a set $\set{M}$ of tuples $(i, j)$, with $i$ representing a treatment-group observations, and $j$ a control-group observations.
@@
 This process involves adding $k$ ``copy'' nodes for each node representing a $T$ datapoint, each with identical edges as the original node, and then perform the optimal matching on the resulting graph.
-However, as \textcite{khan_efficient_2016} notes, this can lead to calculations which quickly become time consuming as the total number of data points increases.
-\textcite{khan_efficient_2016} goes on further to explore efficient approximation algorithms of the $b$-matching, such as the $b$-\textsc{Suitor} algorithm.
+However, as~\textcite{khan_efficient_2016} notes, this can lead to calculations which quickly become time consuming as the total number of data points increases.~\textcite{khan_efficient_2016} goes on further to explore efficient approximation algorithms of the $b$-matching, such as the $b$-\textsc{Suitor} algorithm.
 
@@
 
-Greedy algorithms have the benefit of running much quicker than optimal matching algorithms, and are often the default in other observational data matching software, such as \texttt{MatchIt} \parencite{ho_matchit_2011}.
+Greedy algorithms have the benefit of running much quicker than optimal matching algorithms, and are often the default in other observational data matching software, such as \texttt{MatchIt}~\parencite{ho_matchit_2011}.
 These methods are equivalent to sorting the edges of the graph in ascending order, and naively popping off the top of the list the next match.
 While these methods benefit from speed, they guarantee nothing with regards to optimality.
-This can be particularly problematic in the $1:k$ scenario \parencite{rosenbaum_optimal_1989}, where the order in matches are assigned can have matches implications; compare this to optimal matching, which is a simultaneous match.
+This can be particularly problematic in the $1:k$ scenario~\parencite{rosenbaum_optimal_1989}, where the order in matches are assigned can have matches implications; compare this to optimal matching, which is a simultaneous match.
 
@@
 
-Researchers may wish to assess the balance of their matching, particularly when attempting to heuristically optimize a hyperparameter, such as the caliper.
-Common numerical methods include standardized mean-differences, variance ratios, and empirical CDF (eCDF) comparison between the two groups \parencite{greifer_assessing_2022}.
-Most researchers do not check for convergence of moments beyond the second, and as \textcite{basu_use_2008} report, these moments exhibit much slower convergence, the effects of which can be particularly detrimental when outcome depends on some non-linear function of the joint distribution of $\mat{X}$.
-For this reason, \textcite{zhu_kernel-based_2018} recommend using density-based metrics, such as total variation distance or Kolmogorov-Smirnov distance.
-It is also important to note that balance of the $i$th moment does not imply balance of the $j$th moment across the two groups for $j \geq i$ \parencite{garrido_methods_2014}.
+Researchers should be sure to assess the balance of their matching, particularly when attempting to heuristically optimize a hyperparameter, such as the caliper.
+Common numerical methods include standardized mean-differences, variance ratios, and empirical CDF (eCDF) comparison between the two groups~\parencite{greifer_assessing_2022}.
+Most researchers do not check for convergence of moments beyond the second; it is important to note that balance of the $i$th moment does not imply balance of the $j$th moment across the two groups for $j \geq i$~\parencite{garrido_methods_2014}.
+Furthermore, as~\textcite{basu_use_2008} report, higher moments exhibit much slower convergence in balance, which can have particularly detrimental effects when outcome depends on some non-linear function of the joint distribution of $\mat{X}$.
+For this reason,~\textcite{zhu_kernel-based_2018} recommend using density-based metrics, such as total variation distance or Kolmogorov-Smirnov distance, to assess balance.
 
 There are also many graphical diagnostics for assessing balance.
-QQ-plots, eCDFs, and density estimates for for the unmatched data and the matched subset, perhaps overlaid, effectively visually convey the balance improvement from matching.
-One may also plot the improvement in standardized mean differences between the two groups, referred to as a Love plot \parencite{greifer_assessing_2022}.
+QQ-plots, eCDFs, and density estimates for the unmatched data and the matched subset, perhaps overlaid, effectively visually convey the balance improvement from matching~\parencite{greifer_assessing_2022}.
 One could also plot the bipartite graph to understand where clusters are appearing, or how large groups of ``similar'' observations are, or which observations are most ``dissimilar'' from the rest.
@@
 
-The most popular implementation of bipartite matching for observational studies is the \texttt{MatchIt} package for R \parencite{ho_matchit_2011}.
-The GitHub repository\footnote{\url{https://github.com/kosukeimai/MatchIt}} boasts an impressive $37000$ downloads per month, and with good reason: the authors of the package are also authors of many of the most popular papers in the field of matching (e.g.
-\textcite{ho_matching_2007, imai_causal_2004, king_why_2019, stuart_matching_2010}).
-
-\texttt{MatchIt} is largely used through a single function: \texttt{Matchit::matchit}, which minimally accepts a \texttt{formula} (like those used in \texttt{lm}), \texttt{data}, and a \texttt{method}.
-The function also allows for iterative matching procedures, complex filters on the suitability of potential matches, and a selection of bipartite matching algorithms.
-The output of the function is furthermore easily inspected via the ubiquitous \texttt{summary} function, or plotted via \texttt{plot}.
+The most popular implementation of bipartite matching for observational studies is the \texttt{MatchIt} package for R~\parencite{ho_matchit_2011}.
+The GitHub repository\footnote{\url{https://github.com/kosukeimai/MatchIt}} boasts an impressive $37000$ downloads per month, and with good reason: the authors of the package are also authors of many of the most popular papers in the field of matching (e.g.~\textcite{ho_matching_2007, imai_causal_2004, king_why_2019, stuart_matching_2010}).
 
@@
 \end{figure}
-
 \begin{figure}[h!]
   \center{%
-    \includegraphics[width=\textwidth]{../docs/mermaid/img/distance_uml.png}%
+    \includegraphics[width=0.9\textwidth]{../docs/mermaid/img/distance_uml.png}%
   }
@@
 
-\paragraph{\texttt{DoWhy}.} An ``End-to-End Library for Causal Inference'' developed by Microsoft \parencite{dowhypaper}, \texttt{DoWhy} supports matching on a variety of distance metrics, in addition to numerous other methods in the field of Causal Inference (e.g.
+\paragraph{\texttt{DoWhy}.} An ``End-to-End Library for Causal Inference'' developed by Microsoft~\parencite{dowhypaper}, \texttt{DoWhy} supports matching on a variety of distance metrics, in addition to numerous other methods in the field of Causal Inference (e.g.
 instrumental variables), and can match the data, build a predictive model for outcomes, and estimate the treatment effect in one shot.
@@
 
-\paragraph{\texttt{pymatch}.} Purpose-built by researchers working on an observational study \parencite{miroglio_pymatch_2022}, \texttt{pymatch} supports only one distance metric, the propensity score, and the only matching algorithms on offer are ``random'' and ``greedy''; no optimal matching is available.
+\paragraph{\texttt{pymatch}.} Purpose-built by researchers working on an observational study~\parencite{miroglio_pymatch_2022}, \texttt{pymatch} supports only one distance metric, the propensity score, and the only matching algorithms on offer are ``random'' and ``greedy''; no optimal matching is available.
 The package also lacks support iterative matching procedures.
@@
 
-\paragraph{\texttt{dame-flame}.} Developed by Duke's \emph{Almost Matching Exactly Lab} \parencite{gupta_dame-flame_2021}, \texttt{dame-flame} implements its two titular matching procedures: DAME \parencite{liu_interpretable_2019} and FLAME \parencite{wang_flame_2021}.
+\paragraph{\texttt{dame-flame}.} Developed by Duke's \emph{Almost Matching Exactly Lab}~\parencite{gupta_dame-flame_2021}, \texttt{dame-flame} implements its two titular matching procedures: DAME~\parencite{liu_interpretable_2019} and FLAME~\parencite{wang_flame_2021}.
 As discussed in \S\ref{subsubsec:exactmatching}, these methods extend CEM using machine learning, calculating a weight vector \vec{w} indicating the relative importance of matching on each feature.
@@
 
-The data were generated according to the following hyperparameters. This follows similar frameworks in the literature, such as those of \textcite{austin_optimal_2011}.
+The data were generated according to the following hyperparameters. This follows similar data-generating frameworks in the literature, such as those of~\textcite{austin_optimal_2011}.
 \begin{itemize}
@@
 \subsection{Experiment 1: Caliper vs Imbalance.}
-This experiment seeks to investigate the practical truth of the ``Paradox of Propensity Score Matching'' \parencite{king_why_2019}.
+This experiment seeks to investigate the practical truth of the ``Paradox of Propensity Score Matching''~\parencite{king_why_2019}.
 The Paradox suggests that matching quality could (not necessarily \emph{must}) deteriorate as the imbalance between the treatment and control groups decreases; this is a logical extrapolation from that earlier-noted fact that when $\mat{X} \perp \vec{z}$, the classifier will assign a constant propensity score to all observations.
@@
 This experiment seeks to assess the impact that varying correlation among the features has on selection of an optimal caliper.
-Specifically, it hypothesizes that the optimal choice of caliper will vary significantly as correlation changes; this would be in line with results of the aforementioned \textcite{austin_optimal_2011}.
+Specifically, it hypothesizes that the optimal choice of caliper will vary significantly as correlation changes; this would be in line with results of the aforementioned~\textcite{austin_optimal_2011}.
 
@@
 This analysis suggests that propensity score matching was able to produce high-quality matches even when the data generating model consisted of highly correlated features.
-This is in line with results presented in other research, such as \textcite{austin_optimal_2011}.
+This is in line with results presented in other research, such as~\textcite{austin_optimal_2011}.
 
@@
 In the case of the experiments conducted in \S\ref{sec:experiments}, mean absolute standardized mean difference was useful because we knew the true difference between $\theta_1$ and $\theta_0$; in practice, this quantity is unknown.
-As such, researchers should use robust methods of balance assessment; for example, checking for convergence of higher-order moments, or using density-based metrics such as those of \textcite{zhu_kernel-based_2018}.
+As such, researchers should use robust methods of balance assessment; for example, checking for convergence of higher-order moments, or using density-based metrics such as those of~\textcite{zhu_kernel-based_2018}.
 But even then, one cannot possibly hope to analyze the balance of the \emph{unobserved} confounding features; for these, we are only promised balance \emph{on average}.

### 2022-05-05 presentation/presentation.tex
last few updates

@@
 	\begin{itemize}
-		\item Observational studies frequent in econometrics, psychology, and medical research
+		\item Observational studies frequent in econometrics, psychology, and medical research.
 		\item This presentation: binary treatment/control variable, $\vec{z}$.
-		\item Goal is to estimate ATE: $\mathbbmss E [y_{i1} - y_{i0}]$
+		\item Most often, goal is to estimate ATE: $\mathbbmss E [y_{i1} - y_{i0}]$.
 		\item Solve this problem via \emph{matching} treatment observations to control observations
+			\begin{itemize}
+				\item Produce a matched subset suitable for analysis
+			\end{itemize}
 		\item We break this down into two orthogonal problems:
@@
 	\begin{itemize}
-		\item $\mat{X} \in \reals^{(n + m) \times p}$ of features; $\mat{X}_T \in \reals^{n \times p}, \mat{X}_C \in \reals^{m \times p}$
+		\item $\mat{X} \in \reals^{(n + m) \times p}$ of features; $\mat{X}_T \in \reals^{n \times p}, \mat{X}_C \in \reals^{m \times p}$, $n \leq m$;
 		\item $\vec{z} \in \{0, 1\}^{n + m}$ of binary treatment assignments;
@@
 	\begin{itemize}
+		\item Implicitly ``weights'' features by heterogeneity wrt $\vec{z}$.
+		\item Often prune potential matches with a caliper $c$.
 		\item Even if not used for matching, a useful diagnostic~\parencite{dehejia_causal_1999}.
-		\item Implicitly ``weights'' features by heterogeneity wrt $\vec{z}$.
 		\item Continuous scores: \textcite{imai_causal_2004}.
-		\item Often prune potential matches with a caliper $c$.
 	\end{itemize}
@@
 		\begin{itemize}
+			\item No information about ATE
 			\item Previous attempts to match on outcomes failed, such as \textcite{miettinen_stratification_1976}.
+			\item  ``conditionality principle'' \textcite{hansen_prognostic_2008}.
 		\end{itemize}
 		\item Implicitly ``weights'' features by heterogeneity wrt $\vec{y}$.
-		\item  ``conditionality principle'' suggests usefulness \textcite{hansen_prognostic_2008}.
-		\item Joint use of propensity and prognostic scores \textcite{leacy_joint_2014}.
 	\end{itemize}
@@
 
+\subsection{Matching Algorithms}
 \begin{frame}{Bipartite Matching Algorithms}{}
@@
 		\item Extendable to $1:k$ matching via $b$-matching.
-		\item Slow: $O((n + m)^3)$; approximation: $b$-\textsc{Suitor}~\parencite{khan_efficient_2016}.
+		\item Slow: $O((k \cdot n + m)^3)$; approximation: $b$-\textsc{Suitor}~\parencite{khan_efficient_2016}.
 	\end{itemize}
@@
 	\begin{itemize}
-		\item No optimality assured.
+		\item When control observations may not be reused, no optimality assured.
 		\item Problems when matching $1:k$ \parencite{rosenbaum_optimal_1989}.
@@
 	\end{align}
-		\item Balance metrics: mean ASMD\footnote{Absolute Standardized Mean Difference}, maximum ASMD, proportion of treatment observations matched.
+		\item Metrics: mean ASMD\footnote{Absolute Standardized Mean Difference}, proportion of treatment observations matched.
 		\begin{itemize}
@@
 \end{frame}
-\begin{frame}{Experiments}{Caliper vs. Imbalance: Maximum Absolute Standardized Mean Difference}
-	\begin{center}
-		\includegraphics[width=\textwidth]{../paper/img/output30/caliper_vs_imbalance_big/plots/theta1_calipermax_abs_smd.png}
-	\end{center}
-\end{frame}
 \begin{frame}{Experiments}{Caliper vs. Imbalance: Proportion of Treatment Observations Matched}
@@
 		\includegraphics[width=\textwidth]{../paper/img/output30/caliper_vs_correlation_big/plots/rho_caliper_mean_abs_smd.png}
-	\end{center}
-\end{frame}
-\begin{frame}{Experiments}{Caliper vs. Correlation: Maximum Absolute Standardized Mean Difference}
-	\begin{center}
-		\includegraphics[width=\textwidth]{../paper/img/output30/caliper_vs_correlation_big/plots/rho_calipermax_abs_smd.png}
 	\end{center}

### 2022-05-06 paper/main.tex
typo in literally the second sentence lol

@@
 Observational studies are of ever-increasing importance as the numerous amount of data collected each year continues to grow.
-However, of observational studies are at a major disadvantage compared to blocking-based study design or randomized control trials, because the researchers do not have control over the distribution of confounding variables, known or unknown, in the data itself.
+However, observational studies are at a major disadvantage compared to blocking-based study design or randomized control trials, because the researchers do not have control over the distribution of confounding variables, known or unknown, in the data itself.
 As such, researchers must take preliminary steps to alter the data in some way in order to minimize the effects of selection bias, which results from heterogeneous distributions present in the observational data.

### 2022-05-06 paper/main.tex
more small edits

@@
 
-\paragraph{The Prognostic Score.}~\textcite{hansen_prognostic_2008} discussed the \emph{prognostic score}, a balancing metric on outcomes.
+\paragraph{The Prognostic Score.}~\textcite{hansen_prognostic_2008} proposed the \emph{prognostic score}, a balancing metric on outcomes.
 The only difference between the construction in~\eqref{eqn:defbalancingscore} is that we exchange $\vec{z}$ for $\vec{y}$, a vector of outcomes.
@@
 \end{equation}
-The prognostic score is constructed nearly in exactly the same way as the propensity score; however, in order to preserve~\eqref{eqn:prognosticscoredef}, the model (linear/logistic regression, random forest, etc.) is trained on the \emph{control data alone}.
-This model is then used to predict scores over the \emph{whole dataset}.
+The prognostic score is constructed nearly in exactly the same way as the propensity score; however, in order to ensure no information about the ATE is encoded in the score, the prognostic score is defined as $\mathbbmss E_C [\vec{y} | \mat{X} ]$.
+In practice, this means fitting the model using only $\mat{X}_C, \vec{y}_C$, and then ``predicting'' over all $\mat{X}$.
 
-The idea of using outcomes, either raw or ``preprocessed'', does not see unanimous support;~\textcite{garrido_methods_2014} argues against incorporating any information about outcome into the matching process.~\textcite{miettinen_stratification_1976} was a seminal paper in matching that used outcome stratification; this method has since been found to be suboptimal, and has been effectively replaced by prognostic scores and other methods~\parencite{hansen_bias_2006} (Miettinen's ``multivariate confounder score'' used the \emph{full} dataset, not just the control, when fitting).~\textcite{hansen_bias_2006} argues for its inclusion via a ``conditionality principle'', which suggests that if some statistic (in this case, $b(\mat{X})$) is known to be uninformative about the parameter being estimated (i.e.
+The idea of using outcomes, either raw or ``preprocessed'', does not see unanimous support;~\textcite{garrido_methods_2014} argues against incorporating any information about outcome into the matching process.~\textcite{miettinen_stratification_1976} was a seminal paper in matching that used outcome stratification; this method has since been found to be suboptimal, and has been effectively replaced by prognostic scores and other methods~\parencite{hansen_bias_2006} (Miettinen's ``multivariate confounder score'' used the \emph{full} dataset, not just the control, when fitting).~\textcite{hansen_bias_2006} argues for the prognostic score's inclusion via a ``conditionality principle'', which suggests that if some statistic (in this case, $b(\mat{X})$) is known to be uninformative about the parameter being estimated (i.e.
 treatment effects), it should be included in inferential models.

### 2022-05-06 paper/main.tex
more small edits

@@
 These methods take a vector of tuning parameters $\vec{\uppi}$, such as a vector of calipers.
-These methods afford the benefit that the number of matches is a function of the tuning parameters alone, and given monotonicity of the maximum distance function $\gamma_{\mathcal{D}_{d}}(\vec{\uppi})$, can be each adjusted independently other others: if $\gamma$ is vector-valued, the element-wise differences of feature vectors can be independently compared in a system of inequalities.
+The benefit is that the number of matches is a function of the tuning parameters alone, and given monotonicity of the maximum distance function $\gamma_{\mathcal{D}_{d}}(\vec{\uppi})$, each entry of $\vec{\uppi}$ can be each adjusted independently other others: if $\gamma$ is vector-valued, the element-wise differences of feature vectors can be independently compared in a system of inequalities.
 CEM is an example of an MIB distance measure, but this class is clearly much more flexible, allowing for precise adjustments.

### 2022-05-06 paper/main.tex
more small edits

@@
 \textcite{iacus_multivariate_2011} introduces a formal definition for a class of distance measures which they call ``monotonic imbalance bounding'', or MIB.
-These methods are designed to require no assumptions about the distribution of the data, and to emphasize minimizing \emph{in-sample} imbalance, as opposed to \emph{sample} balance.
+These methods are designed to require no assumptions about the distribution of the data, and to emphasize minimizing \emph{in-sample} imbalance, as opposed to \emph{expected} balance.
 These methods take a vector of tuning parameters $\vec{\uppi}$, such as a vector of calipers.
 The benefit is that the number of matches is a function of the tuning parameters alone, and given monotonicity of the maximum distance function $\gamma_{\mathcal{D}_{d}}(\vec{\uppi})$, each entry of $\vec{\uppi}$ can be each adjusted independently other others: if $\gamma$ is vector-valued, the element-wise differences of feature vectors can be independently compared in a system of inequalities.
-CEM is an example of an MIB distance measure, but this class is clearly much more flexible, allowing for precise adjustments.
+CEM is an example of an MIB distance measure, but this class is clearly much more flexible, allowing for precise adjustments of hyperparameters specific to each feature.
 

### 2022-05-06 paper/main.tex
more small edits

@@
 
-\paragraph{Almost Exact Matching.} Some of the most modern methods are DAME~\parencite{liu_interpretable_2019} and FLAME \cite{wang_flame_2021}, which incorporate machine-learning generated weight vectors indicating the relative importance of features to extend exact matching into the field of \emph{almost exact matching}.
+\paragraph{Almost Exact Matching.} Some of the most modern methods are DAME~\parencite{liu_interpretable_2019} and FLAME~\parencite{wang_flame_2021}.
 Each of these algorithms incorporates machine learning to compute a weight vector $\vec{w}$, indicating the relative importance of exact-matching on each feature.

### 2022-05-06 paper/main.tex
more small edits

@@
 
-Another valid distance measure is the $L^p$ norms of the difference between feature vectors, i.e. $\| \vec{x}_T - \vec{x}_C \|_p$. 
+A rather obvious distance measure is the $L^p$ norm of the difference between feature vectors, i.e. $\| \vec{x}_T - \vec{x}_C \|_p$. 
 As~\textcite{king_why_2019} note, norm-based distance metrics attempt to replicate fully-blocked design, and thus offer some theoretical advantages.
@@
 \textcite{baltar_mahalanobis_2014} explored the efficacy of this iterative method in practice, relative to propensity score and Mahalanobis distance alone.
-In the paper, they compare seek to estimate the effect of Brazilian towns' social agenda on public health.
+In the paper, they seek to estimate the effect of Brazilian towns' social agenda on public health.
 They found both the propensity score alone and iterative method to perform admirably, whereas Mahalanobis distance performed poorly; they do not take this as an indictment of Mahalanobis distance matching, and pose for further research to identify conditions under which Mahalnobis distance may perform \emph{better}.
@@
 \textcite{rosenbaum_optimal_1989} argues for optimal matching according to the Hungarian algorithm, also known as the Kuhn-Munkres or simply the Munkres algorithm~\parencite{munkres_algorithms_1957}.
-We will denote a matching a set $\set{M}$ of tuples $(i, j)$, with $i$ representing a treatment-group observations, and $j$ a control-group observations.
+We will denote a matching a set $\set{M}$ of tuples $(i, j)$, with $i$ representing a treatment-group observation $T_i$, and $j$ a control-group observation $C_j$.
 The Hungarian algorithm solves the \emph{linear assignment problem}, which in this setting may be expressed as
@@
 This process involves adding $k$ ``copy'' nodes for each node representing a $T$ datapoint, each with identical edges as the original node, and then perform the optimal matching on the resulting graph.
-However, as~\textcite{khan_efficient_2016} notes, this can lead to calculations which quickly become time consuming as the total number of data points increases.~\textcite{khan_efficient_2016} goes on further to explore efficient approximation algorithms of the $b$-matching, such as the $b$-\textsc{Suitor} algorithm.
+However, as~\textcite{khan_efficient_2016} notes, this can lead to calculations which quickly become time consuming as the total number of data points increases; the runtime is $O((k \cdot n + m)^3)$.~\textcite{khan_efficient_2016} goes on further to explore efficient approximation algorithms of the $b$-matching, such as the $b$-\textsc{Suitor} algorithm.
 

### 2022-05-06 paper/main.tex
more small edits

@@
 While these methods benefit from speed, they guarantee nothing with regards to optimality.
-This can be particularly problematic in the $1:k$ scenario~\parencite{rosenbaum_optimal_1989}, where the order in matches are assigned can have matches implications; compare this to optimal matching, which is a simultaneous match.
+This can be particularly problematic in the $1:k$ scenario~\parencite{rosenbaum_optimal_1989}, where the order in matches are assigned can have huge implications on the overall balance; compare this to optimal matching, which is a simultaneous match.
 

### 2022-05-06 paper/main.tex
more small edits

@@
   \center{%
-    \includegraphics[width=0.8\textwidth]{../docs/mermaid/img/relationships.png}%
+    \includegraphics[width=0.7\textwidth]{../docs/mermaid/img/relationships.png}%
   }
@@
   \center{%
-    \includegraphics[width=0.9\textwidth]{../docs/mermaid/img/distance_uml.png}%
+    \includegraphics[width=0.8\textwidth]{../docs/mermaid/img/distance_uml.png}%
   }
@@
 \end{minted}
-\paragraph{Setting Edges.} The user should then make a call to the \texttt{set\_edges} method to calculate the distance between each observation the two groups, and set the graph nodes and edge weights.
+\paragraph{Setting Edges.} The user should then make a call to the \texttt{set\_edges} method to calculate the distance between each observation in the two groups, and set the graph nodes and edge weights.
 Here, the user supplies a \texttt{\_Distance} measure from \texttt{matching.distance}.
@@
 
-\paragraph{Matching.} Once the user has set the edges of the graph and completed their filtering, they may wish to conduct a $1:k$ matching between treatment and control, for some integer $k \geq 1$\footnote{This is not always necessary, particularly in the case of CEM.}
+\paragraph{Matching.} Once the user has set the edges of the graph and completed their filtering, they may wish to conduct a $1:k$ matching between treatment and control, for some integer $k \geq 1$\footnote{This is not always necessary, particularly in the case of CEM and other strata-based matching methods}.
 The user does this via a call to the \texttt{match} method.
@@
 
+\newpage
 \section{Experiments}
@@
 This heatmap makes clear that when heterogeneity of the features is too large (even just a single standard deviation apart), the proportion of treatment observations matched may decrease substantially when using the same caliper $c$ (here, we see a decrease of 50\% for $c = 0.05$).
+A common ``rule-of-thumb'' desired level of balance is $0.1$ absolute standardized mean difference, which is achieved by larger calipers which match more of the data.
+Researchers will have to consider this practical trade off between balance and the size of the matched subset.
 \paragraph{Analysis of Subfigure~\ref{subfig:exp1_mean_abs_smd}.} First, we note that the mean absolute standardized mean difference is nearly constant down the leftmost column (with calipers of $c=0.05$); the cost of imbalance was primarily paid in the size of the matched subset.
@@
 
+\newpage
 \subsection{Experiment 2: Caliper vs Correlation.}
@@
 $\rho=0.3$); this is no doubt correlated with the high-proportion of treatment group observations matched for those same values of $\rho$.
-
-This analysis suggests that propensity score matching was able to produce high-quality matches even when the data generating model consisted of highly correlated features.
+This experiment suggests that the propensity score, in combination with a reasonable caliper $c$, is a useful and robust matching metric under the correlated data regimes.
 This is in line with results presented in other research, such as~\textcite{austin_optimal_2011}.
